### 2023-05-02 paper/main.tex

\documentclass[twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{natbib}
\usepackage{graphicx, caption}
\usepackage{authblk}
\usepackage{textcomp}

% \usepackage{moreverb,savetrees}

\usepackage{multicol}
\usepackage[margin=0.6in]{geometry}

\graphicspath{{./figures/}{./}}

\title{Title}
\author[1]{PALAGRAM}

\affil[1]{Affiliations}

\date{}

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

\begin{document}

\twocolumn[
	\maketitle

	\begin{abstract}
		Particle-laden gravity currents (PLGCs) are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid. We focus our attention on finite volume current, generated here by lock-release device over a controlled constant slope bottom at the laboratory scale. The main objective of the present chapter is to provide a comprehensive flow regime associated with the relevant dimensionless parameters which characterize the current at the macroscopic scale, i.e. average over the current depth.
	\end{abstract}
	\vskip 0.7cm
]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 1
\begin{figure*}[h]
	\centering
	\includegraphics{figure1.pdf}
	\caption{Caption 1}
	\label{fig:fig1}
\end{figure*}

\section{Introduction}

\subsection{Control parameters}

Particle-laden gravity currents (PLGCs) are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid. A major difference between compositional currents and PLGCs is therefore the existence of typical length and time scales associated with the particle dynamics. Particularly, the own settling velocity of a heavy particle only allows the existence of these gravity currents on finite time and length scales. Nevertheless, prediction of these finite scales is difficult as it shall strongly depends on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface, which modifies entrainment and associated dissipation, and in turn can delay or enhance settling. Accordingly, their dynamics are controlled by several dimensionless parameters leading to a wide variety of flow regimes. In particular, beyond the slope $\alpha$, the dimensionless density difference between the current and the ambient $At$ and the ratio between available inertia and dissipation, say a Reynolds number $Re$, as usually introduced as a control parameter for compositional currents, the initial solid fraction of particles $\phi$ and a dimensionless settling velocity St have now to be considered as potentially affecting flow regimes. It shall be noted that in some specific situations, some of these dimensionless parameters could be linked as, for instance, $\phi$ and $At$ for heavy particles in a single-phase liquid for the carrying and ambient fluids

\subsection{Literature}

\begin{itemize}
	\item literature slope
	\item literature settling
	\item literature volume fraction
	\item other effects: lock geometry, non-boussinesq, entrainment, erosion/deposition $\rightarrow$ ignored here
\end{itemize}

\section*{Methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 2
\begin{figure}
	\centering
	% \includegraphics{Figures/Figure_lit_plat/Figure_lit_plat.pdf}
	\includegraphics{figure2.pdf}
	\caption{Caption.}
	\label{fig:fig2}
\end{figure}

\subsection{Experimental setups}

Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed on the front dynamics, often referred to as the $\mathcal{F}r$ number, and the front shape. Then prediction of the different flow regimes associated with PLGCs shall be entirely included into these observables. In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at “rest”, 3 dam-break-type devices have been used to cover a wide range of control parameters, as $\alpha \in [0^o,30^o]$, $\phi\in[0.1\%,20\%]$, $At\in[???]$, $Re\in[???]$, $St\in[???]$, $a\in??$, as specifically defined in section \ref{sec:dimensionlessmap}. These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.

\subsection{Datasets}

On these 3 devices, a total of $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below). They are classified into 4 different datasets, described below.

\paragraph{Dataset 1}

Description Dataset LEGI

\paragraph{Dataset 2}

Description Dataset IMFT

\paragraph{Dataset 3}

Description Dataset LEGI/IMFT

\paragraph{Dataset 4}

Description Dataset LEMTA

\paragraph{SedFoam}

Description Dataset SedFoam

\subsection{Fitting procedure and front position analysis}

For each experiment/simulation, we extract the current front position as a function of time (see Fig. \ref{fig:fig3}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 3
\begin{figure*}
	\centering
	\includegraphics{figure3.pdf}
	\caption{Caption}
	\label{fig:fig3}
\end{figure*}

\section{Dimensional analysis at the macroscopic scale of the current}
\label{sec:dimensionlessmap}

Here, the dimensionless parameters are based on a characteristic length $L_0$ and a characteristic velocity $U_0=\sqrt{g' H_0}$ where $L_0$ and $H_0$ are the horizontal length and depth of the initial container, respectively, $g'$ is the modified gravity $r g \cos{\alpha}$ and $r = (\rho_m - \rho_a)/\rho_a$ with $\rho_m$ and $\rho_a$ the current density and ambient fluid density respectively. Note that $\rho_m = \phi \rho_p+(1-\phi)\rho_f$ with $\rho_p$ the particles density and $\rho_f$ the fluid phase density initially in the current. In the set of experiments presented here $\rho_f\le \rho_a$. The velocity scale $U_0$ corresponds to an idealized transfer of initial potential energy to kinetic energy along the slope, i.e. a collapse of the initial column.

Accordingly,
\begin{equation}
	\displaystyle a =\frac{H_0}{L_0},
\end{equation}
\begin{equation}
	\displaystyle Re = \frac{U_0 H_0\rho_m}{\eta_f},
\end{equation}
\begin{equation}
	\displaystyle At = \frac{\rho_m-\rho_a}{\rho_a},
\end{equation}
\begin{equation}
	\displaystyle St = \frac{V_s}{U_0},
\end{equation}
with $\eta_f$ the interstitial fluid viscosity and $V_s$ the Stokes settling velocity in the interstitial fluid phase labelled $f$.

Then, the dynamics of the current are characterized by several dimensionless variables. The most common one to be analyzed in gravity-driven avalanches is the dimensionless front velocity, also referred to as the Froude number,
\begin{equation}
	\displaystyle \mathcal{F}r =\frac{U_c}{U_0},
\end{equation}
where $U_c$ is the front velocity of the current. Note that $U_c$ is time-dependent, and so is $\mathcal{F}r$, i.e. that $\mathcal{F}r$ as defined here is not a control parameter. (The Froude number which would be defined as a control parameter for a lock-release configuration would be $Fr\equiv 1$, and is thus not relevant to predict the flow dynamics.)

\section{Results}

\subsection{Influence of the slope on the front velocity}

The influence of the slope angle $\alpha$ on the dimensionless front velocity $\mathcal{F}r$ is presented in figure \ref{fig:fig4}. Even if the entire set of experiments and numerical simulations are reported here, results are highlighted for a nearly constant $St\approx 5\times 10^{-2}$ and $\phi < 0.4$ for clarity.

\begin{figure*}
	\centering
	\includegraphics{figure4.pdf}
	\caption{Caption}
	\label{fig:fig4}
\end{figure*}

\subsection{Influence of $St$}

\begin{figure*}
	\centering
	\includegraphics{figure5.pdf}
	\caption{Caption}
	\label{fig:fig5}
\end{figure*}

Discuss figure \ref{fig:fig5}

\subsection{Influence of volume fraction, $\phi$}

\begin{figure}
	\centering
	\includegraphics{figure6.pdf}
	\caption{Caption}
	\label{fig:fig6}
\end{figure}

Discuss figure \ref{fig:fig6}

\subsection*{Acknowledgment}

\subsection*{Fundings}

\bibliographystyle{apalike}
\bibliography{biblio}

% \end{multicols*}

\newpage
%TC:endignore

\end{document}

### 2023-05-02 figure_scripts/figure3.py
Fixing mistake on fig 3

@@ -38,8 +38,11 @@ fig, axarr = plt.subplot_mosaic(layout, figsize=(tp.large_figure_width, 0.55*tp.
 ax = axarr['(a)']
 for d in datasets[mask]:
     ax.scatter(d.variables['t'][:].data, d.variables['x_front'][:].data,
-               s=2 if d.author == 'Julien' else 0.3, color=tp.color_setups[d.author],
-               zorder=zorder_setups[d.author], rasterized=True)
+               s=1,
+               color=tp.color_setups[d.author],
+               zorder=zorder_setups[d.author],
+               marker='s' if d.author == 'Julien' else None,
+               rasterized=True)
 
 ax.set_ylabel(r'Front position, $x_{\rm f}$ [m]')
 ax.set_xlabel(r'Time, $t$ [s]')
@@ -48,7 +51,7 @@ ax.set_ylim(bottom=0)
 
 # selected run, non-dimensional
 ax = axarr['(b)']
-runs = [0, 158, 94, 196, 209, 201]
+runs = [0, 158, 94, 197, 210, 202]
 
 for run in runs:
     d = datasets[list_runs.index(os.path.join(
@@ -61,7 +64,9 @@ for run in runs:
     x_axis = d.variables['t'][:].data/t_ad
     y_axis = d.variables['x_front'][:].data/x_ad
     ax.scatter(x_axis, y_axis,
-               color=tp.color_setups[d.author], s=2, rasterized=True)
+               color=tp.color_setups[d.author], s=3,
+               marker='s' if d.author == 'Julien' else None,
+               rasterized=True)
     # plotting fit
     fitresult = load_modelresult(os.path.join(
         path_data, 'fitresult_run_{:03d}.save'.format(run)))

### 2023-05-04 figure_scripts/figure4.py
Specifying phi < 0.45 for fig 5, and adjusting the same for fig 4

@@ -46,7 +46,7 @@ author_zorder = ['Rastello', 'Cyril', 'Cyril/Marie', 'Jean', 'Julien']
 mask_Stokes = (St > 4e-2) & (St < 6e-2)
 # mask_phi = (phi > 0.12) & (phi < 0.18)
 # mask_Stokes = (St < 5.5e-2)
-mask_phi = (phi < 0.40)
+mask_phi = (phi < 0.45)
 
 # %% graphic vector for plots
 alphas = np.ones_like(Fr)

### 2023-05-04 figure_scripts/figure5.py

@@ -44,7 +44,7 @@ for a0, axarr_sub in zip(alpha0, axarr):
     mask_alpha = (alpha > a0 - alpha_pad) & (alpha < a0 + alpha_pad)
     for i, (var, ax) in enumerate(zip([Fr, L], axarr_sub.flatten())):
         for author in author_zorder:
-            mask = (authors == author) & mask_alpha
+            mask = (authors == author) & mask_alpha & (phi < 0.45)
             marker = 's' if author == 'Julien' else None
             #
             ax.scatter(St[mask]/a[mask], var[mask], c=tp.color_setups[author],

### 2023-05-04 paper/main.tex
Adding captions to figures

@@
 \usepackage{multicol}
-\usepackage[margin=0.6in]{geometry}
+\usepackage[textwidth=492.982pt, top=2cm, bottom=2cm]{geometry}
 
@@
 	\includegraphics{figure1.pdf}
-	\caption{Caption 1}
+	\caption{\textbf{Experimental set-ups used in this study.} (a--c) Sketches of the three experimental set-ups.
+Below are shown snapshots of corresponding experimental images.
+(d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, $\mathcal{S}t=$ (i--k) Set-up 2, $\alpha = $, $\phi \sim $, $\mathcal{S}t=$ (l--p) Set-up 3, $\alpha = $, $\phi \sim $, $\mathcal{S}t=$.}
 	\label{fig:fig1}
@@
 	\includegraphics{figure2.pdf}
-	\caption{Caption.}
+	\caption{\textbf{Explored parameter space.} Each symbol correspond to a single experimental run.}
 	\label{fig:fig2}
@@
 	\includegraphics{figure3.pdf}
-	\caption{Caption}
+	\caption{(a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
 	\label{fig:fig3}
@@
 	\includegraphics{figure4.pdf}
-	\caption{Caption}
+	\caption{\textbf{Influence of the bottom slope.} (a) Current Froude number as a function of the bottom slope.
+(b) Current Froude number corrected by a prefector $\sqrt{a}$ as a function of the bottom slope.
+Here, plain symbols corresponds to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
 	\label{fig:fig4}
@@
 	\includegraphics{figure5.pdf}
-	\caption{Caption}
+	\caption{\textbf{Influence of particle settling.} Current Froude number $\mathcal{F}r$ and non-dimensional quadratic dissipation $\tilde{\lambda}$ as a function of the ratio between the Stokes number and the lock aspect-ratio, for different ranges of bottom slopes.
+Here, we selected runs for $\phi < 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
 	\label{fig:fig5}
@@
 \subsection{Influence of volume fraction, $\phi$}
+\label{sec:influence_phi}
 
@@
 	\includegraphics{figure6.pdf}
-	\caption{Caption}
+	\caption{\textbf{Influence of particle volume fraction.} Current Froude number as a function of the initial particle volume fraction for (a) $\alpha \approx 0^\circ$, and (b) $\alpha \approx 45^\circ$.
+The inset in (a) shows the plot in linear scale.}
 	\label{fig:fig6}

### 2023-05-04 figure_scripts/figure4.py

@@ -40,18 +40,17 @@ aspect = H0/L0
 
 # %% graphic specifications
 # changing order depending on authors
-author_zorder = ['Rastello', 'Cyril', 'Cyril/Marie', 'Jean', 'Julien']
+author_zorder = ['Cyril', 'Rastello', 'Cyril/Marie', 'Jean', 'Julien']
 
 # %% masks for plot
-mask_Stokes = (St > 4e-2) & (St < 6e-2)
-# mask_phi = (phi > 0.12) & (phi < 0.18)
-# mask_Stokes = (St < 5.5e-2)
+# mask_Stokes = (St > 4e-2) & (St < 6e-2)
+# mask_Stokes = (St < 6e-2)
 mask_phi = (phi < 0.45)
 
 # %% graphic vector for plots
 alphas = np.ones_like(Fr)
-alphas[(~mask_phi) & (authors == 'Jean')] = 0.2
-alphas[(~mask_Stokes) & (authors != 'Jean')] = 0.2
+# alphas[(~mask_phi | ~mask_Stokes)] = 0.2
+alphas[~mask_phi] = 0.2
 
 markers = np.full_like(Fr, 'o', dtype=str)
 markers[particles == 'Hydrogels'] = '*'

### 2023-05-04 paper/main.tex
Modifying fig5 to show angles

@@
 	\includegraphics{figure5.pdf}
-	\caption{\textbf{Influence of particle settling.} Current Froude number $\mathcal{F}r$ and non-dimensional quadratic dissipation $\tilde{\lambda}$ as a function of the ratio between the Stokes number and the lock aspect-ratio, for different ranges of bottom slopes.
+	\caption{\textbf{Influence of particle settling.} Current Froude number $\mathcal{F}r$ and attenuation $\tilde{\lambda}$ as a function of the ratio between the Stokes number and the lock aspect-ratio, for different ranges of bottom slopes.
 Here, we selected runs for $\phi < 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}

### 2023-05-04 figure_scripts/figure5.py

@@ -38,9 +38,9 @@ alpha0 = [1, 7, 15, 45]
 alpha_pad = 1.5
 
 figsize = (tp.large_figure_width, tp.golden*tp.large_figure_width/2)
-fig, axarr = plt.subplots(4, 2, constrained_layout=True,
-                          figsize=figsize, sharex=True)
-for a0, axarr_sub in zip(alpha0, axarr):
+fig, axarr = plt.subplots(4, 3, constrained_layout=True,
+                          figsize=figsize, gridspec_kw={'width_ratios': [0.2, 1, 1]})
+for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
     mask_alpha = (alpha > a0 - alpha_pad) & (alpha < a0 + alpha_pad)
     for i, (var, ax) in enumerate(zip([Fr, L], axarr_sub.flatten())):
         for author in author_zorder:
@@ -71,6 +71,7 @@ for a0, axarr_sub in zip(alpha0, axarr):
     axarr_sub[1].set_ylabel(r'Dissipation, $\tilde{\lambda}$')
 
     axarr_sub[1].set_xscale('log')
+    axarr_sub[0].set_xscale('log')
     # axarr_sub[1].ticklabel_format(
     #     axis='y', style='sci', scilimits=(0, 1), useMathText=True)
 
@@ -91,8 +92,19 @@ for a0, axarr_sub in zip(alpha0, axarr):
         axarr_sub[1].set_xlabel(r'Stokes number, $\mathcal{S}t/a$')
         axarr_sub[0].set_xlabel(r'Stokes number, $\mathcal{S}t/a$')
 
+xline = 0.8
+for ax, a0 in zip(axarr[:, 0].flatten(), alpha0):
+    ax.set_axis_off()
+    ax.set_xlim([0, 1])
+    ax.set_ylim([0, 1])
+    ax.axvline(x=xline, color='k')
+    ax.text(xline/2, 0.5, r'${:.0f}^\circ$'.format(
+        a0), ha='center', va='center')
 
-for ax, l in zip(axarr.flatten(), 'abcdefgh'):
+axarr[0, 0].axhline(y=0.8, xmax=xline, color='k')
+axarr[0, 0].text(xline/2, 0.83, r'$\alpha~[^\circ]$', ha='center', va='bottom')
+
+for ax, l in zip(axarr[:, 1:].flatten(), 'abcdefgh'):
[…]

### 2023-05-04 figure_scripts/figure4.py
Adding Birman data to fig 4

@@ -4,6 +4,7 @@ import sys
 
 import matplotlib.pyplot as plt
 import matplotlib.transforms as mtransforms
+from matplotlib.lines import Line2D
 import numpy as np
 import template as tp
 from netCDF4 import Dataset
@@ -18,6 +19,10 @@ def ang2sin(ang):
 
 def sin2ang(sin):
     return np.degrees(np.arcsin(sin))
+
+
+def Birman(ang):
+    return -0.1824*np.radians(ang)**2 + 0.2781*np.radians(ang) + 0.4871
 
 
 # %% Load data
@@ -63,6 +68,7 @@ layout = [['legend', 'legend'], ['(a)', '(b)']]
 fig, axarr = plt.subplot_mosaic(layout, figsize=tp.large_figure_size, constrained_layout=True,
                                 gridspec_kw={'height_ratios': [0.001, 1]})
 
+alpha_plot = np.linspace(0, 50, 500)
 for label in ['(a)', '(b)']:
     ax = axarr[label]
     for author in author_zorder:
@@ -76,8 +82,13 @@ for label in ['(a)', '(b)']:
         #               for a in alphas[mask]]
         facecolors = [to_rgba(tp.color_setups[author], a)
                       for a in alphas[mask]]
-        tp.mscatter(np.sin(np.radians(alpha[mask])), Y[mask], ax=ax,
+        tp.mscatter(ang2sin(alpha[mask]), Y[mask], ax=ax,
                     edgecolors=edgecolors, label=author, m=markers[mask], facecolors=facecolors)
+    #
+    fact = 1 if label == '(a)' else np.sqrt(0.1)
+    ax.plot(ang2sin(alpha_plot), fact*Birman(alpha_plot),
+            ls='--', color='k', zorder=-8)
+
     ax.set_xlabel(r'$\sin \alpha$')
     ax.set_xlim(left=-0.012)
 
@@ -88,15 +99,13 @@ axarr['(a)'].set_ylabel(r'Froude number, $\mathcal{F}r$')
 axarr['(b)'].set_ylabel(r'$\sqrt{a}\mathcal{F}r$')
 axarr['(a)'].set_ylim(0, 1.35)
 axarr['(b)'].set_ylim(0, 1.85)
+axarr['(a)'].set_xlim(right=ang2sin(48))
+axarr['(b)'].set_xlim(right=ang2sin(48))
 
 axarr['legend'].axis('off')
-leg = axarr['legend'].legend(handles=tp.legend_elements, loc="upper center", ncol=5,
[…]

### 2023-05-04 paper/main.tex
Advancing paper

@@
 \usepackage{textcomp}
+\usepackage{xcolor}
+\usepackage{amsmath}
 
@@
 
-\section*{Methods}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 2
-\begin{figure}
-	\centering
-	% \includegraphics{Figures/Figure_lit_plat/Figure_lit_plat.pdf}
-	\includegraphics{figure2.pdf}
-	\caption{\textbf{Explored parameter space.} Each symbol correspond to a single experimental run.}
-	\label{fig:fig2}
-\end{figure}
-
-\subsection{Experimental setups}
-
-Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed on the front dynamics, often referred to as the $\mathcal{F}r$ number, and the front shape.
-Then prediction of the different flow regimes associated with PLGCs shall be entirely included into these observables.
-In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at “rest”, 3 dam-break-type devices have been used to cover a wide range of control parameters, as $\alpha \in [0^o,30^o]$, $\phi\in[0.1\%,20\%]$, $At\in[???]$, $Re\in[???]$, $St\in[???]$, $a\in??$, as specifically defined in section \ref{sec:dimensionlessmap}.
-These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.
-
-\subsection{Datasets}
-
-On these 3 devices, a total of $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below).
-They are classified into 4 different datasets, described below.
-
-\paragraph{Dataset 1}
-
-Description Dataset LEGI
-
-\paragraph{Dataset 2}
-
-Description Dataset IMFT
-
-\paragraph{Dataset 3}
-
-Description Dataset LEGI/IMFT
-
-\paragraph{Dataset 4}
-
-Description Dataset LEMTA
-
-\paragraph{SedFoam}
-
-Description Dataset SedFoam
-
-\subsection{Fitting procedure and front position analysis}
-
-For each experiment/simulation, we extract the current front position as a function of time (see Fig. \ref{fig:fig3}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 3
-\begin{figure*}
-	\centering
-	\includegraphics{figure3.pdf}
-	\caption{(a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
-	\label{fig:fig3}
-\end{figure*}
-
 \section{Dimensional analysis at the macroscopic scale of the current}
@@
 (The Froude number which would be defined as a control parameter for a lock-release configuration would be $Fr\equiv 1$, and is thus not relevant to predict the flow dynamics.)
+
+\section{Methods}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 2
+\begin{figure}
+	\centering
+	% \includegraphics{Figures/Figure_lit_plat/Figure_lit_plat.pdf}
+	\includegraphics{figure2.pdf}
+	\caption{\textbf{Explored parameter space.} Each symbol correspond to a single experimental run.}
+	\label{fig:fig2}
+\end{figure}
+
+\subsection{Experimental setups}
+
+Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed on the front dynamics, often referred to as the $\mathcal{F}r$ number, and the front shape.
+Then prediction of the different flow regimes associated with PLGCs shall be entirely included into these observables.
+In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at “rest”, 3 dam-break-type devices have been used to cover a wide range of control parameters, as $\alpha \in [0^o,30^o]$, $\phi\in[0.1\%,20\%]$, $At\in[???]$, $Re\in[???]$, $St\in[???]$, $a\in??$, as specifically defined in section \ref{sec:dimensionlessmap}.
+These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.
+
+\paragraph{Set-up 1}
+
+Specifications of IMFT set-up
+
+\paragraph{Set-up 2}
+
+Specifications of LEGI set-up
+
+\paragraph{Set-up 3}
+
+Specifications of LEMTA set-up
+
+\subsection{Datasets}
+
+On these 3 devices, a total of $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below).
+They are classified into 4 different datasets, described below.
+
+\paragraph{Dataset 1}
+
+Description Dataset LEGI
+
+\paragraph{Dataset 2}
+
+Description Dataset IMFT
+
+\paragraph{Dataset 3}
+
+Description Dataset LEGI/IMFT
+
+\paragraph{Dataset 4}
+
+Description Dataset LEMTA
+
+\paragraph{SedFoam}
+
+Description Dataset SedFoam
+
+\subsection{Fitting procedure and front position analysis}
+
+For each experimental/numerical run, we extract the current front position as a function of time (see Fig.
+\ref{fig:fig3}a).
+A few selected runs are shown on figure~\ref{fig:fig3}b in non-dimensional time coordinates.
+After a short transient acceleration phase, most currents exhibit at first a constant velocity corresponding to the so-called \emph{slumping} regime.
+After this, the velocity can decrease, either because of settling induced buoyancy loss, or because the current enters the \emph{inertial} regime.
+
+It can be qualitatively seen that increasing the bottom slope increases the slumping velocity, while increasing the Stokes number shortens the duration of the constant-velocity regime.
+%
+To properly quantify that, we fit all front propagation curves (except Dataset 4) with the following equation:
+
+\begin{equation}
+	\label{eq:fit_eq}
+	\frac{x_{\rm f}}{L_{0}} = x_{i} + \mathcal{F}r \frac{t}{t_{0}} - \frac{\tilde{\lambda}}{2} \left(\frac{t}{t_{0}}\right)^{2},
+\end{equation}
+where $\mathcal{F}r$ corresponds now to the non-dimensional current velocity during the slumping regime, and $\tilde{\lambda}$ to the dissipation.
+The fit is performed between $t/t_{0} = 5$, to avoid the initial transient phase, and $t/t_{0} = 30$, the transition time from the slumping to the inertial regime~\citep{refs}.
+Note that $x_{i}$ is kept constant equal to 0 for Dataset 1, 2 and 3, but had to be left free to adjust for the SedFoam simulations \textcolor{red}{comment this}.
+For runs with strong dissipation ($\tilde{\lambda} > 2 {\times} 10^{-2}$), we add a quartic term to \eqref{eq:fit_eq} to improve the fit quality and obtain a better estimation of $\mathcal{F}r$ and $\tilde{\lambda}$.
+We checked that our results does not depend on the presence of this quartic term, and that the linear and quadratic ones remain much larger.
+
+For Dataset 4, the time resolution as well as the experiment duration does not allow to extract the dissipation. Therefore, we simply fit the whole available points with a linear function.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 3
+\begin{figure*}
+	\centering
+	\includegraphics{figure3.pdf}
+	\caption{(a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
+	\label{fig:fig3}
+\end{figure*}
 

### 2023-05-21 _tests/Cyril/check_fits_analysis.py
Marking all expriments no suspension with red

@@ -82,9 +82,9 @@ SETUPS = {'Cyril': 'IMFT', 'Cyril/Marie': 'LEGI', 'Jean': 'LEMTA',
           'Julien': 'NUM', 'Rastello': 'LEGI'}
 
 # %% Loading data
-list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_JULIEN2/*.nc')))
+# list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_JULIEN2/*.nc')))
 # list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_JEAN/*.nc')))
-# list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_MARIE/*.nc')))
+list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_MARIE/*.nc')))
 # list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_CYRIL/*.nc')))
 datasets = np.array([Dataset(run) for run in list_runs])
 
@@ -94,6 +94,8 @@ grains = np.array([d.particle_type for d in datasets])
 authors = np.array([d.author for d in datasets])
 diams = np.array([d.variables['d'][:].data for d in datasets])
 phi = np.array([d.variables['phi'][:].data for d in datasets])
+H0 = np.array([d.variables['H0'][:].data for d in datasets])
+Ha = np.array([d.variables['H_a'][:].data for d in datasets])
 
 
 authors = np.array([d.author for d in datasets])
@@ -105,10 +107,10 @@ authors = np.array([d.author for d in datasets])
 #     grains == 'silica sand') & (authors == 'Cyril')
 # mask_runs = (authors == 'Cyril/Marie')
 # mask_runs = np.ones_like(alphas).astype('bool')
-mask_runs = (alphas > 6) & (alphas < 8)
+# mask_runs = (alphas > 6) & (alphas < 8)
 # mask_runs = (alphas > 40)
 # mask_runs = (grains == 'PMMA')
-
+mask_runs = (H0/Ha < 0.3)
 # %% Loop over data file and analysis
 
 fig_fit, ax = plt.subplots(1, 1, constrained_layout=True)

### 2023-05-21 figure_scripts/figure2.py

@@ -38,7 +38,9 @@ markers[dataset_idx == 'SedFoam'] = 's'
 
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 edgecolors = np.full_like(facecolors, 'k')
-edgecolors[H0/Ha < 0.2] = 'tab:red'
+
+mask_nosuspended = (authors == 'Rastello') & (H0/Ha < 1)
+edgecolors[mask_nosuspended] = 'tab:red'
 
 zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)

### 2023-05-21 figure_scripts/figure4.py

@@ -43,11 +43,11 @@ authors, particles = np.array([[d.author, d.particle_type,
 
 # %% graphic specifications
 # %% masks for plot
-mask_phi = (phi < 0.45)
+mask_phi = (phi < 0.449)
 
 # %% graphic vector for plots
 alphas = np.ones_like(Fr)
-alphas[~mask_phi] = 0.2
+alphas[~mask_phi] = 0.4
 
 dataset_idx = np.vectorize(tp.datasets.get)(authors)
 
@@ -57,10 +57,10 @@ markers[dataset_idx == 'SedFoam'] = 's'
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 facecolors = np.array([to_rgba(c, a) for c, a in zip(facecolors, alphas)])
 
-# edgecolors = np.copy(facecolors)
+mask_nosuspended = (authors == 'Rastello') & (H0/Ha < 1)
 edgecolors = np.array([to_rgba('k', a) for a in alphas])
-edgecolors[H0/Ha < 0.2] = np.array([to_rgba('tab:red', 1)
-                                   for a in alphas[H0/Ha < 0.2]])
+edgecolors[mask_nosuspended] = np.array(
+    [to_rgba('tab:red', 0.4) for a in alphas[mask_nosuspended]])
 
 zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
@@ -88,7 +88,7 @@ secax = ax.secondary_xaxis('top', functions=(sin2ang, ang2sin))
 secax.set_xlabel(r'angle, $\alpha$ [deg.]')
 
 ax.set_ylabel(r'Froude number, $\mathcal{F}r$')
-ax.set_ylim(0, 1.6)
+ax.set_ylim(0, 1.59)
 ax.set_xlim(right=ang2sin(48))
 
 other_elements = [Line2D([0], [0], color='k', ls='--',

### 2023-05-21 figure_scripts/figure5.py

@@ -37,7 +37,9 @@ markers[dataset_idx == 'SedFoam'] = 's'
 
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 edgecolors = np.full_like(facecolors, 'k')
-edgecolors[H0/Ha < 0.2] = 'tab:red'
+
+mask_nosuspended = (authors == 'Rastello') & (H0/Ha < 1)
+edgecolors[mask_nosuspended] = 'tab:red'
 
 zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
@@ -54,7 +56,7 @@ fig, axarr = plt.subplots(4, 3, constrained_layout=True,
                           figsize=figsize, gridspec_kw={'width_ratios': [0.2, 1, 1]})
 for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
     mask_alpha = (alpha > a0 - alpha_pad) & (alpha < a0 + alpha_pad)
-    mask = (mask_alpha & (phi < 0.45))[plot_idxs]
+    mask = (mask_alpha & (phi < 0.449))[plot_idxs]
     for i, (var, ax) in enumerate(zip([Fr, L], axarr_sub.flatten())):
         tp.mscatter((St/a)[plot_idxs][mask], var[plot_idxs][mask], ax=ax, m=markers[plot_idxs][mask],
                     facecolors=facecolors[plot_idxs][mask], edgecolors=edgecolors[plot_idxs][mask], lw=0.5)
@@ -62,7 +64,7 @@ for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
         if i == 0:
             moy, std = np.nanmean(
                 var[plot_idxs][mask]), np.nanstd(var[plot_idxs][mask])
-            ax.axhline(moy, color='k', ls=':', zorder=-10)
+            ax.axhline(moy, color='k', ls=':', zorder=-10, lw=1)
             # ax.axhspan(moy - std, moy + std, color='k', alpha=0.2, zorder=-10)
         else:
             ax.axhline(0, color='k', ls=':', zorder=-10, lw=1)
@@ -72,7 +74,7 @@ for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
                            & mask_alpha)
             moy, std = np.nanmean(
[…]

### 2023-05-21 figure_scripts/figure6.py

@@ -36,7 +36,9 @@ markers[dataset_idx == 'SedFoam'] = 's'
 
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 edgecolors = np.full_like(facecolors, 'k')
-edgecolors[H0/Ha < 0.2] = 'tab:red'
+
+mask_nosuspended = (authors == 'Rastello') & (H0/Ha < 1)
+edgecolors[mask_nosuspended] = 'tab:red'
 
 zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
@@ -67,7 +69,7 @@ for alpha0, ax in zip(alphas, axarr.flatten()):
                 facecolors=facecolors[plot_idxs][mask], edgecolors=edgecolors[plot_idxs][mask], lw=0.5)
     #
     ax.set_ylabel(r'Froude number, $\mathcal{F}r$')
-    ax.set_ylim([0, 1.6])
+    ax.set_ylim([0, 1.59])
 
 ax.set_xlabel(r'Volume fraction, $\phi$')
 ax.set_xscale('log')

### 2023-05-22 figure_scripts/figure4.py
update fig5 with phi > 0.45 points in transparent

@@ -3,12 +3,11 @@ import os
 import sys
 
 import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
 import numpy as np
 import template as tp
-from netCDF4 import Dataset
 from matplotlib.colors import to_rgba
-
+from matplotlib.lines import Line2D
+from netCDF4 import Dataset
 
 plt.rcParams['xtick.top'] = False
 

### 2023-05-22 figure_scripts/figure5.py

@@ -6,6 +6,7 @@ import matplotlib.pyplot as plt
 import matplotlib.transforms as mtransforms
 import numpy as np
 import template as tp
+from matplotlib.colors import to_rgba
 from netCDF4 import Dataset
 
 plt.rcParams['figure.constrained_layout.hspace'] = 0
@@ -32,14 +33,20 @@ a = H0/L0
 # %% graphic specifications
 dataset_idx = np.vectorize(tp.datasets.get)(authors)
 
+mask_phi = (phi < 0.449)
+alphas = np.ones_like(Fr)
+alphas[~mask_phi] = 0.4
+
 markers = np.vectorize(tp.marker_style.get)(particles)
 markers[dataset_idx == 'SedFoam'] = 's'
 
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
-edgecolors = np.full_like(facecolors, 'k')
+facecolors = np.array([to_rgba(c, a) for c, a in zip(facecolors, alphas)])
 
 mask_nosuspended = (authors == 'Rastello') & (H0/Ha < 1)
-edgecolors[mask_nosuspended] = 'tab:red'
+edgecolors = np.array([to_rgba('k', a) for a in alphas])
+edgecolors[mask_nosuspended] = np.array(
+    [to_rgba('tab:red', 0.4) for a in alphas[mask_nosuspended]])
 
 zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
@@ -56,14 +63,14 @@ fig, axarr = plt.subplots(4, 3, constrained_layout=True,
                           figsize=figsize, gridspec_kw={'width_ratios': [0.2, 1, 1]})
 for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
     mask_alpha = (alpha > a0 - alpha_pad) & (alpha < a0 + alpha_pad)
-    mask = (mask_alpha & (phi < 0.449))[plot_idxs]
+    mask = (mask_alpha)[plot_idxs]
     for i, (var, ax) in enumerate(zip([Fr, L], axarr_sub.flatten())):
         tp.mscatter((St/a)[plot_idxs][mask], var[plot_idxs][mask], ax=ax, m=markers[plot_idxs][mask],
[…]

### 2023-05-22 figure_scripts/figure5.py
Adding trend to fig 5

@@ -6,11 +6,17 @@ import matplotlib.pyplot as plt
 import matplotlib.transforms as mtransforms
 import numpy as np
 import template as tp
+from lmfit import Model
 from matplotlib.colors import to_rgba
 from netCDF4 import Dataset
 
 plt.rcParams['figure.constrained_layout.hspace'] = 0
 plt.rcParams['figure.constrained_layout.h_pad'] = 0.0005
+
+
+def lambda_var(x, a, c, th):
+    return np.piecewise(x, [x < th, x >= th], [lambda x: c, lambda x: a*(x - th) + c])
+
 
 # %% Load data
 path_data = '../data/output_data'
@@ -54,9 +60,26 @@ rng = np.random.default_rng(1994)
 rng.shuffle(random_order)
 plot_idxs = np.lexsort((random_order, zorders))
 
-# %% masks for plot
 alpha0 = [0, 7, 15, 45]
 alpha_pad = 1.5
+
+# %% fit for alpha = 7
+
+model = Model(lambda_var)
+params = model.make_params()
+p0 = {'a': 1, 'c': 0, 'th': 1e-2}
+
+for par in params.keys():
+    params[par].set(value=p0[par])
+
+params['th'].vary = False
+params['c'].vary = False
+
+mask_alpha = (alpha > alpha0[1] - alpha_pad) & (alpha < alpha0[1] + alpha_pad)
+
+result = model.fit(L[mask_alpha], params, x=(St/a)[mask_alpha])
+
+# # %% Figure
 
 figsize = (tp.large_figure_width, tp.golden*tp.large_figure_width/2)
 fig, axarr = plt.subplots(4, 3, constrained_layout=True,
@@ -75,6 +98,9 @@ for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
             # ax.axhspan(moy - std, moy + std, color='k', alpha=0.2, zorder=-10)
         else:
             ax.axhline(0, color='k', ls=':', zorder=-10, lw=1)
+            x_plot = np.logspace(np.log10(1.52e-2), 0, 300)
+            ax.plot(x_plot, result.params['a']*(x_plot - result.params['th']
+                                                ) + result.params['c'], color='k', lw=1, ls='-')
         #
         if a0 == 7:
             mask_saline = ((particles == 'saline water')

### 2023-05-22 paper/main.tex
Update .tex from Overleaf

@@
 \usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
 \usepackage{natbib}
@@
 \usepackage{textcomp}
-\usepackage{xcolor}
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage{amsmath}
-
-% \usepackage{moreverb,savetrees}
 
@@
 
-\graphicspath{{./figures/}{./}}
+\newcommand*{\cyril}{\textcolor{ForestGreen}}
 
-\title{Title}
-\author[1]{PALAGRAM}
+\graphicspath{{./figures}{./}}
 
-\affil[1]{Affiliations}
-
-\date{}
-
-\newenvironment{Figure}
-  {\par\medskip\noindent\minipage{\linewidth}}
-  {\endminipage\par\medskip}
+\title{Macroscopic description of particle-laden gravity currents: overall perspectives}
+\author{Gadal C., Schneider J., Bonamy C., Chauchat J., Dossmann Y., Kiesgen de Richter S., Mercier M., Naaim F., Rastello M., Lacaze L.}
+\date{{\it\color{red}Deadline: May 31$^{st}$}}
 
@@
 \twocolumn[
+
 	\maketitle
@@
 	\begin{abstract}
-		Particle-laden gravity currents (PLGCs) are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid.
+		Particle-laden gravity currents (PLGCs) are driven by a mass difference between an heavy fluid-particle mixture and a lighter ambient liquid.
 		We focus our attention on finite volume current, generated here by lock-release device over a controlled constant slope bottom at the laboratory scale.
@@
 average over the current depth.
+
 	\end{abstract}
-	\vskip 0.7cm
+
+	\vspace{0.3cm}
+	{\it \color{red}REMINDER: Word limit of about 10,000 per article, plus 7 figures.}\\
+	\vspace{0.3cm}
 ]
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 1
-\begin{figure*}[h]
+\section{Introduction}
+\label{sec:intro}
+
+\cyril{
+	\begin{itemize}
+		\item définir $U_{0} = \sqrt{g' H_{0}}$ dans l'intro déjà ?
+		\item inclure le $\cos\alpha$ dans cete definition directement ? $U_{0} = \sqrt{g' H_{0} \cos\alpha}$
+	\end{itemize}
+}
+
+\subsection{Finite volume density currents}
+
+Density currents have been widely considered at the laboratory scale for their ubiquitous occurrence in geophysical flows \citep{Hopfinger1983,Simpson1999,Dufek2016}.
+As they are mostly associated with unsteady and non-uniform dynamics in nature, the lock-release device appeared to be a popular canonical configuration to mimic their behaviour at the laboratory scale \citep{Hacker1996,Shin2004,Nogueira2014}.
+In particular, studies often focused on characterising the front dynamics and the front shape as a macroscopic description of the underlying complex dynamics \citep{Rottman1983,Marino2005,Hogg2006}.
+
+The basic approach to describe a lock-release evolution is built upon characteristics solutions of the hyperbolic shallow water equations for an inviscid dense liquid of density $\rho_c$ slumping on a horizontal bottom under an inertialess and weightless ambient fluid.
+Among others, this specifically allows quantifying a most obvious scaling analysis solution reading for the front velocity $U_c$ reading
+\begin{equation}
+	U_c\propto \sqrt{gH_0},
+\end{equation}
+with $g$ the gravity and $H_0$ the initial height of the fluid volume being released.
+The method of characteristics suggests a coefficient of proportionality of exactly $2$ between velocities $U_c$ and $\sqrt{gH_0}$.
+An extension towards a weighted ambient fluid of density $\rho_a$ shall lead to
+\begin{equation}
+	U_c\propto \sqrt{g'H_0},\quad\mbox{with}\quad g'=\frac{\rho_c-\rho_a}{\rho_c}g.
+	\label{eq:solUc}
+\end{equation}
+However, the coefficient of proportionality now requires slightly more attention as it is clearly found smaller than $2$ in experiments or simulations \citep{??}.
+The main reason is that most density currents are characterized by a small density ratio between current and ambient, making the assumption of inertialess ambient fluid unlikely.
+Knowledge of the front velocity then requires a local description of the flow close to the front, which in turn adds the characteristic height of the ambient as a required control parameter, as suggested by \citet{Benjamin1968}.
+In the case of lock-release configuration, it would suggest an influence of the ratio between reservoir height $H_0$ and ambient height $H$.
+Nevertheless, obtained results allow linking the front velocity to the height of the current close to the front, which is a priori unknown in a lock-release device, i.e.
+the front relation includes two unknowns \citep{Benjamin1968}.
+Models have been proposed to circumvent this issue, leading for instance to a coefficient of proportionality $0.5 (\rho_c/\rho_a)^{1/2}(H/H_0)^{1/3}$ between the two velocities in \eqref{eq:solUc} \citep{Huppert1980}.
+The latter coefficient is relevant for not two deep ambient $H$ compared to the reservoir height $H_0$, which is actually the case in most experiments.
+Most configurations correspond to $H=H_0$, which would lead to a value of $0.5$.
+Even if several experiments and numerical simulations confirm such a value around $0.5(\rho_c/\rho_a)^{1/2}$, its exact value remains still unclear, and probably dependent on several physical processes.
+In particular, these velocities relationships during the slumping regime are based on negligible dissipation during this regime, which shall be violated in some cases, as significant mixing with the ambient at the miscible interface,  settling particles, etc.
+
+The slumping regime ends when either the rarefaction wave associated with finite volume or dissipation induced by bottom friction and entrainment at the miscible interface, overcomes the current dynamics \citep{???}.
+From dimensional analysis, a transition length scale $X_{\rm t}$ towards these regimes would lead, respectively, to
+\begin{equation}
+	\label{eq:Xt_inertial}
+	\displaystyle    X_{\rm t}=  \frac{2L_0 }{\sqrt{g'H_0}/U_c - 1},
+\end{equation}
+\begin{equation}
+	\label{eq:Xt_drag}
+	\displaystyle    X_{\rm t}=H_0  \frac{g'H_0}{C_dU_c^2},
+\end{equation}
+with $C_d$ some drag coefficient induced by friction at the bottom and by entrainment at the interface.
+This would give with the previous estimation of the velocity ratio $0.5(\rho_c/\rho_a)^{1/2}$ during the slumping regime, respectively,
+$X_{\rm t}=2L_0/[2(\rho_a/\rho_c)^{1/2}-1]$ \textcolor{blue}{A voir? } and $X_{\rm t}=[4(\rho_a/\rho_c)H_0]/C_d$.
+%\begin{equation}
+%\displaystyle    X_{\rm t}=2L_0/[2(\rho_a/\rho_c)^{1/2}-1],
+%\end{equation}
+Details of such transitions actually depend on the complex details of the flow and drag induced by the current intruding the ambient and the friction at the bottom.
+These transitions then become case-dependent and required parametrization.
+For instance, following experiments of \cite{Rottman1983}, it was found that???
+%Moreover, another transition has also been reported in the literature, referred to as the buoyancy-inertia regime \citep{Hoult1972}.
+Yet, this regime is hardly visible, probably due to the succession of different temporal law evolution of the front from the constant slumping regime towards the decelerating viscous regime, and can be disregarded for the purpose of the present study \citep{Huppert1980}.
+
+%The slumping regime ends when either the rarefaction wave associated with finite volume or dissipation induced by bottom friction and entrainment at the miscible interface, overcome the current dynamics \citep{???}.
+%By dimensional analysis, they lead to transition length scale $X_{\rm t}$, respectively
+%\begin{equation}
+%\displaystyle    X_{\rm t}= 2L_0  \frac{1}{\sqrt{g'H_0}/U_c - 1},
+%\end{equation}
+%\begin{equation}
+%\displaystyle    X_{\rm t}=H_0  \frac{g'H_0}{C_dU_c^2},
+%\end{equation}
+%with $C_d$ some drag coefficient induced by friction at the bottom and by entrainment at the interface.
+This would give with the previous estimation of the velocity ratio $0.5$ during the slumping regime, respectively
+%\begin{equation}
+%\displaystyle    X_{\rm t}=2L_0,
+%\end{equation}
+%\begin{equation}
+%\displaystyle    X_{\rm t}=\frac{4H_0}{C_d}.
+%\end{equation}
+
+Constant bottom slope $\alpha$, often relevant to characterise the main process of natural flow over topography, adds complexity to this front analysis.
+In particular, including a bottom slope into the previous slumping regime is not obvious as it would suggest a constant acceleration of the flow, at least until balanced by friction at both the interface and the bottom.
+Yet, constant regime flows are observed at early stages in experiments and simulations prior to such balance \citep{Birman2007,???}.
+It has been shown that the reason for that is the relatively long time or length scales over which acceleration imposed by the slope overcome flow induced by the lock-release pressure gradient.
+A way to highlight that is a simple dimensional balance between slope and pressure gradient.
+This can be written as
+%\begin{equation}
+%\displaystyle    (\rho_c - \rho_a)g \sin{(\alpha)} %= \frac{(\rho_c - \rho_a)g \cos{(\alpha)}H_0}{X_{\rm t}},
+%\end{equation}
+\begin{equation}
+	\label{eq:Xt_slope}
+	X_{\rm t}=H_0\tan^{-1}{(\alpha)}.
+\end{equation}
+where $X_{\rm t}$ corresponds to the length scale over which slope effect and pressure gradient balance, and can be seen as a transition length scale over which slope drives the current until reaching the next equilibrium between slope and friction.
+Nevertheless, the slope can affect the specific relationship \eqref{eq:solUc} linking the velocity scales during the slumping regime \citep{Birman2007,???}.
+Such a relationship still needs to be clarified.
+
+%%%%%%Figure 1
+\begin{figure*}[ht!]
 	\centering
@@
 Below are shown snapshots of corresponding experimental images.
-(d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, $\mathcal{S}t=$ (i--k) Set-up 2, $\alpha = $, $\phi \sim $, $\mathcal{S}t=$ (l--p) Set-up 3, $\alpha = $, $\phi \sim $, $\mathcal{S}t=$.}
+(d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, $\mathcal{S}t=$ (i--k) Set-up 2, $\alpha = $, $\phi \sim $, $\mathcal{S}t=$ (l--p) Set-up 3, $\alpha = $, $\phi \sim 55\%$, $\mathcal{S}t=$.}
 	\label{fig:fig1}
 \end{figure*}
+%%%%%%%%
 
-\section{Introduction}
+\subsection{Particle-Laden Gravity Currents}
 
-\subsection{Control parameters}
+Particle-Laden Gravity Currents (PLGCs) refer to as gravity currents, or density currents, are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid \citep{Hopfinger1983,Middleton1993,Kneller2000,Meiburg2010,wells2021}.
+Different subclasses of PLGCS can be distinguished depending on the relative density of the different phases constituting the system; i.e.
+the particle phase, the carrier fluid phase and the ambient fluid phase.
+In the following, we will focus on two of them, characterized by only two distinct densities.
+At one end, the carrier phase and ambient phase are the same fluid and particles are heavier.
+This is the most commonly studied subclass of PLGCs, usually referred to as turbidity currents.
+At the other end, PLGCs can also be induced by a neutrally-buoyant fluid-particle mixture flowing into a lighter ambient fluid.
+A major difference between PLGCs and compositional currents is therefore the existence of typical length and time scales associated with the particle dynamics.
+%At one end, the own settling velocity of heavy particle only allow the existence of these gravity currents on finite time scale and finite length scale.
+Nevertheless, the prediction of these finite scales is difficult as it shall strongly depends on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface, which modifies entrainment and associated dissipation.
+In turn, this can modify the macroscopic behaviour of the current, as for instance the dynamics of the front.
+Accordingly, their dynamics shall be controlled by several dimensionless parameters leading to a wide variety of flow regimes.
+In particular, beyond the slope $\alpha$, the dimensionless density difference between the current and the ambient $At$ and the ratio between available inertia and dissipation, say a Reynolds number $Re$, as usually introduced as a control parameter for any density currents, the initial solid fraction of particles $\phi$ and a dimensionless settling velocity $St$ have now to be considered as potentially affecting flow regimes.
+It shall be noted that in some specific situations, some of these dimensionless parameters could be linked, for instance, $\phi$ and $At$ for heavy particles in a single-phase liquid for the carrying and ambient fluids.
 
-Particle-laden gravity currents (PLGCs) are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid.
-A major difference between compositional currents and PLGCs is therefore the existence of typical length and time scales associated with the particle dynamics.
-Particularly, the own settling velocity of a heavy particle only allows the existence of these gravity currents on finite time and length scales.
-Nevertheless, prediction of these finite scales is difficult as it shall strongly depends on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface, which modifies entrainment and associated dissipation, and in turn can delay or enhance settling.
-Accordingly, their dynamics are controlled by several dimensionless parameters leading to a wide variety of flow regimes.
-In particular, beyond the slope $\alpha$, the dimensionless density difference between the current and the ambient $At$ and the ratio between available inertia and dissipation, say a Reynolds number $Re$, as usually introduced as a control parameter for compositional currents, the initial solid fraction of particles $\phi$ and a dimensionless settling velocity St have now to be considered as potentially affecting flow regimes.
-It shall be noted that in some specific situations, some of these dimensionless parameters could be linked as, for instance, $\phi$ and $At$ for heavy particles in a single-phase liquid for the carrying and ambient fluids
-
-\subsection{Literature}
-
-\begin{itemize}
-	\item literature slope
-	\item literature settling
-	\item literature volume fraction
-	\item other effects: lock geometry, non-boussinesq, entrainment, erosion/deposition $\rightarrow$ ignored here
-\end{itemize}
+Following the balance scaling as done far, a transition between the equivalent slumping regime as described for gravity current and deposition of particle.
+Instead of dynamical scaling as done previously, here a kinematic scaling is obtained as the balancing time scale of current propagation and time scale of particle settling.
+This leads to
+\begin{equation}
+	\label{eq:Xt_settling}
+	X_{\rm t}=H_0\frac{U_c}{V_s},
+\end{equation}
+in which $V_S$ is the settling velocity of the solid particles within the PLGCs. Of course, such transition disappears for the neutrally-buoyant situation described previously.
+Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear.
+Overall, the specific ratio between velocities in \eqref{eq:solUc} in the slumping inertial regime remains to clarify varying the influence of the mechanisms discussed so far.
 
@@
 
-Here, the dimensionless parameters are based on a characteristic length $L_0$ and a characteristic velocity $U_0=\sqrt{g' H_0}$ where $L_0$ and $H_0$ are the horizontal length and depth of the initial container, respectively, $g'$ is the modified gravity $r g \cos{\alpha}$ and $r = (\rho_m - \rho_a)/\rho_a$ with $\rho_m$ and $\rho_a$ the current density and ambient fluid density respectively.
-Note that $\rho_m = \phi \rho_p+(1-\phi)\rho_f$ with $\rho_p$ the particles density and $\rho_f$ the fluid phase density initially in the current.
+%%%%%%Figure 2  
+\begin{figure}
+	\centering
+	\includegraphics{figure2.pdf}
+	\caption{\textbf{Explored parameter space.} Each symbol corresponds to a single experimental run.}
+	\label{fig:fig2}
+\end{figure}
+%%%%%%%%%%
+
+Here, the dimensionless parameters are based on a characteristic length $L_0$ and a characteristic velocity $U_0=\sqrt{g' H_0}$ where $L_0$ and $H_0$ are the horizontal length and depth of the initial container, respectively, $g'$ is the modified gravity $ g \cos{\alpha}(\rho_m-\rho_a)/\rho_m$ with $\rho_m$ the current mixture density (analogous to $\rho_c$ as for current in \eqref{eq:solUc}) and $\rho_a$ the current density and ambient fluid density respectively.
+Note that $g'$ is slightly different from the definition in section \ref{sec:intro}.
+Here $g'$ is written with the ambient density at the denominator, i.e.
+as to include the influence of density ratio on front velocity as discussed in section \ref{sec:intro}.
+The mixture density can be quantified as $\rho_m = \phi \rho_p+(1-\phi)\rho_f$ with $\rho_p$ the particles density and $\rho_f$ the fluid phase density initially in the current.
 In the set of experiments presented here $\rho_f\le \rho_a$.
@@
 a collapse of the initial column.
-
 Accordingly,
@@
 \end{equation}
+%\textcolor{blue}{Rem: avant on avait $(\rho_m-\rho_a)/\rho_a$. qu est ce que cela change.}
 \begin{equation}
@@
 with $\eta_f$ the interstitial fluid viscosity and $V_s$ the Stokes settling velocity in the interstitial fluid phase labelled $f$.
+Note that $At$ usually intervenes only in the modified gravity $g'$ for Boussinesq currents, which is mostly the case for density currents \citep{Bonometti2011,???}
 
-Then, the dynamics of the current are characterized by several dimensionless variables.
-The most common one to be analyzed in gravity-driven avalanches is the dimensionless front velocity, also referred to as the Froude number,
+Then, the dynamics of the current is characterised by several dimensionless variables.
+The most common one to be analysed in gravity-driven avalanches is the dimensionless front velocity, also referred to as the Froude number,
 \begin{equation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 2
-\begin{figure}
-	\centering
-	% \includegraphics{Figures/Figure_lit_plat/Figure_lit_plat.pdf}
-	\includegraphics{figure2.pdf}
-	\caption{\textbf{Explored parameter space.} Each symbol correspond to a single experimental run.}
-	\label{fig:fig2}
-\end{figure}
+\subsection{Experimental setups}
+Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed in the front dynamics, often referred to as the $\mathcal{F}r$ number, and the front shape.
+Then prediction of the different flow regimes associated with PLGCs shall be entirely included in these observables.
+In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at ``rest'', 3 dam-break-type devices have been used to cover a wide range of control parameters, These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.
 
-\subsection{Experimental setups}
-
-Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed on the front dynamics, often referred to as the $\mathcal{F}r$ number, and the front shape.
-Then prediction of the different flow regimes associated with PLGCs shall be entirely included into these observables.
-In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at “rest”, 3 dam-break-type devices have been used to cover a wide range of control parameters, as $\alpha \in [0^o,30^o]$, $\phi\in[0.1\%,20\%]$, $At\in[???]$, $Re\in[???]$, $St\in[???]$, $a\in??$, as specifically defined in section \ref{sec:dimensionlessmap}.
-These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.
+All setups are based on the same typical lock-release design (see figure~\ref{fig:fig1}).
+In particular, a tank of length $L_0+L_1$ along the streamwise direction $x$, width $W_0$ and height $H_0$, is separated by a sluice gate at a distance $L_0$ from one end along $x$, allowing to distinguish an initial reservoir in which the particle-laden mixture is prepared from the ambient fluid.
+The sluice gate is lifted up manually at $t=0$ to initiate the gravity current.
 
@@
 
-Specifications of IMFT set-up
+The first set-up (figure~\ref{fig:fig1}(a)) has $L_0+L_1=150~\textup{cm}$ with $L_0 = 10~\textrm{cm}$.
+The upper boundary is a free surface and the tank can be tilted at an angle $\alpha=[0^\circ,7.5^\circ]$ with respect to the horizontal as shown in the 2D sketch of figure~\ref{fig:fig1}(a).
+The suspension is made by strongly stirring a known mass of particles within the whole water column using a chemical mixer, which is stopped just before opening the gate.
+In this inclined tank, $H_{0}$ is defined as the average suspension height inside the lock, equal to $20~\textrm{cm}$.
+%
+The current evolution is followed by a camera at 120 Hz and using a backlight as a light source.
 
@@
 
-Specifications of LEMTA set-up
+The third set-up (figure~\ref{fig:fig1}(c)) has $L_0+L_1=150~\textup{cm}$ with $L_0 = 20~\textrm{cm}$.
+The upper boundary is a free surface and the tank remains horizontal ($\alpha = 0^\circ$).
+Here, $H_{0}$ is between 0.2~cm and 0.3~cm depending on experiments.
+The current evolution is followed by a camera at 10 Hz and using a backlight as a light source.
+
+% The suspension is made by strongly stirring a known mass of particles within the whole water column using a chemical mixer, which is stopped just before opening the gate.
 
@@
 
-On these 3 devices, a total of $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below).
-They are classified into 4 different datasets, described below.
+On these 3 devices, $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below). They are classified into 4 different datasets, described below.
 
@@
 
-Description Dataset LEGI
+The first dataset is gathered using set-up 1.
+A first series of experiments is made with silica sand grains of $d \sim 120~\mu\textrm{m}$ while varying the bottom slope from $0^\circ$ to $7^\circ$.
+Then, for $\alpha = 7^\circ$, the settling velocity is varied by using glass beads (©Silibeads) of mean diameter ranging from $60~\mu\textrm{m}$ to $250~\mu\textrm{m}$, corresponding to $V_{\rm s} \in [0.3, 3.2]~\textrm{cm}~\textrm{s}^{-1}$.
+A few experiments with saline water (homogeneous gravity currents) performed at this $\alpha$-value are also included in the dataset.
 
@@
 
-Description Dataset IMFT
+Description Dataset LEGI
 
@@
 
-Description Dataset LEGI/IMFT
-
-\paragraph{Dataset 4}
-
-Description Dataset LEMTA
+The third dataset set is gathered using setup 3.
+The experiments are performed with polystyrene beads $d \sim 1~\textrm{mm}$ in salt water (isodense gravity currents) on a slope of $0^\circ$.
+The density difference between the lock and the ambient fluid is constant.
+Only the density varies from $~3 \%$ to $~55 \%$.
 
@@
 
+The set of dimensionless numbers, as defined in section \ref{sec:dimensionlessmap} and covered with the different datasets, is reported in figure~\ref{fig:fig2}, with in particular (a) $\phi$, (b) $\alpha$ and (c) $St$, all as a function of $\mathcal{R}e$.
+It is shown that they roughly cover the range $\phi\in[0.1\%,50\%]$, $\alpha \in [0^o,45^o]$, $St\in[10^{-3}, 10]$ and $Re\in[10^3, 10^6]$.
+The aspect ratio is only varied in the narrow range $a\in[2,3]$ (not shown in the figure).
+Its influence will not be specifically discussed in the following as it is found to be a second-order parameter in the range of parameters covered here.
+
 \subsection{Fitting procedure and front position analysis}
+
+%%%%%%%%%%%%%%%%%%%% Figure 3
+\begin{figure*}[ht]
+	\centering
+	\includegraphics{figure3.pdf}
+	\caption{\cyril{Panel (b) will be back to normal soon} (a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
+	\label{fig:fig3}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
 
@@
 After a short transient acceleration phase, most currents exhibit at first a constant velocity corresponding to the so-called \emph{slumping} regime.
-After this, the velocity can decrease, either because of settling induced buoyancy loss, or because the current enters the \emph{inertial} regime.
+After this, the velocity can decrease, either because of settling-induced buoyancy loss, or because the current enters the \emph{inertial} regime.
 
@@
 The fit is performed between $t/t_{0} = 5$, to avoid the initial transient phase, and $t/t_{0} = 30$, the transition time from the slumping to the inertial regime~\citep{refs}.
-Note that $x_{i}$ is kept constant equal to 0 for Dataset 1, 2 and 3, but had to be left free to adjust for the SedFoam simulations \textcolor{red}{comment this}.
+Note that $x_{i}$ is kept constant equal to 0 for datasets 1 and 2, but had to be left free to adjust for the SedFoam simulations \cyril{(comment this ?)}.
 For runs with strong dissipation ($\tilde{\lambda} > 2 {\times} 10^{-2}$), we add a quartic term to \eqref{eq:fit_eq} to improve the fit quality and obtain a better estimation of $\mathcal{F}r$ and $\tilde{\lambda}$.
-We checked that our results does not depend on the presence of this quartic term, and that the linear and quadratic ones remain much larger.
+We checked that our results do not depend on the presence of this quartic term and that the linear and quadratic ones remain much larger.
 
-For Dataset 4, the time resolution as well as the experiment duration does not allow to extract the dissipation. Therefore, we simply fit the whole available points with a linear function.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 3
-\begin{figure*}
-	\centering
-	\includegraphics{figure3.pdf}
-	\caption{(a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
-	\label{fig:fig3}
-\end{figure*}
+For dataset 3, the time resolution and the experiment duration do not allow for extracting the dissipation. Therefore, we simply fit the whole available points with a linear function.
 
@@
 
-The influence of the slope angle $\alpha$ on the dimensionless front velocity $\mathcal{F}r$ is presented in figure \ref{fig:fig4}.
-Even if the entire set of experiments and numerical simulations are reported here, results are highlighted for a nearly constant $St\approx 5\times 10^{-2}$ and $\phi < 0.4$ for clarity.
+The influence of the slope angle $\alpha$ on the dimensionless front velocity $\mathcal{F}r$ is presented in figure~\ref{fig:fig4}.
+Even if the entire set of experiments and numerical simulations are reported here, results are highlighted for $\phi < 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
 
-\begin{figure*}
+\begin{figure*}[ht]
 	\centering
 	\includegraphics{figure4.pdf}
-	\caption{\textbf{Influence of the bottom slope.} (a) Current Froude number as a function of the bottom slope.
-(b) Current Froude number corrected by a prefector $\sqrt{a}$ as a function of the bottom slope.
-Here, plain symbols corresponds to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
+	\caption{\textbf{Influence of the bottom slope} on the current Froude number.
+Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
 	\label{fig:fig4}
@@
 
-\begin{figure*}
+\begin{figure*}[ht]
 	\centering
@@
 	\caption{\textbf{Influence of particle settling.} Current Froude number $\mathcal{F}r$ and attenuation $\tilde{\lambda}$ as a function of the ratio between the Stokes number and the lock aspect-ratio, for different ranges of bottom slopes.
-Here, we selected runs for $\phi < 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
+Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
+Dotted lines mark either the average of points with $\phi < 0.45$ (a, c, e, g) or the 0 baseline (b, d, f, h).
+On (e, f), the dashed line indicates the average value for homogeneous saline currents ($St \equiv 0$) of dataset 1.}
 	\label{fig:fig5}
@@
 
-Discuss figure \ref{fig:fig5}
+In this section, we focus on the influence of settling on the current dynamics.
+The influence of the Stokes number $St$ on the dimensionless front velocity $\mathcal{F}r$ and attenuation $\tilde{\lambda}$ is presented in figure~\ref{fig:fig5} for different bottom slope $\alpha$.
+
+As shown by figure~\ref{fig:fig5}(a, c, e, f), the dimensionless front velocity does not depend on $St$ at the first order.
+However, this is not the case for the attenuation parameter $\tilde{\lambda}$.
+For $\alpha \approx 7^\circ$, the attenuation is constant below a threshold $St/a \sim 10^{-2}$, but increases rather linearly above this threshold (see the black line on figure~\ref{fig:fig5}(d), note the semi-log x-scale).
+
+The attenuation parameter can be linked to the transition length scale $X_{\rm t}$ as:
+\begin{equation}
+	\label{eq:link_attenuation_transition_length}
+	\frac{X_{\rm t}}{t_{\rm t}^{2}} \sim \lambda,
+\end{equation}
+where $t_{\rm t} = X_{\rm t}/U_{\rm c}$ is the corresponding transition time scale.
+%
+Assuming that the current decelerates due to its entry inside the \emph{inertial} regime, we use \eqref{eq:Xt_inertial}, which leads to $\tilde{\lambda} \sim cste$.
+On the other hand, if the current decelerates due to a settling-induced loss of buoyancy, we use \eqref{eq:Xt_settling}, which results in:
+\begin{equation}
+	\tilde{\lambda} \sim \frac{St}{a}.
+\end{equation}
 
@@
 \end{figure}
-
-Discuss figure \ref{fig:fig6}
+The influence of the parameter $\Phi$ on the Froude number is presented in Figure~\ref{fig:fig6}.
+The results demonstrate the existence of a critical fraction $\Phi_{c}$ above which the Froude number decreases.
+This reduction is characterized by an increase particle-particle and particle-fluid interactions which leads in wall friction and a decrease in inertial forces, resulting in a corresponding increase in viscous forces.
+This results in the decrease of the front velocity and the collapse of the current with the presence of a strong mass fraction gradient near the front.
 
 \subsection*{Acknowledgment}
+Sylvain Dauge for the full designing of Setup 2. Hervé Bellot and Brivaël Collin are thanked for experimental support.
 
 \subsection*{Fundings}
+ANR PALAGRAM
 
@@
 
-% \end{multicols*}
-
-\newpage
-%TC:endignore
-
 \end{document}

### 2023-05-23 _tests/Cyril/check_fits_analysis.py
Updating with new Julien's data

@@ -22,7 +22,7 @@ def determine_fit_props(author, alpha, run, params):
     params['L'].vary = True
     params['c'].vary = False
     params['d'].vary = False
-    t_bounds = [5, 30]
+    t_bounds = [5, 25]
     #
     if author == 'Jean':
         t_bounds = [0, 40]
@@ -82,9 +82,9 @@ SETUPS = {'Cyril': 'IMFT', 'Cyril/Marie': 'LEGI', 'Jean': 'LEMTA',
           'Julien': 'NUM', 'Rastello': 'LEGI'}
 
 # %% Loading data
-# list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_JULIEN2/*.nc')))
+list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_JULIEN2/*.nc')))
 # list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_JEAN/*.nc')))
-list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_MARIE/*.nc')))
+# list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_MARIE/*.nc')))
 # list_runs = np.array(glob.glob(os.path.join(input_path, 'runs_CYRIL/*.nc')))
 datasets = np.array([Dataset(run) for run in list_runs])
 
@@ -95,7 +95,7 @@ authors = np.array([d.author for d in datasets])
 diams = np.array([d.variables['d'][:].data for d in datasets])
 phi = np.array([d.variables['phi'][:].data for d in datasets])
 H0 = np.array([d.variables['H0'][:].data for d in datasets])
-Ha = np.array([d.variables['H_a'][:].data for d in datasets])
+# Ha = np.array([d.variables['H_a'][:].data for d in datasets])
 
 
 authors = np.array([d.author for d in datasets])
@@ -108,9 +108,9 @@ authors = np.array([d.author for d in datasets])
 # mask_runs = (authors == 'Cyril/Marie')
 # mask_runs = np.ones_like(alphas).astype('bool')
 # mask_runs = (alphas > 6) & (alphas < 8)
-# mask_runs = (alphas > 40)
+mask_runs = (alphas > 40)
 # mask_runs = (grains == 'PMMA')
-mask_runs = (H0/Ha < 0.3)
+# mask_runs = (H0/Ha < 0.3)
 # %% Loop over data file and analysis
 
 fig_fit, ax = plt.subplots(1, 1, constrained_layout=True)

### 2023-05-23 analysis.py

@@ -98,11 +98,7 @@ for par in params.keys():
                     max=higher_bounds[par])
 
 # %% Loading data
-list_runs = sorted(glob.glob(os.path.join(input_path, 'runs_CYRIL/*.nc'))) + \
-    sorted(glob.glob(os.path.join(input_path, 'runs_JEAN/*.nc'))) + \
-    sorted(glob.glob(os.path.join(input_path, 'runs_JULIEN2/*.nc'))) + \
-    sorted(glob.glob(os.path.join(input_path, 'runs_MARIE/*.nc')))
-
+list_runs = sorted(glob.glob(os.path.join(input_path, 'runs*/*.nc')))
 datasets = [Dataset(run) for run in list_runs]
 
 # %% Loop over data file and analysis

### 2023-05-23 figure_scripts/figure2.py

@@ -42,7 +42,7 @@ edgecolors = np.full_like(facecolors, 'k')
 mask_nosuspended = (authors == 'Rastello') & (H0/Ha < 1)
 edgecolors[mask_nosuspended] = 'tab:red'
 
-zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
+zorders = np.vectorize(lambda dataset: tp.dataset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
 rng = np.random.default_rng(1994)
 rng.shuffle(random_order)

### 2023-05-23 figure_scripts/figure4.py

@@ -61,7 +61,8 @@ edgecolors = np.array([to_rgba('k', a) for a in alphas])
 edgecolors[mask_nosuspended] = np.array(
     [to_rgba('tab:red', 0.4) for a in alphas[mask_nosuspended]])
 
-zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
+zorders = np.vectorize(
+    lambda dataset: tp.dataset_zorder2[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
 rng = np.random.default_rng(1994)
 rng.shuffle(random_order)

### 2023-05-23 figure_scripts/figure5.py

@@ -54,7 +54,7 @@ edgecolors = np.array([to_rgba('k', a) for a in alphas])
 edgecolors[mask_nosuspended] = np.array(
     [to_rgba('tab:red', 0.4) for a in alphas[mask_nosuspended]])
 
-zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
+zorders = np.vectorize(lambda dataset: tp.dataset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
 rng = np.random.default_rng(1994)
 rng.shuffle(random_order)

### 2023-05-23 figure_scripts/figure6.py

@@ -40,7 +40,7 @@ edgecolors = np.full_like(facecolors, 'k')
 mask_nosuspended = (authors == 'Rastello') & (H0/Ha < 1)
 edgecolors[mask_nosuspended] = 'tab:red'
 
-zorders = np.vectorize(lambda dataset: tp.datset_zorder[dataset])(dataset_idx)
+zorders = np.vectorize(lambda dataset: tp.dataset_zorder[dataset])(dataset_idx)
 random_order = np.arange(zorders.size)
 rng = np.random.default_rng(1994)
 rng.shuffle(random_order)

### 2023-05-23 figure_scripts/template.py

@@ -63,7 +63,8 @@ color_datasets = {'SedFoam': '#780116', '3': '#FE7F2D',
 marker_style = {'glass beads': 'o', 'silica sand': 'h', 'Hydrogels': '*',
                 'PMMA': 'D', 'polystyren beads': 'X', 'SedFoam': 's'}
 
-datset_zorder = {'1': 0, '2': 1, '3': 2, 'SedFoam': 2}
+dataset_zorder = {'1': 0, '2': 1, '3': 2, 'SedFoam': 2}
+dataset_zorder2 = {'1': 0, '2': 2, '3': 2, 'SedFoam': 1}
 
 # %% corresponding legend
 

### 2023-05-23 analysis.py
Changing lambda to tau in fit

@@ -23,8 +23,8 @@ def create_variable(netcdf_group, name, data, dimensions=None, std=None,
         var.comments = comments
 
 
-def polyFULL(t, Fr, L, xi, c, d):
-    return xi + Fr*t - L*t**2/2 + c*t**3 + d*t**4
+def polyFULL(t, Fr, tau, xi, c, d):
+    return xi + Fr*(t - t**2/tau) + c*t**3 + d*t**4
 
 
 def Stokes_Velocity(d, mu, rho_p, rho_f, g):
@@ -34,14 +34,14 @@ def Stokes_Velocity(d, mu, rho_p, rho_f, g):
 def determine_fit_props(author, alpha, run, params):
     params['xi'].vary = False
     params['Fr'].vary = True
-    params['L'].vary = True
+    params['tau'].vary = True
     params['c'].vary = False
     params['d'].vary = False
     t_bounds = [5, 30]
     #
     if author == 'Jean':
         t_bounds = [0, 40]
-        params['L'].vary = False
+        params['tau'].vary = False
         params['xi'].vary = True
     if author == 'Julien':
         params['xi'].vary = True
@@ -86,11 +86,15 @@ model = Model(polyFULL)
 params = model.make_params()
 
 # parameter properties (Non dim.)
-p0 = {'Fr': 0.4, 'xi': 0, 'L': 0, 'c': 0, 'd': 0}
+p0 = {'Fr': 0.4, 'xi': 0, 'tau': 100, 'c': 0, 'd': 0}
+
+# lower_bounds = {'Fr': 0, 'xi': -np.inf,
+#                 'L': -0.05, 'c': -1e-3, 'd': -1e-4}
+# higher_bounds = {'Fr': 2, 'xi': np.inf, 'L': 0.1, 'c': 1e-3, 'd': 1e-4}
 
 lower_bounds = {'Fr': 0, 'xi': -np.inf,
-                'L': -0.05, 'c': -1e-3, 'd': -1e-4}
-higher_bounds = {'Fr': 2, 'xi': np.inf, 'L': 0.1, 'c': 1e-3, 'd': 1e-4}
+                'tau': -np.inf, 'c': -1e-3, 'd': -1e-4}
+higher_bounds = {'Fr': 2, 'xi': np.inf, 'tau': np.inf, 'c': 1e-3, 'd': 1e-4}
 
 # set parameter bounds
 for par in params.keys():
@@ -148,19 +152,19 @@ for i, d in enumerate(datasets):
     # Make fit
[…]

### 2023-05-23 figure_scripts/figure5.py

@@ -24,11 +24,11 @@ list_runs = glob.glob(os.path.join(path_data, '*.nc'))
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% create data vectors
-alpha, phi, St, H0, L0, Fr, L = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
+alpha, phi, St, H0, L0, Fr, tau = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
                                            d.variables['St'][:].data, d.variables['H0'][:].data,
                                            d.variables['L0'][:].data, d.variables['Fr'][:].data,
-                                           d.variables['L'][:].data
-                                           ] for d in datasets]).T
+                                           d.variables['tau'][:].data
+                                             ] for d in datasets]).T
 
 Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
               else d.variables['H0'][:].data for d in datasets])
@@ -77,7 +77,7 @@ params['c'].vary = False
 
 mask_alpha = (alpha > alpha0[1] - alpha_pad) & (alpha < alpha0[1] + alpha_pad)
 
-result = model.fit(L[mask_alpha], params, x=(St/a)[mask_alpha])
+result = model.fit(1/tau[mask_alpha], params, x=(St/a)[mask_alpha])
 
 # # %% Figure
 
@@ -87,7 +87,7 @@ fig, axarr = plt.subplots(4, 3, constrained_layout=True,
 for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
     mask_alpha = (alpha > a0 - alpha_pad) & (alpha < a0 + alpha_pad)
     mask = (mask_alpha)[plot_idxs]
-    for i, (var, ax) in enumerate(zip([Fr, L], axarr_sub.flatten())):
+    for i, (var, ax) in enumerate(zip([Fr, 1/tau], axarr_sub.flatten())):
         tp.mscatter((St/a)[plot_idxs][mask], var[plot_idxs][mask], ax=ax, m=markers[plot_idxs][mask],
                     facecolors=facecolors[plot_idxs][mask], edgecolors=edgecolors[plot_idxs][mask], lw=0.5)
         #
@@ -98,10 +98,10 @@ for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
[…]

### 2023-05-23 analysis.py
Re-changing tau to lamba (only for fit to avoid infinte values (lambda sim 0)

@@ -23,8 +23,8 @@ def create_variable(netcdf_group, name, data, dimensions=None, std=None,
         var.comments = comments
 
 
-def polyFULL(t, Fr, tau, xi, c, d):
-    return xi + Fr*(t - t**2/tau) + c*t**3 + d*t**4
+def polyFULL(t, Fr, lamb, xi, c, d):
+    return xi + Fr*(t - lamb*t**2) + c*t**3 + d*t**4
 
 
 def Stokes_Velocity(d, mu, rho_p, rho_f, g):
@@ -34,14 +34,14 @@ def Stokes_Velocity(d, mu, rho_p, rho_f, g):
 def determine_fit_props(author, alpha, run, params):
     params['xi'].vary = False
     params['Fr'].vary = True
-    params['tau'].vary = True
+    params['lamb'].vary = True
     params['c'].vary = False
     params['d'].vary = False
     t_bounds = [5, 30]
     #
     if author == 'Jean':
         t_bounds = [0, 40]
-        params['tau'].vary = False
+        params['lamb'].vary = False
         params['xi'].vary = True
     if author == 'Julien':
         params['xi'].vary = True
@@ -86,15 +86,15 @@ model = Model(polyFULL)
 params = model.make_params()
 
 # parameter properties (Non dim.)
-p0 = {'Fr': 0.4, 'xi': 0, 'tau': 100, 'c': 0, 'd': 0}
+p0 = {'Fr': 0.4, 'xi': 0, 'lamb': 0, 'c': 0, 'd': 0}
 
 # lower_bounds = {'Fr': 0, 'xi': -np.inf,
 #                 'L': -0.05, 'c': -1e-3, 'd': -1e-4}
 # higher_bounds = {'Fr': 2, 'xi': np.inf, 'L': 0.1, 'c': 1e-3, 'd': 1e-4}
 
 lower_bounds = {'Fr': 0, 'xi': -np.inf,
-                'tau': -np.inf, 'c': -1e-3, 'd': -1e-4}
-higher_bounds = {'Fr': 2, 'xi': np.inf, 'tau': np.inf, 'c': 1e-3, 'd': 1e-4}
+                'lamb': -0.02, 'c': -1e-3, 'd': -1e-4}
+higher_bounds = {'Fr': 2, 'xi': np.inf, 'lamb': 0.2, 'c': 1e-3, 'd': 1e-4}
 
 # set parameter bounds
 for par in params.keys():
[…]

### 2023-05-23 figure_scripts/figure5.py

@@ -24,11 +24,11 @@ list_runs = glob.glob(os.path.join(path_data, '*.nc'))
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% create data vectors
-alpha, phi, St, H0, L0, Fr, tau = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
-                                           d.variables['St'][:].data, d.variables['H0'][:].data,
-                                           d.variables['L0'][:].data, d.variables['Fr'][:].data,
-                                           d.variables['tau'][:].data
-                                             ] for d in datasets]).T
+alpha, phi, St, H0, L0, Fr, lamb = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
+                                              d.variables['St'][:].data, d.variables['H0'][:].data,
+                                              d.variables['L0'][:].data, d.variables['Fr'][:].data,
+                                              d.variables['lamb'][:].data
+                                              ] for d in datasets]).T
 
 Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
               else d.variables['H0'][:].data for d in datasets])
@@ -77,7 +77,7 @@ params['c'].vary = False
 
 mask_alpha = (alpha > alpha0[1] - alpha_pad) & (alpha < alpha0[1] + alpha_pad)
 
-result = model.fit(1/tau[mask_alpha], params, x=(St/a)[mask_alpha])
+result = model.fit(lamb[mask_alpha], params, x=(St/a)[mask_alpha])
 
 # # %% Figure
 
@@ -87,7 +87,7 @@ fig, axarr = plt.subplots(4, 3, constrained_layout=True,
 for a0, axarr_sub in zip(alpha0, axarr[:, 1:]):
     mask_alpha = (alpha > a0 - alpha_pad) & (alpha < a0 + alpha_pad)
     mask = (mask_alpha)[plot_idxs]
-    for i, (var, ax) in enumerate(zip([Fr, 1/tau], axarr_sub.flatten())):
+    for i, (var, ax) in enumerate(zip([Fr, lamb], axarr_sub.flatten())):
         tp.mscatter((St/a)[plot_idxs][mask], var[plot_idxs][mask], ax=ax, m=markers[plot_idxs][mask],
[…]

### 2023-05-24 figure_scripts/figure4.py
Adding Froude model + theory on fig 6

@@ -8,6 +8,7 @@ import template as tp
 from matplotlib.colors import to_rgba
 from matplotlib.lines import Line2D
 from netCDF4 import Dataset
+from models import Birman
 
 plt.rcParams['xtick.top'] = False
 
@@ -18,10 +19,6 @@ def ang2sin(ang):
 
 def sin2ang(sin):
     return np.degrees(np.arcsin(sin))
-
-
-def Birman(ang):
-    return -0.1824*np.radians(ang)**2 + 0.2781*np.radians(ang) + 0.4871
 
 
 # %% Load data

### 2023-05-24 figure_scripts/figure6.py

@@ -7,6 +7,7 @@ import matplotlib.transforms as mtransforms
 import numpy as np
 import template as tp
 from netCDF4 import Dataset
+from models import Froude, Krieger_viscosity
 
 plt.rcParams['figure.constrained_layout.hspace'] = 0
 plt.rcParams['figure.constrained_layout.h_pad'] = 0.005
@@ -46,6 +47,12 @@ rng = np.random.default_rng(1994)
 rng.shuffle(random_order)
 plot_idxs = np.lexsort((random_order, zorders))
 
+# %% theory
+phi_m = 0.585
+phi_plot = np.logspace(-3, phi_m, 200)
+eta = Krieger_viscosity(phi_plot, phi_m)
+Fr_th = Froude(0, eta, 7e4)
+
 # %% masks for plot
 alphas = [0, 45]
 alpha_pad = 1.5
@@ -71,8 +78,13 @@ for alpha0, ax in zip(alphas, axarr.flatten()):
     ax.set_ylabel(r'Froude number, $\mathcal{F}r$')
     ax.set_ylim([0, 1.59])
 
+
+axarr[0].plot(phi_plot, Fr_th, ls='-', color='k')
+axins.plot(phi_plot, Fr_th, ls='-', color='k')
+
 ax.set_xlabel(r'Volume fraction, $\phi$')
 ax.set_xscale('log')
+ax.set_xlim([0.0035, 1.2])
 
 for ax, l in zip(axarr.flatten(), 'ab'):
     trans = mtransforms.ScaledTranslation(

### 2023-05-24 figure_scripts/models.py

@@ -1,0 +1,76 @@
+import numpy as np
+
+# %% Birman et al 2007
+
+
+def Birman(ang):
+    return -0.1824*np.radians(ang)**2 + 0.2781*np.radians(ang) + 0.4871
+
+
+# %% energetic bilan model (extension of Gadal et al. 2023)
+
+def Froude(theta, eta, Re, Fr0=0.5, a=1, r=1, Cd=0.4, E=500, A=3):
+    """
+    Calculate $Fr = U_c/u_0$, with $u_0 = \sqrt{\cos\theta\delta_\rho g h_0/rho_c}$ using an energetic bilan similar to that of Gadal et al. 2023. 
+
+    Parameters
+    ----------
+    Fr0 : scalar, array_like
+        Limit for $r=1$, $theta \to 0$, $Cd \to 0$, $Re \to \infty$.
+    theta : scalar, array_like
+        Bottom slope, in degree.
+    eta : scalar, array_like
+        Non-dimensinal viscosity $\eta/\eta_f$.
+    Re : scalar, array_like
+        Reynlods number $Re = \rho_0 u_0 h_{0}/eta_{f}$
+    a : scalar, array_like
+        Lock aspect ratio, $a = h_0/L_0$
+    r : scalar, array_like
+        Density ratio, $r = \rho_0/rho_a$
+    Cd : scalar, array_like
+        Drag coefficient
+    E : scalar, array_like
+        Viscous dissipation coefficient
+    A : scalar, array_like
+        Slope effect coefficient
+
+    Returns
+    -------
+    scalar, array_like
+        Non-dimensional velocity $Fr = U_c/u_0$, with $u_0 = \sqrt{\cos\theta\delta_\rho g h_0/rho_c}$.
+    """
+    drag = (1 + Cd*a/r)
+    viscous = E*eta/Re
+    Fr = (np.sqrt(viscous**2 + Fr0**2*drag*(1 +
+          A*np.tan(np.radians(theta)))) - viscous)/drag
+    return Fr
+
+# %% Viscosity models
+
+
+def Krieger_viscosity(phi, phi_m=0.585, eta_cr=5/2):
+    # (Krieger and Dougherty (1959) model, from Stickel and Powell (2005))
+    return (1 - phi/phi_m)**(-eta_cr*phi_m)
+
+
[…]

### 2023-05-31 paper/main.tex
Update paper

@@
 \usepackage{amsmath}
+\usepackage{cuted}
 
@@
 \newcommand*{\cyril}{\textcolor{ForestGreen}}
+\newcommand*{\marie}{\textcolor{Cyan}}
+
+\definecolor{jeanc}{RGB}{253,129,43}
+\newcommand*{\jean}{\textcolor{jeanc}}
+
+\newcommand*{\julien}{\textcolor{brown}}
 
@@
 \title{Macroscopic description of particle-laden gravity currents: overall perspectives}
-\author{Gadal C., Schneider J., Bonamy C., Chauchat J., Dossmann Y., Kiesgen de Richter S., Mercier M., Naaim F., Rastello M., Lacaze L.}
+
+\author[1]{Gadal C.}
+\author[2]{Schneider J.}
+\author[3]{Bonamy C.}
+\author[3]{Chauchat J.}
+\author[2]{Dossmann Y.}
+\author[2]{Kiesgen de Richter S.}
+\author[1]{Mercier M.}
+\author[4]{Naaim F.}
+\author[3]{Rastello M.}
+\author[1]{Lacaze L.}
+
+\affil[1]{Institut de Mécanique des Fluides de Toulouse (IMFT), Université de Toulouse, CNRS, Toulouse, France}
+\affil[2]{Affil1}
+\affil[3]{Univ. Grenoble Alpes, CNRS, Grenoble INP, LEGI, 38000 Grenoble, France}
+\affil[4]{Affil1}
+
 \date{{\it\color{red}Deadline: May 31$^{st}$}}
@@
 		The main objective of the present chapter is to provide a comprehensive flow regime associated with the relevant dimensionless parameters which characterize the current at the macroscopic scale, i.e.
-average over the current depth.
-
+average over the current depth \marie{in the early moments of the flow.
+(?)}
 	\end{abstract}
@@
 The basic approach to describe a lock-release evolution is built upon characteristics solutions of the hyperbolic shallow water equations for an inviscid dense liquid of density $\rho_c$ slumping on a horizontal bottom under an inertialess and weightless ambient fluid.
-Among others, this specifically allows quantifying a most obvious scaling analysis solution reading for the front velocity $U_c$ reading
+Among others, this specifically allows quantifying a most obvious scaling analysis solution for the front velocity $U_c$ reading
 \begin{equation}
@@
 \end{equation}
-with $g$ the gravity and $H_0$ the initial height of the fluid volume being released.
-The method of characteristics suggests a coefficient of proportionality of exactly $2$ between velocities $U_c$ and $\sqrt{gH_0}$.
+with $g$ \marie{gravitational acceleration} and $H_0$ the initial height of the fluid volume being released.
+The method of characteristics suggests a coefficient of proportionality of exactly $2$ between velocities $U_c$ and $\sqrt{g H_0}$.
 An extension towards a weighted ambient fluid of density $\rho_a$ shall lead to
 \begin{equation}
-	U_c\propto \sqrt{g'H_0},\quad\mbox{with}\quad g'=\frac{\rho_c-\rho_a}{\rho_c}g.
+	U_c\propto \sqrt{\tilde{g}H_0},\quad\mbox{with}\quad \tilde{g}=\frac{\rho_c-\rho_a}{\rho_c}g.
 	\label{eq:solUc}
@@
 Knowledge of the front velocity then requires a local description of the flow close to the front, which in turn adds the characteristic height of the ambient as a required control parameter, as suggested by \citet{Benjamin1968}.
-In the case of lock-release configuration, it would suggest an influence of the ratio between reservoir height $H_0$ and ambient height $H$.
+In the case of lock-release configuration, it would suggest an influence of the ratio between reservoir height $H_0$ and ambient height $H_\textup{a}$.
 Nevertheless, obtained results allow linking the front velocity to the height of the current close to the front, which is a priori unknown in a lock-release device, i.e.
 the front relation includes two unknowns \citep{Benjamin1968}.
-Models have been proposed to circumvent this issue, leading for instance to a coefficient of proportionality $0.5 (\rho_c/\rho_a)^{1/2}(H/H_0)^{1/3}$ between the two velocities in \eqref{eq:solUc} \citep{Huppert1980}.
-The latter coefficient is relevant for not two deep ambient $H$ compared to the reservoir height $H_0$, which is actually the case in most experiments.
-Most configurations correspond to $H=H_0$, which would lead to a value of $0.5$.
+Models have been proposed to circumvent this issue, leading for instance to a coefficient of proportionality $0.5 (\rho_c/\rho_a)^{1/2}(H_\textup{a}/H_0)^{1/3}$ between the two velocities in \eqref{eq:solUc} \citep{Huppert1980}.
+The latter coefficient is relevant for not too deep ambient $H_\textup{a}$ compared to the reservoir height $H_0$, which is actually the case in most experiments.
+\marie{Many} configurations correspond to $H_\textup{a}=H_0$, which would lead to a value of $0.5$.
 Even if several experiments and numerical simulations confirm such a value around $0.5(\rho_c/\rho_a)^{1/2}$, its exact value remains still unclear, and probably dependent on several physical processes.
@@
 	\label{eq:Xt_inertial}
-	\displaystyle    X_{\rm t}=  \frac{2L_0 }{\sqrt{g'H_0}/U_c - 1},
+	\displaystyle    X_{\rm t}=  \frac{2L_0 }{\sqrt{\tilde{g}H_0}/U_c - 1},
 \end{equation}
@@
 	\label{eq:Xt_drag}
-	\displaystyle    X_{\rm t}=H_0  \frac{g'H_0}{C_dU_c^2},
+	\displaystyle    X_{\rm t}=H_0  \frac{\tilde{g}H_0}{C_dU_c^2},
 \end{equation}
+\marie{(Pas super clair pour moi mais je vais réfléchir au pourquoi du comment. ;-))}
 with $C_d$ some drag coefficient induced by friction at the bottom and by entrainment at the interface.
@@
 Details of such transitions actually depend on the complex details of the flow and drag induced by the current intruding the ambient and the friction at the bottom.
-These transitions then become case-dependent and required parametrization.
+These transitions then become case-dependent and require parametrization.
 For instance, following experiments of \cite{Rottman1983}, it was found that???
@@
 %At one end, the own settling velocity of heavy particle only allow the existence of these gravity currents on finite time scale and finite length scale.
-Nevertheless, the prediction of these finite scales is difficult as it shall strongly depends on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface, which modifies entrainment and associated dissipation.
+Nevertheless, the prediction of these finite scales is difficult as it shall strongly depend on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface, which modifies entrainment and associated dissipation.
 In turn, this can modify the macroscopic behaviour of the current, as for instance the dynamics of the front.
@@
 
-Following the balance scaling as done far, a transition between the equivalent slumping regime as described for gravity current and deposition of particle.
-Instead of dynamical scaling as done previously, here a kinematic scaling is obtained as the balancing time scale of current propagation and time scale of particle settling.
+\cyril{Problème cette phrase: Following the balance scaling as done far, a transition between the equivalent slumping regime as described for gravity current and deposition of particle.} Instead of dynamical scaling as done previously, here a kinematic scaling is obtained as the balancing time scale of current propagation and time scale of particle settling.
 This leads to
@@
 \end{equation}
-in which $V_S$ is the settling velocity of the solid particles within the PLGCs. Of course, such transition disappears for the neutrally-buoyant situation described previously.
+\marie{Ne faudrait il pas un angle dans le rapport cosinus ou sinus?}
+in which $V_S$ is the settling velocity of the solid particles within the PLGCs.
+Note that such transition shall disappear for the neutrally-buoyant situation of PLGCs.
+However, it has been observed that mixing at the interface can modify local particle concentration at the front leading to a somehow similar behaviour as settling \cite{Schneider???}.
 Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear.
@@
 	\includegraphics{figure2.pdf}
-	\caption{\textbf{Explored parameter space.} Each symbol corresponds to a single experimental run.}
+	\caption{\textbf{Explored parameter space.} Each symbol corresponds to a single experimental run.\marie{(Définir ici le pourquoi des symboles avec un contour rouge.)}}
 	\label{fig:fig2}
@@
 
-Here, the dimensionless parameters are based on a characteristic length $L_0$ and a characteristic velocity $U_0=\sqrt{g' H_0}$ where $L_0$ and $H_0$ are the horizontal length and depth of the initial container, respectively, $g'$ is the modified gravity $ g \cos{\alpha}(\rho_m-\rho_a)/\rho_m$ with $\rho_m$ the current mixture density (analogous to $\rho_c$ as for current in \eqref{eq:solUc}) and $\rho_a$ the current density and ambient fluid density respectively.
+According to the previous discussion, the dimensionless parameters are based on a characteristic length $L_0$ and a characteristic velocity $U_0=\sqrt{g' H_0}$ where $L_0$ and $H_0$ are the horizontal length and depth of the initial container, respectively.
+$g'$ is the modified gravity $ g \cos{\alpha}(\rho_m-\rho_a)/\rho_a$ with $\rho_m$ the current mixture density (analogous to $\rho_c$ as for current in \eqref{eq:solUc}) and $\rho_a$ the ambient fluid density.
 Note that $g'$ is slightly different from the definition in section \ref{sec:intro}.
@@
 The velocity scale $U_0$ corresponds to an idealized transfer of initial potential energy to kinetic energy along the slope, i.e.
-a collapse of the initial column.
-Accordingly,
+a collapse of the initial column.\marie{Si on dit idealized transfer il faudrait mettre le facteur 2 non?} Accordingly,
 \begin{equation}
@@
 \end{equation}
+\marie{j'ai perdu le fil à un moment dans les figures.
+Est ce qu'on a essayé de prendre en compte l'angle dans le Stokes ou pas du genre $Vscos\alpha/U_0$.
+Je suppose que oui et que cela n'a rien donné et que par conséquent on a opté pour revenir à du standard?}
 with $\eta_f$ the interstitial fluid viscosity and $V_s$ the Stokes settling velocity in the interstitial fluid phase labelled $f$.
-Note that $At$ usually intervenes only in the modified gravity $g'$ for Boussinesq currents, which is mostly the case for density currents \citep{Bonometti2011,???}
+Note that $At$ usually intervenes only in the modified gravity $g'$ for Boussinesq currents, which is mostly the case for density currents \citep{Bonometti2011,???}\marie{mostly le case quand on est dans l'eau mais ceux qui font des gaz lourds ou légers par exemple (Hanouzet 2019) ne sont pas Boussinesq.
+Peut-être modérer un peu en frequently plutôt que mostly?}
 
@@
 Then prediction of the different flow regimes associated with PLGCs shall be entirely included in these observables.
-In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at ``rest'', 3 dam-break-type devices have been used to cover a wide range of control parameters, These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.
+In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at ``rest'', 3 dam-break-type devices have been used to cover a wide range of control parameters.
+These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.
 
@@
 The current evolution is followed by a camera at 120 Hz and using a backlight as a light source.
+\marie{donner largeur $W$ et $H_a$?}
 
 \paragraph{Set-up 2}
-
-Specifications of LEGI set-up
+\marie{The second set-up (figure~\ref{fig:fig1}(b)) has $L_0+L_1=370~\textup{cm}$ with $L_0 = 30~\textrm{cm}$ for runs with PMMA particles and $L_0 = 10~\textrm{cm}$ for runs with silica sand or hydrogel particles (see section~\ref{datasets} for particles specifications).
+The overall lock and flume width is $W= 10~\textrm{cm}$.
+The
+upper boundary in the flume is a lid and the only free surface is in the lock region.
+The ambient fluid is tap water at room temperature and $H_a=50~\textrm{cm}$.
+The flume can be tilted at an angle $\alpha =[0^\circ,45.4^\circ]$ as shown in the 2D sketch of figure~\ref{fig:fig1}(b).
+$H_0$ is defined as the average suspension height perpendicular to the flume floor inside the lock at gate opening time.
+LED plates illuminate the flume from above.
+The flow dynamic is recorded with a Phantom$^\copyright$ Miro C110 camera at 50 frames per second.
+Recorded images are treated with a Matlab$^\copyright$ image processing.}
 
@@
 
-The third set-up (figure~\ref{fig:fig1}(c)) has $L_0+L_1=150~\textup{cm}$ with $L_0 = 20~\textrm{cm}$.
-The upper boundary is a free surface and the tank remains horizontal ($\alpha = 0^\circ$).
-Here, $H_{0}$ is between 0.2~cm and 0.3~cm depending on experiments.
-The current evolution is followed by a camera at 10 Hz and using a backlight as a light source.
-
+\jean{The third set-up (figure~\ref{fig:fig1}(c)) has $L_0+L_1=150~\textup{cm}$ with $L_0 = 20~\textrm{cm}$.
+The upper boundary is a free surface and the tank remains horizontal ($\alpha = 0^\circ$) as shown in the 2D sketch of figure~\ref{fig:fig1}(c).
+Here, $H_{0}$ is between 20~cm and 30~cm depending on experiments.
+The overall lock and flume width is $W= 10~\textrm{cm}$.
+The current evolution is followed by a camera at 10 Hz and using a backlight as a light source.}
+\marie{donner largeur $W$ et $H_a$?}
 % The suspension is made by strongly stirring a known mass of particles within the whole water column using a chemical mixer, which is stopped just before opening the gate.
 
-\subsection{Datasets}
+\paragraph{Two-fluid model}
 
-On these 3 devices, $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below). They are classified into 4 different datasets, described below.
+A two-fluid model, SedFoam \citep{chauchat2017}, is used to simulate lock-release configuration from set-up 1 and 2.
+This model available for download at Zenodo \citep{bonamy2023} is implemented in the open-source computational fluid dynamics toolbox OpenFOAM and has been extensively validated for various sediment transport problems such as sheet-flow \citep{cheng2016,chauchat2022,mathieu2021,mathieu2022}, scour \citep{nagel2020,tsai2022}, ripples \citep{salimi2021a,salimi2021b}, avalanches \citep{montella2021} and granular collapse \cite{montella2023}.
+The simulations presented herein are 2D vertical and turbulence-averaged, the turbulence model is a $k-\epsilon$ model with additional two-phase flow terms for density stratification and turbulent drag dissipation.
+The closure for the particle-particle interactions is a frictional-collisional model based on a Coulomb rheology for the frictional part and the kinetic theory of granular flows for the collisional part.
+The two contributions are simply added as described in \cite{chauchat2017}.
+In order to describe the influence of the slope, the total pressure is solved and the gravity vector is rotated while the mesh is kept aligned with the $x$ and $y$ axis.
+
+\subsection{Datasets\label{datasets}}
+
+On these 3 devices, $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below).
+They are classified into 4 different datasets, described below.\marie{Utilisation de $N$ deux fois peut donner l'impression qu'il y a le même nombre de simus que
+	d'expés non?}
 
@@
 \paragraph{Dataset 2}
-
-Description Dataset LEGI
+\marie{The second dataset is gathered using set-up 2.
+Three different types of particles have been used in order to vary $d$, $\rho_p$ and as a direct consequence $V_s$ and $St$.
+A first series of experiments is made of the same silica sand grains as in dataset 1 with the same stirring procedure and varying the bottom slope from $0^\circ$ to $15^\circ$.
+A second series of experiments is made of PMMA particles with $\rho_p=1190~\textrm{kg/m}^{3}$ and diameter $d\sim 250~\mu\textrm{m}$.
+Part of the runs had a manual stirring with a rod and $H_0=H_a$, while others deal with a release of a deposited layer of particles and no mixing leading to $H_0<H_a$ \cyril{(see symbols with red edges on each figure).}.
+The bottom slope is varied between $0^\circ$ and $45.4^\circ$.
+A third type of particles corresponds to hydrogel particles with densities $\rho_p\in [1003, 1045]~\textrm{kg/m}^{3}$ and $d\in [0.3, 1.6]~\textrm{cm}$.
+For most of them the flume tilt was set to $45.4^\circ$ and experiments began with the opening of the lock on a deposited layer ($H_0<H_a$) A couple runs only had a lower tilt and were released from a stirred configuration.}
 
@@
 
-The third dataset set is gathered using setup 3.
-The experiments are performed with polystyrene beads $d \sim 1~\textrm{mm}$ in salt water (isodense gravity currents) on a slope of $0^\circ$.
-The density difference between the lock and the ambient fluid is constant.
-Only the density varies from $~3 \%$ to $~55 \%$.
+\jean{The third dataset is gathered using setup 3.
+The experiments are performed using a neutrally-buoyant fluid-particle mixture made of polystyrene beads $d \sim 1~\textrm{mm}$ with density $\rho_p \sim 1038.5~\textrm{kg/m}^{3}$ and saltwater for the heavy fluid kept inside the lock.
+The density difference between the lock and the ambient fluid is constant for each experiments.
+Only the mass fraction varies from $~3 \%$ to $~55 \%$.}
 
-\paragraph{SedFoam}
+\paragraph{Dataset 4}
 
-Description Dataset SedFoam
+This dataset has been obtained using the two-fluid model described above.
+The domain size is 3.8m long and 0.5m height and the typical grid resolution is 1.25mm in the streamwise direction and it varies between 0.25 and 3.5 mm in the wall-normal direction.
+The time step is adaptative following a criteria on the Courant number which must remain smaller than 0.3 for the phase velocities as well as for the relative velocity between the phases (the maximum allowed time-step during the simulation is set to $10^{-3}$s).
+The lock is $W_0=0.3$m wide and $H_0=0.5$m high corresponding to the domain height.
+The numerical schemes are second order in time and space : backward for time, a blend of centered and upwind schemes for advection (limitedLinear in OpenFOAM) and centered scheme for diffusion.
+The initial condition is fluid and particle phase at rest ($u_f=u_p=0$) and the initial volume fraction of particles $\phi_0$ is set to a constant value in the region $x\in[-0.5 ; 0]$ m and $y\in[0;0.5]m$.
+At t=0, the governing equations are numerical integrated until $t_{end}=80$s maximum.
+The front position is determined as the most downstream position of the isocontour of volume fraction $\phi_f=10^{-3}$ at a frequency of 10Hz.
+This criteria has been demonstrated to be robust for all the simulations presented in this work.
+The particles properties are $e=0.8$, the restitution coefficient for binary collisions, $\mu_s=0.53$, the macroscopic friction coefficient.
+Two sets of simulations have been performed, one using glass beads in water $\rho_p/\rho_f=2.65$ with diameter $d_p=60\ \mu$m and initial volume fractions $\phi_0\in[0.015 ; 0.1]$ ; the other one with PMMA particles in water $\rho_p/\rho_f=1.19$ with diameters $d_p \in [ 100 ; 650]\ \mu$m and initial volume fractions $\phi_0\in[0.0144 ; 0.5]$.
+For both configurations the slope angle $\alpha$ has been varied in the range $\alpha \in [0 ; 45]$ degrees.\\
 
 The set of dimensionless numbers, as defined in section \ref{sec:dimensionlessmap} and covered with the different datasets, is reported in figure~\ref{fig:fig2}, with in particular (a) $\phi$, (b) $\alpha$ and (c) $St$, all as a function of $\mathcal{R}e$.
-It is shown that they roughly cover the range $\phi\in[0.1\%,50\%]$, $\alpha \in [0^o,45^o]$, $St\in[10^{-3}, 10]$ and $Re\in[10^3, 10^6]$.
-The aspect ratio is only varied in the narrow range $a\in[2,3]$ (not shown in the figure).
-Its influence will not be specifically discussed in the following as it is found to be a second-order parameter in the range of parameters covered here.
-
-\subsection{Fitting procedure and front position analysis}
+It is shown that they roughly cover the range $\phi\in[0.1\%,50\%]$\marie{(A voir Cyril comment tu veux gérer le $\phi$ pour les petites hydrogels qui ont un $\phi$ plus élevé.)}, $\alpha \in [0^o,45^o]$, $St\in[10^{-3}, 10]$ and $Re\in[10^3, 10^6]$.
+The aspect ratio is only varied in the narrow range $a\in[2,3]$ (not shown in the figure)\marie{(pour les manips LEGI on n'est pas toujours dans cette gamme, on est des fois plus petit entre autres quand on n'a pas mixé et également parfois plus grand (a priori $a\in[0.5,5]$)}.
+Its influence will not be specifically discussed in the following as it is found to be a second-order parameter in the range of parameters covered here.\marie{(found ou assumed?)}
 
@@
 	\includegraphics{figure3.pdf}
-	\caption{\cyril{Panel (b) will be back to normal soon} (a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
+	\caption{(a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
 	\label{fig:fig3}
@@
 
+\subsection{Fitting procedure and front position analysis}
+
 For each experimental/numerical run, we extract the current front position as a function of time (see Fig.
 \ref{fig:fig3}a).
-A few selected runs are shown on figure~\ref{fig:fig3}b in non-dimensional time coordinates.
-After a short transient acceleration phase, most currents exhibit at first a constant velocity corresponding to the so-called \emph{slumping} regime.
+A few selected runs are shown in figure~\ref{fig:fig3}b in non-dimensional time coordinates.\marie{($t_0$ n'est pas défini dans le papier avant je pense sauf erreur de ma part.)} After a short transient acceleration phase, most currents exhibit at first a constant velocity corresponding to the so-called \emph{slumping} regime.
 After this, the velocity can decrease, either because of settling-induced buoyancy loss, or because the current enters the \emph{inertial} regime.
 
-It can be qualitatively seen that increasing the bottom slope increases the slumping velocity, while increasing the Stokes number shortens the duration of the constant-velocity regime.
+It can be qualitatively seen that increasing the bottom slope increases the slumping velocity while increasing the Stokes number shortens the duration of the constant-velocity regime.
 %
-To properly quantify that, we fit all front propagation curves (except Dataset 4) with the following equation:
+To properly quantify that, we fit all front propagation curves (except dataset 3) with the following equation:
 
@@
 	\label{eq:fit_eq}
-	\frac{x_{\rm f}}{L_{0}} = x_{i} + \mathcal{F}r \frac{t}{t_{0}} - \frac{\tilde{\lambda}}{2} \left(\frac{t}{t_{0}}\right)^{2},
+	\frac{x_{\rm f}}{L_{0}} = \frac{x_{i}}{L_0} + \mathcal{F}r \left[\frac{t}{t_{0}} - \frac{1}{\tau} \left(\frac{t}{t_{0}}\right)^{2}\right],
 \end{equation}
-where $\mathcal{F}r$ corresponds now to the non-dimensional current velocity during the slumping regime, and $\tilde{\lambda}$ to the dissipation.
+where $\mathcal{F}r$ corresponds now to the non-dimensional current velocity during the slumping regime and $\tau$ to a characteristic attenuation time.
 The fit is performed between $t/t_{0} = 5$, to avoid the initial transient phase, and $t/t_{0} = 30$, the transition time from the slumping to the inertial regime~\citep{refs}.
 Note that $x_{i}$ is kept constant equal to 0 for datasets 1 and 2, but had to be left free to adjust for the SedFoam simulations \cyril{(comment this ?)}.
-For runs with strong dissipation ($\tilde{\lambda} > 2 {\times} 10^{-2}$), we add a quartic term to \eqref{eq:fit_eq} to improve the fit quality and obtain a better estimation of $\mathcal{F}r$ and $\tilde{\lambda}$.
+For runs with strong attenuation ($\tau < 10^{2}$), we add a quartic term to \eqref{eq:fit_eq} to improve the fit quality and obtain a better estimation of $\mathcal{F}r$ and $\tau$.
 We checked that our results do not depend on the presence of this quartic term and that the linear and quadratic ones remain much larger.
 
-For dataset 3, the time resolution and the experiment duration do not allow for extracting the dissipation. Therefore, we simply fit the whole available points with a linear function.
+For dataset 3, the time resolution and the experiment duration do not allow for inferring the quadratic term. Therefore, we simply fit the whole available points with a linear function.
 
@@
 
-\subsection{Influence of the slope on the front velocity}
+\subsection{Influence of the bottom slope, $\alpha$}
 
@@
 Even if the entire set of experiments and numerical simulations are reported here, results are highlighted for $\phi < 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
+
+Describe results
+
+\medskip
+Although the slope-induced acceleration has been shown \marie{(???)} not to modify the slumping buoyancy-inertia equilibrium, the effect of the bottom slope on the corresponding dimensionless front velocity is clear.
+Following \citet{Gadal2023}, the increase of the dimensionless front velocity may instead result from the complicated current dynamics during the early transient acceleration phase, involving strong vertical motions and significant interfacial friction~\citep{Cantero2007}.
+The energetic balance between the initial state and the end of this early acceleration phase can be written as:
+\begin{equation}
+	\label{eq:energetic_balance}
+	\left[\frac{1}{2}\rho_{0}u_{\rm c}^{2} - B \Delta\rho g \sin\alpha L\right] - A \Delta\rho g \cos\alpha h_{0} =  - \frac{1}{2}C_{\rm d}\rho_{0}u_{\rm c}^{2} \frac{L}{h_{0}},
+\end{equation}
+where $A$, $B$ and $C_{\rm d}$ are scaling constants, and $L$ accounts for the alongslope distance over which the current moves during this phase.
+\citet{Cantero2007} found it to be independent of the lock aspect ratio or initial buoyancy, $L = 0.3 h_{0}$.
+Hence, we use here $L = D h_{0}$ with $D$ constant, which results in:
+\begin{equation}
+	\label{eq:Fr_th}
+	Fr = \frac{U_{\rm c}}{U_0} = \frac{Fr_0}{\sqrt{1 + C_{\rm D}}} \sqrt{1 + \frac{\tan\alpha}{S}},
+\end{equation}
+where $Fr_0$, $C_{\rm D}$ and $S$ are new constants depending on $A$, $B$, $C_{\rm d}$ and $D$.
+As shown by fig~\ref{fig:fig4}, this simple model is able to reproduce the observed variation of the dimensionless velocity in the bottom slope in the range [$5^\circ$, $35^\circ$].
+For $\alpha < 5^\circ$, [...].
+For $\alpha > 35^\circ$, [...].
 
@@
 	\caption{\textbf{Influence of the bottom slope} on the current Froude number.
-Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
+Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
+The black line indicates the prediction of \eqref{eq:Fr_th} for $Fr_{0} = 0.5$, $C_{\rm D} = 0.4$ and $S = 0.25$.}
 	\label{fig:fig4}
@@
 
-\subsection{Influence of $St$}
+\subsection{Influence of the Stokes number, $St$}
+\label{sec:influence_stokes}
 
@@
 Dotted lines mark either the average of points with $\phi < 0.45$ (a, c, e, g) or the 0 baseline (b, d, f, h).
-On (e, f), the dashed line indicates the average value for homogeneous saline currents ($St \equiv 0$) of dataset 1.}
+On (e, f), the dashed line indicates the average value for homogeneous saline currents ($St \equiv 0$) of dataset 1.
+On (b, d, f, h), the black line is $\tilde{\lambda} = 0.27 (St/a - 0.015)$.}
 	\label{fig:fig5}
@@
 In this section, we focus on the influence of settling on the current dynamics.
-The influence of the Stokes number $St$ on the dimensionless front velocity $\mathcal{F}r$ and attenuation $\tilde{\lambda}$ is presented in figure~\ref{fig:fig5} for different bottom slope $\alpha$.
+The influence of the Stokes number $St$ on the dimensionless front velocity $\mathcal{F}r$ and attenuation $\tau$ is presented in figure~\ref{fig:fig5} for different bottom slopes $\alpha$.
 
 As shown by figure~\ref{fig:fig5}(a, c, e, f), the dimensionless front velocity does not depend on $St$ at the first order.
-However, this is not the case for the attenuation parameter $\tilde{\lambda}$.
+However, this is not the case for the attenuation $1/\tau$.
 For $\alpha \approx 7^\circ$, the attenuation is constant below a threshold $St/a \sim 10^{-2}$, but increases rather linearly above this threshold (see the black line on figure~\ref{fig:fig5}(d), note the semi-log x-scale).
 
-The attenuation parameter can be linked to the transition length scale $X_{\rm t}$ as:
+The attenuation time can be linked to the transition length scale $X_{\rm t}$ as:
 \begin{equation}
 	\label{eq:link_attenuation_transition_length}
-	\frac{X_{\rm t}}{t_{\rm t}^{2}} \sim \lambda,
+	\frac{X_{\rm t}}{L_0} \sim \tau Fr,
 \end{equation}
-where $t_{\rm t} = X_{\rm t}/U_{\rm c}$ is the corresponding transition time scale.
 %
-Assuming that the current decelerates due to its entry inside the \emph{inertial} regime, we use \eqref{eq:Xt_inertial}, which leads to $\tilde{\lambda} \sim cste$.
-On the other hand, if the current decelerates due to a settling-induced loss of buoyancy, we use \eqref{eq:Xt_settling}, which results in:
+Then, assuming that the current decelerates due to its transition from the slumping regime towards the so-called \emph{inertial} regime, \eqref{eq:Xt_inertial} suggests $\tau \sim cste$.
+\textcolor{blue}{Similarly, one would obtain $\tau\sim a/Fr^{3}$ for the viscous transition.} On the other hand, if the current decelerates due to a settling-induced loss of buoyancy, then \eqref{eq:Xt_settling} results in:
 \begin{equation}
-	\tilde{\lambda} \sim \frac{St}{a}.
+	\frac{1}{\tau} \sim \frac{St}{a}.
 \end{equation}
+Hence, this explains the two different trends observed in figure~\ref{fig:fig5}(d).
+Similar results can also be extrapolated for $\alpha = 0^\circ$ and $\alpha = 15^\circ$, although additional points are required to assess the dependence of the transitional threshold and of the slope of the linear increase with $\phi$ or $\alpha$.
+
+For larger slopes, our experiments do not allow us to discriminate the influence of the Stokes number (see figure~\ref{fig:fig5}(g,h)).
+\cyril{Further exploring these regimes might arise singular effects as, at large slopes, a significant component of the settling occurs along the slope, and might therefore increase the current inertia, or at least decrease the expected attenuation.}
 
@@
 	\caption{\textbf{Influence of particle volume fraction.} Current Froude number as a function of the initial particle volume fraction for (a) $\alpha \approx 0^\circ$, and (b) $\alpha \approx 45^\circ$.
-The inset in (a) shows the plot in linear scale.}
+The inset in (a) shows the plot in linear scale.
+The black line indicates the prediction of \eqref{eq:Fr_th_eta} for $Fr_{0} = 0.5$, $C_{\rm D} = 0.4$, $S = 0.25$ and $Re_c = 500$.}
 	\label{fig:fig6}
 \end{figure}
-The influence of the parameter $\Phi$ on the Froude number is presented in Figure~\ref{fig:fig6}.
+
+\jean{The influence of the parameter $\Phi$ on the Froude number is presented in Figure~\ref{fig:fig6}.
 The results demonstrate the existence of a critical fraction $\Phi_{c}$ above which the Froude number decreases.
 This reduction is characterized by an increase particle-particle and particle-fluid interactions which leads in wall friction and a decrease in inertial forces, resulting in a corresponding increase in viscous forces.
-This results in the decrease of the front velocity and the collapse of the current with the presence of a strong mass fraction gradient near the front.
+This results in the decrease of the front velocity and the collapse of the current with the presence of a strong mass fraction gradient near the front.}
+
+The energetic balance presented in section~\ref{sec:influence_stokes} can be adapted to include viscous dissipation effects by replacing the constant $C_{\rm d}$ as:
+\begin{equation}
+	C_{\rm d} \equiv  C_{\rm d}\left(1 + \frac{E}{Re}\frac{U_0}{U_{\rm c}}\frac{\eta_{\rm eff}}{\eta_{\rm f}} \right),
+\end{equation}
+where $E$ is a new scaling constant, and $\eta_{\rm eff}$ is the effective suspension viscosity taking into account the presence of particles. This results in:
+\begin{equation}
+	\label{eq:Fr_th_eta}
+	\begin{split}
+		Fr & =  \frac{1}{1 + C_{\rm D}} \left[-a\frac{Re_{\rm c}}{Re}\frac{\eta_{\rm eff}}{\eta_{\rm f}} \right. \\
+		&\left. +\sqrt{\left(\frac{Re_{\rm c}}{Re}\frac{\eta_{\rm eff}}{\eta_{\rm f}} \right)^{2} + Fr_0^{2} \left(1 + \left[1 + C_{\rm D}\right]\frac{\tan\alpha}{S} \right)}  \right],
+	\end{split}
+\end{equation}
+where $Re_{\rm c}$ is new constant. The effect of the volume fraction on the effective viscosity is taken into account following \citet{krieger1959mechanism}:
+\begin{equation}
+	\label{eq:Krieger}
+	\frac{\eta_{\rm eff}}{\eta_{\rm f}} = \left(1 - \frac{\phi}{\phi_{\rm m}}\right)^{-(5/2)\phi_{\rm m}},
+\end{equation}
+where $\phi_{\rm m} = 0.585$ is the maximum volume fraction~\citep{boyer2011unifying}.
+
+As shown by figure~\ref{fig:fig6}(a), the combination of \eqref{eq:Fr_th_eta} and \eqref{eq:Krieger} is able to reproduce quantitatively the observed influence of $\phi$ of the dimensionless velocity for $\alpha=0^\circ$.
+Note that we checked that this agreement does not depend on the choice of the viscosity model by trying other available in the literature~\citep{stickel2005fluid, boyer2011unifying}.
+
+\section{Conclusion and perspectives}
+
+\julien{
+	Influence of kinetic theory closures suggests that the particle-particle interactions play a role in the dynamics of the PLGC.\\
+	3D effects matter on the turbulent mixing at the current interface (add a figure with 2 snapshots)\\
+	Influence of the initial condition in the numerical simulations: noise on volume fraction, velocities, ... Relevant for 2D RANS simulations ?
+}
+
+\subsection*{Data availability}
+
+[...]
 
 \subsection*{Acknowledgment}
-Sylvain Dauge for the full designing of Setup 2. Hervé Bellot and Brivaël Collin are thanked for experimental support.
+\marie{Sylvain Dauge is acknowledged for the full designing of Setup 2. Hervé Bellot and Brivaël Collin are thanked for experimental support.}
+
+JC and CB would like to acknowledge Hugo Divel for his contribution on the preliminary lock-release simulations using SedFoam.
 

### 2023-06-01 figure_scripts/figure1.py
Adding Sedofoam snapshots to fig1

@@ -11,8 +11,8 @@ from matplotlib.path import Path
 
 plt.rcParams['figure.constrained_layout.hspace'] = 0
 plt.rcParams['figure.constrained_layout.wspace'] = 0
-plt.rcParams['figure.constrained_layout.h_pad'] = 0
-plt.rcParams['figure.constrained_layout.w_pad'] = 0
+plt.rcParams['figure.constrained_layout.h_pad'] = 0.005
+plt.rcParams['figure.constrained_layout.w_pad'] = 0.005
 
 
 def sind(x):
@@ -57,16 +57,15 @@ water_height_facts = [0.89, 1, 0.89]  # proportion of tank height
 width_ratios = tank_length*cosd(SLOPES) + np.abs(tank_height*sind(SLOPES))
 
 
-figsize = (tp.large_figure_width, 0.73*tp.large_figure_width)
-fig = plt.figure(constrained_layout=True, figsize=figsize)
+figsize = (tp.large_figure_width, 0.84*tp.large_figure_width)
+fig = plt.figure(figsize=figsize, layout='compressed')
 
-subfigs = fig.subfigures(2, 1, height_ratios=(1, 1.32), hspace=0.005)
-axs_sketches = subfigs[0].subplots(
-    1, 3, width_ratios=width_ratios, gridspec_kw={'wspace': 0.02})
-subfigures_images = subfigs[1].subfigures(
-    1, 3, width_ratios=width_ratios)
+subfigs = fig.subfigures(2, 1, height_ratios=(0.364, 0.636), hspace=0.005)
 
 # %%%%%%% Sketches
+
+axs_sketches = subfigs[0].subplots(
+    1, 3, width_ratios=width_ratios, gridspec_kw={'wspace': 0.02})
 
 ax_labels = ['a', 'b', 'c']
 for i, (ax, ax_label, slope, water_height_fact) in enumerate(zip(axs_sketches.flatten(),
@@ -215,14 +214,19 @@ for i, (ax, ax_label, slope, water_height_fact) in enumerate(zip(axs_sketches.fl
             r'$L_{0}$', ha='center', va='top')
 
 # %%%%%%% experimental images
+subfigures_images = subfigs[1].subfigures(
+    2, 1, height_ratios=(0.77, 0.20), hspace=0.022)
+
+subfigures_images_top = subfigures_images[0].subfigures(
[…]

### 2023-06-01 figure_scripts/template.py

@@ -41,6 +41,7 @@ rcParams['figure.figsize'] = large_figure_size
 
 # cmap_images = cmo.cm.ice
 cmap_images = cmo.cm.gray
+cmap_images2 = cmo.cm.gray_r
 # cmap_images = cmo.cm.dense_r
 # cmap_images = cmr.arctic
 

### 2023-07-21 paper/main.tex
SUBMISSION

--- 2023-05-04 paper/main.tex
@@
 \usepackage{textcomp}
-\usepackage{xcolor}
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage{amsmath}
-
-% \usepackage{moreverb,savetrees}
+\usepackage{cuted}
 
@@
 
-\graphicspath{{./figures/}{./}}
+\newcommand*{\cyril}{\textcolor{ForestGreen}}
+\newcommand*{\marie}{\textcolor{Cyan}}
+\newcommand*{\julien}{\textcolor{Brown}}
+\newcommand*{\yvan}{\textcolor{Blue}}
 
-\title{Title}
-\author[1]{PALAGRAM}
+\definecolor{jeanc}{RGB}{253,129,43}
+\newcommand*{\jean}{\textcolor{jeanc}}
 
-\affil[1]{Affiliations}
+\graphicspath{{./figures}{./}}
 
-\date{}
+\title{Chapter 16: Particle-laden gravity currents: the lock-release slumping regime at the laboratory scale}
 
-\newenvironment{Figure}
-  {\par\medskip\noindent\minipage{\linewidth}}
-  {\endminipage\par\medskip}
+% \title{Chapter 16: Lock-release particle-laden gravity currents at the laboratory scale}
+
+% \title{Macroscopic description of particle-laden gravity currents: overall perspectives}
+
+% \title{Macroscopic description of lock-release particle-laden gravity currents}
+
+\author[1]{Gadal C.}
+\author[2]{Schneider J.}
+\author[3]{Bonamy C.}
+\author[3]{Chauchat J.}
+\author[2]{Dossmann Y.}
+\author[2]{Kiesgen de Richter S.}
+\author[1]{Mercier M. J.}
+\author[4]{Naaim-Bouvet F.}
+\author[3]{Rastello M.}
+\author[1]{Lacaze L.}
+
+\affil[1]{Institut de Mécanique des Fluides de Toulouse (IMFT), Université de Toulouse, CNRS, Toulouse, France}
+\affil[2]{Laboratoire Énergies et Mécanique Théorique
+et Appliquée (LEMTA), Université de Lorraine, CNRS, 54500, Nancy, France}
+\affil[3]{Univ. Grenoble Alpes, CNRS, Grenoble INP, LEGI, 38000 Grenoble, France}
+\affil[4]{Univ. Grenoble Alpes, INRAE, CNRS, IRD, Grenoble INP, IGE, 38000 Grenoble, France}
+
+\date{{\it\color{red}Deadline: May 31$^{st}$ - repoussé (mi juillet ou EuroMech?)}}
 
@@
 	\begin{abstract}
-		Particle-laden gravity currents (PLGCs) are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid.
-We focus our attention on finite volume current, generated here by lock-release device over a controlled constant slope bottom at the laboratory scale.
-The main objective of the present chapter is to provide a comprehensive flow regime associated with the relevant dimensionless parameters which characterize the current at the macroscopic scale, i.e.
-average over the current depth.
+
+		Particle-laden gravity currents (PLGCs) are driven by the mass difference between a heavy fluid-particle mixture and a lighter ambient liquid.
+They often occur in natural and industrial situations, among which a typical situation is the release of a finite volume.
+Here, we focus on such `dam-break' situations, which are studied using lock-release devices at the laboratory scale.
+		%
+		The objective of the present chapter is to provide a description at the macroscopic scale of the early moments of the flow, namely the \emph{slumping regime}, with respect to the relevant dimensionless parameters.
+For this, we combine a total of 288 runs from three different lock-release devices and from two-fluids numerical simulations, which allow us to cover a large range of particle types (size and density), volume fractions, bottom slopes and geometries.
+		%
+		By tracking the front propagation through time, we extract the dimensionless slumping velocity $\mathcal{F}r$ and dimensionless characteristic slumping duration $\tau$, on which we base our description.
+		%
+		Our results show that the slumping velocity increases with the bottom slope, but decreases with the particle volume fraction when the latter exceeds a critical value.
+However, it remains independent of particle settling processes, which on the other hand affects the slumping duration.
+Hence, above a critical threshold, $\tau$ decreases as the ratio between the settling velocity and characteristic current velocity increases.
+For all these regimes, we derive scalings and energetic balances that reproduce the observed trends.
+The latter comparison confirms the role of initial energy transfer from the initial state towards the slumping phase on the resulting dynamics.
+This initial process and its characterisation remain crucial to prescribe relevant initial conditions for large-scale predictive modelling.
+
 	\end{abstract}
-	\vskip 0.7cm
+
+	\vspace{0.3cm}
+	{\it \color{red}REMINDER: Word limit of about 10,000 per article, plus 7 figures.}\\
+	\vspace{0.3cm}
 ]
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 1
-\begin{figure*}[h]
+\section{Introduction}
+\label{sec:intro}
+
+\subsection{Finite volume density currents}
+\label{sec:intro_lockrelease}
+
+Density currents have been widely considered at the laboratory scale for their ubiquitous occurrence in geophysical flows \citep{Hopfinger1983,Simpson1999,Dufek2016}.
+As they are mostly associated with unsteady and non-uniform dynamics in nature, the lock-release device appeared to be a popular canonical configuration to mimic their behaviour at the laboratory scale \citep{Hacker1996,Shin2004,Nogueira2014}.
+In particular, studies often focused on characterising the front dynamics and the front shape as a macroscopic description of the underlying complex dynamics \citep{Rottman1983,Marino2005,Hogg2006}.
+
+The basic approach to describe a lock-release evolution is built upon characteristics solutions of the hyperbolic shallow water equations for an inviscid dense liquid of density $\rho_{\text c}$ slumping on a horizontal bottom under an inertialess and weightless ambient fluid.
+In particular, this allows predicting the front velocity $U_{\text{c}}$ only based on the initial condition of the released mass.
+At the early stages of the slumping phase, it is found to be simply related to the only velocity scale of the system as
+\begin{equation}
+	U_{\text{c}}\propto \sqrt{gH_0},
+\end{equation}
+with $g$ gravitational acceleration and $H_0$ the initial height of the fluid volume being released.
+The method of characteristics suggests a coefficient of proportionality of exactly $2$ between velocities $U_{\text{c}}$ and $\sqrt{g H_0}$.
+An extension towards a weighted ambient fluid of density $\rho_{\text a}$ could lead to
+\begin{equation}
+	\label{eq:solUc}
+	U_{\text{c}}=F\sqrt{\tilde{g}H_0},\quad\mbox{with}\quad \tilde{g}=\frac{\rho_{\text c}-\rho_{\text a}}{\rho_{\text c}}g,
+\end{equation}
+However, the coefficient of proportionality $F$, similar to a Froude number, now requires slightly more attention as it is clearly found smaller than $2$ in experiments or simulations.
+The main reason is that most density currents are characterized by a small density ratio between current and ambient, making the assumption of inertialess ambient fluid unlikely.
+A dynamic front condition is now required, and is often inspired from the steady front configuration derived by \citet{Benjamin1968}, now also linking the ambient height $H_{\text a}$ to the front velocity.
+This front relation includes two unknowns, a front velocity and a front height, and is often used as a front boundary condition in shallow water numerical solver for gravity currents \cite[e.g.][]{Ungarish2009}.
+Then, past studies managed to link the front velocity to the current height close to the front.
+However, the link between the front velocity and the initial conditions then becomes much more difficult to predict.
+Models have been proposed to circumvent this issue, leading for instance to \citep{Huppert1980}
+\begin{equation}
+	\label{eq:Froude_Huppert1980}
+	F = 0.5 \sqrt{\frac{\rho_{\text c}}{\rho_{\text a}}}\left(\frac{H_{\text a}}{H_0}\right)^{1/3}.
+\end{equation}
+%
+The latter is relevant for not too deep ambient $H_{\text a}$ compared to the reservoir height $H_0$, which is actually the case in most experiments.
+Many configurations correspond to full-depth release ($H_{\text a}=H_0$), which leads to $F = 0.5(\rho_{\text c}/\rho_{\text a})^{1/2}$.
+Even if several experiments and numerical simulations confirm such a result, the exact value of $F$ remains still unclear and probably dependent on several physical processes.
+In particular, these relationships during the slumping regime are based on assuming negligible dissipation, which shall be violated in some cases, for instance in the presence of settling particles or significant mixing with the ambient at the miscible interface.
+
+For a finite volume lock-release device on a mild enough slope, the slumping regime necessarily ends when the rarefaction wave associated with finite volume reaches the current front \citep[e.g.][]{Rottman1983, Hogg2006}.
+This leads to a new regime usually referred to as the inertial regime in the literature.
+This is observed on the front evolution as $U_{\text{c}}$ is no longer time independent but evolves as $t^{-1/3}$.
+A simple approach can be useful to provide a rapid estimation of the position $X_{\text t}$ or time $T_{\text t}$ at which this transition occurs.
+For that purpose, it is worth mentioning that such a transition occurs as a rarefaction wave evolves faster than the front.
+Then, assuming the front velocity $U_{\text{c}}$ to be different from the rarefaction wave velocity at the interface $\sqrt{\tilde{g}H_0}$, which is assumed to be constant along its entire propagation path for simplicity, this leads to a transition length scale $X_{\text t}$ written as
+\begin{equation}
+	\label{eq:Xt_inertial}
+	\displaystyle X_{\text t}\propto  \frac{2L_0 F}{1 - F}.
+\end{equation}
+Note that here $2$ comes from the path of propagation of the wave: emanating from the gate, it propagates backwards to the end of the reservoir, reflects on the wall and then propagates towards the front.
+During the slumping regime, the previous estimation of $F$ \eqref{eq:Froude_Huppert1980} for full depth release $H_0=H_{\text a}$ leads to $X_{\text t}\propto 2L_0/[2(\rho_{\text a}/\rho_{\text c})^{1/2}-1]$.
+Note that, although more advanced methods based on single-layer shallow-water models predict a similar result, $X_{\text t} = 2 L_0 F (1 + F/2)^{1/2}$ \citep{Hogg2006}, full depth-release experiments have found much larger transitional positions, i.e $X_{\text t} \approx 10 L_{0}$~\citep{Rottman1983}.
+
+Other mechanisms can modify the dynamics of the slumping regime which could lead to other transitions of the front velocity at different positions $X_{\text t}$ or times $T_{\text t}$, prior to the influence of the rarefaction wave.
+This is basically part of the topic of the present chapter.
+
+For a gravity current on a horizontal bottom, a classic transition discussed in the literature is the one dominated by viscous dissipation.
+The flow eventually reaches a fully viscous regime \citep{Huppert1980}.
+This transition typically holds for fully laminar flow.
+A similar regime can be induced by inertial friction, for which the full derivation remains less obvious \citep[see for instance][for the case of a constant-influx current]{Hogg2001}.
+In any case, such transition can be seen as the time or position at which the current is strongly affected by dissipation, which actually encompasses entrainment for miscible gravity currents.
+The slumping regime then ends when dissipation induced by bottom friction and/or entrainment at the miscible interface, overcomes the current dynamics.
+%
+Again using a simple idealised approach, the latter transition can be estimated from a dynamic condition balancing an estimation of the shear stress $\rho_{\text c} C_d U_{\text{c}}^2$ ($C_d$ being some drag coefficient induced by friction at the bottom and/or by entrainment at the interface) with inertia over the current height $h_{\text c}$, $\rho_{\text c} U_{\text{c}}^2 h_{\text c} /X_t$.
+In the latter expression, $h_{\text c}$ can have different scaling.
+In the case of a clearly detached current head, the constant height scales with the initial height $H_0$.
+When assuming that the current remains attached to the initial column at transition \citep[similarly to][but for inertial drag instead of viscous dissipation]{Bougouin2022}, mass conservation between the initial column and a simplified box geometry implies $h_{\text c} = H_0 L_0/X_t$.
+%
+Both cases respectively lead to
+\begin{equation}
+	\label{eq:Xt_drag}
+	\displaystyle X_{\text t}= {\frac{H_0}{C_d}}\quad \mbox{or} \quad X_{\text t}=  \sqrt{\frac{H_0 L_0}{C_d}}.
+\end{equation}
+
+To finish with, constant bottom slope $\alpha$, often relevant to characterize the main process of natural flow over topography, adds complexity to this front analysis.
+First, note that, when considering an inclined bottom, the reduced gravity and therefore the characteristic velocity \eqref{eq:solUc} shall be updated with a $\cos{\alpha}$ term accounting for the reduced effect of the initial column pressure normal to the inclined bottom.
+Second, including a bottom slope also adds a constant forcing along the slope suggesting a constant acceleration of the flow, at least until balanced by friction at both the interface and the bottom.
+Hence, it could affect the existence of the aforementioned slumping constant-velocity regime.
+%
+Yet, constant velocities associated with a slumping regime are clearly observed at early stages in experiments and simulations prior to such balance \citep{Birman2007, Gadal2023}.
+This is due to the relatively long time or length scales required for slope-induced acceleration to overcome the flow induced by the initial pressure gradient.
+A way to highlight that is a simple balance between slope and pressure gradient, resulting in
+\begin{equation}
+	\label{eq:Xt_slope}
+	X_{\text t}=H_0\tan^{-1}{(\alpha)}.
+\end{equation}
+The latter can be seen as a transition length scale after which the slope starts driving the current until reaching the next equilibrium between slope and friction.
+
+%%%%%%Figure 1
+\begin{figure*}[ht!]
 	\centering
@@
 	\caption{\textbf{Experimental set-ups used in this study.} (a--c) Sketches of the three experimental set-ups.
-Below are shown snapshots of corresponding experimental images.
-(d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, $\mathcal{S}t=$ (i--k) Set-up 2, $\alpha = $, $\phi \sim $, $\mathcal{S}t=$ (l--p) Set-up 3, $\alpha = $, $\phi \sim $, $\mathcal{S}t=$.}
+Below are shown snapshots of corresponding experimental images: (d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, ${S}t=0.027$; (i--k) Set-up 2, $\alpha = 45.4^\circ$, $\phi \sim 1.5~\%$, ${S}t=0.06 $; (l--p) Set-up 3, $\alpha = 0^\circ$, $\phi \sim 55\%$, ${S}t \equiv 0$; (q--v) SedFoam, $\alpha = 45.4^\circ$, $\phi \sim 1.7~\%$, $\mathcal{S}t=0.06$.}
 	\label{fig:fig1}
 \end{figure*}
+%%%%%%%%
 
-\section{Introduction}
+\subsection{Particle-Laden Gravity Currents}
+\label{sec:intro_PLGC}
 
-\subsection{Control parameters}
+Particle-Laden Gravity Currents (PLGCs) refer to as gravity currents, or density currents, are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid \citep[e.g.][]{Hopfinger1983,Middleton1993,Kneller2000,Meiburg2010,wells2021}.
+Different subclasses of PLGCs can be distinguished depending on the relative density of the different phases constituting the system; i.e.
+the particle phase, the carrier fluid phase and the ambient fluid phase.
+In the following, we will focus on two of them, characterized by only two distinct densities.
+At one end, the carrier phase and ambient phase are the same fluid and particles are heavier.
+This is the most commonly studied subclass of PLGCs, usually referred to as turbidity currents.
+At the other end, PLGCs can also be induced by a neutrally-buoyant fluid-particle mixture flowing into a lighter ambient fluid.
 
-Particle-laden gravity currents (PLGCs) are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid.
-A major difference between compositional currents and PLGCs is therefore the existence of typical length and time scales associated with the particle dynamics.
-Particularly, the own settling velocity of a heavy particle only allows the existence of these gravity currents on finite time and length scales.
-Nevertheless, prediction of these finite scales is difficult as it shall strongly depends on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface, which modifies entrainment and associated dissipation, and in turn can delay or enhance settling.
-Accordingly, their dynamics are controlled by several dimensionless parameters leading to a wide variety of flow regimes.
-In particular, beyond the slope $\alpha$, the dimensionless density difference between the current and the ambient $At$ and the ratio between available inertia and dissipation, say a Reynolds number $Re$, as usually introduced as a control parameter for compositional currents, the initial solid fraction of particles $\phi$ and a dimensionless settling velocity St have now to be considered as potentially affecting flow regimes.
-It shall be noted that in some specific situations, some of these dimensionless parameters could be linked as, for instance, $\phi$ and $At$ for heavy particles in a single-phase liquid for the carrying and ambient fluids
+A major difference between PLGCs and compositional currents is therefore the existence of typical length and time scales associated with the particle dynamics.
+Nevertheless, the prediction of these finite scales is difficult as it shall strongly depend on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface which modifies entrainment and associated dissipation.
+In turn, this can modify the macroscopic behaviour of the current, for instance, the dynamics of the front.
+Accordingly, these dynamics shall be controlled by several dimensionless parameters that can lead to a wide variety of flow regimes.
+In particular, beyond the slope $\alpha$, the dimensionless density difference between the current and the ambient $\mathcal{A}t$, the ratio between available inertia and dissipation, say a Reynolds number $\mathcal{R}e$ as usually introduced as a control parameter for any density currents, the initial solid fraction of particles $\phi$ and a dimensionless settling velocity $S=V_{\text s}/U_{\text{c}}$ with $V_{\text s}$ the settling velocity of particles into the carrying fluid phase, have now to be considered as potentially affecting flow regimes.
+It shall be noted that in some specific situations, some of these dimensionless parameters could be linked (for instance $\phi$ and $\mathcal{A}t$ for heavy particles in a single-phase liquid for the carrying and ambient fluids).
 
-\subsection{Literature}
+Beyond the transition from the slumping regime to other regimes discussed in the previous section (still relevant for PLGCs although parametrisation of the processes involves more complex mechanisms induced by the fluid-particle interaction), here, a new transition towards the deposition of the particle can emerge.
+In particular, based on a kinematic approach balancing time scales related to current propagation and particle settling, leads to
+%
+\begin{equation}
+	\label{eq:Xt_settling}
+	X_{\text t} \propto H_0\frac{U_{\text{c}}}{V_{\text s}}.
+\end{equation}
+%
+Note that such transition shall disappear for the neutrally-buoyant situation of PLGCs.
+However, it has been observed that mixing at the interface can modify the local concentration of particles at the front, resulting in behaviour similar to that of settling~\cite{Schneider2023}.
+Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear.
+Overall, the value of $F$ in \eqref{eq:solUc} during the slumping inertial regime remains to clarify varying the influence of the mechanisms discussed so far.
 
-\begin{itemize}
-	\item literature slope
-	\item literature settling
-	\item literature volume fraction
-	\item other effects: lock geometry, non-boussinesq, entrainment, erosion/deposition $\rightarrow$ ignored here
-\end{itemize}
+Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed in the front dynamics, often referred to as the Froude number, and the front shape.
+Then prediction of the different flow regimes associated with PLGCs shall be somewhat included in these observables.
+After presenting the experimental and numerical methods used to investigate PLGCs dynamics in section~\ref{sec:methods}, we introduce the main dimensionless quantities used and macroscopic observations in section~\ref{sec:current_description}.
+We discuss our results in section~\ref{sec:results} and conclude in section~\ref{sec:conclusion}.
 
-\section{Dimensional analysis at the macroscopic scale of the current}
+\section{Methods}
+\label{sec:methods}
+
+In order to characterize the slumping regime, and its transition to other regimes, of PLGCs induced by the finite-volume release of a suspension initially at ``rest'', 3 lock-release devices have been used to cover a wide range of control parameters.
+These experimental devices are supplemented with two-fluid numerical simulations to confirm the relevance of such a flow description to characterize the dynamics of PLGCs.
+
+\subsection{Experimental set-ups}
+
+All set-ups are based on the same typical lock-release design (see figure~\ref{fig:fig1}).
+The tank has a width $W$, a height $H$ and a length $L_0 + L_1$ along the $x$ direction.
+It is split into two parts by a sluice gate at a distance $L_0$ from the left wall.
+%More specifically, a tank of width $W$, height $H$ and length $L_0+L_1$ along the streamwise direction $x$ is split in two along this direction by a sluice gate at a distance $L_0$ from one end.
+On the $L_0$ side, a reservoir is filled up to a height $H_0$ with a fluid/particle mixture having a density $\rho_{\text{m}} = \phi \rho_{\text{p}} + (1 - \phi)\rho_{\text{f}}$, with $\rho_{\text{p}}$ and $\rho_{\text{f}}$ the particle and fluid densities, respectively.
+On the other side of height $H_{\text a}$ is the ambient fluid of density $\rho_{\text{a}}$, with typically $\rho_{\text{a}} < \rho_{\text{m}}$.
+At $t = 0$, the sluice gate is lifted up manually to initiate the gravity current.
+
+\paragraph{Set-up 1}
+
+The first set-up (figure~\ref{fig:fig1}(a)) has $L_0+L_1=150~{\text cm}$ with $L_0 = 10~\textrm{cm}$, $W = 10~\textrm{cm}$ and $H=49~\textrm{cm}$.
+The upper boundary is a free surface and the tank can be inclined at an angle $\alpha\in[0^\circ,7.5^\circ]$ with respect to the horizontal as shown in the 2D sketch of figure~\ref{fig:fig1}(a).
+For inclined configurations, $H_{\text a}$ then increases along the flume, and the reservoir is filled such as both heights at the gate, reservoir one and ambient one, are the same (see figure~\ref{fig:fig1}(a)).
+The suspension in the reservoir is made by strongly stirring a known mass of particles within the whole water column using a mechanical mixer, which is stopped just before opening the gate.
+$H_{0}$ is defined as the average suspension height inside the lock, equal to $20~\textrm{cm}$.
+%
+The current evolution is followed by an sCMOS pco.edge$^\copyright$ camera at 120 Hz, using a LED backlight as a light source.
+Additional details on the set-up and the image processing can be found in \citet{Gadal2023}.
+
+\paragraph{Set-up 2}
+The second set-up (figure~\ref{fig:fig1}(b)) has $L_0+L_1=3.8~{\text m}$ with $L_0 = 30~\textrm{cm}$ for runs with PMMA particles and $L_0 = 10~\textrm{cm}$ for runs with silica sand or hydrogel particles (see section~\ref{sec:datasets} for particles specifications).
+The overall flume width and depth are $W= 10~\textrm{cm}$ and $H= 50~\textrm{cm}$.
+The upper boundary in the flume is a rigid lid, and the only free surface is in the lock region.
+The ambient fluid is tap water at room temperature and $H_{\text a}=50~\textrm{cm}$.
+The flume can be inclined at an angle $\alpha \in [0^\circ,45.4^\circ]$ as shown in figure~\ref{fig:fig1}(b).
+Here, $H_0$ is defined as the average suspension depth perpendicular to the flume floor inside the lock at the gate opening time.
+LED plates illuminate the flume from above.
+The flow dynamic is recorded with a Phantom$^\copyright$ Miro C110 camera at 50 Hz.
+
+\paragraph{Set-up 3}
+
+The third set-up (figure~\ref{fig:fig1}(c)) has $L_0+L_1=150~{\text cm}$ with $L_0 = 20~\textrm{cm}$, $W= 10~\textrm{cm}$ and $H = 10~\textrm{cm}$.
+The upper boundary is a free surface and the tank remains horizontal ($\alpha = 0^\circ$), as shown in the 2D sketch of figure~\ref{fig:fig1}(c).
+Here, the suspension and ambient heights are equal, i.e.
+$H_{a}=H_{0}$, and vary between 20~cm and 30~cm depending on experiments.
+The current evolution is followed by a Nikon$^\copyright$ D500 camera at 10 Hz and using a white LED panel as a light source.
+Additional details on the set-up and the image processing can be found in \citet{Schneider2023}.
+
+\paragraph{Two-fluid model}
+
+A two-fluid model, SedFoam \citep{chauchat2017}, is used to simulate lock-release configurations from set-ups 1 and 2.
+This model available for download at Zenodo \citep{bonamy2023} is implemented in the open-source computational fluid dynamics toolbox OpenFOAM and has been extensively validated for various sediment transport problems such as sheet-flow \citep{cheng2016,chauchat2022,mathieu2021,mathieu2022}, scour \citep{nagel2020,tsai2022}, ripples \citep{salimi2021a,salimi2021b}, avalanches \citep{montella2021} and granular collapse \citep{montella2023}.
+The simulations presented herein are 2D vertical and turbulence-averaged.
+The turbulence model is a $k-\epsilon$ model with additional two-phase flow terms for density stratification and turbulent drag dissipation.
+The closure for the particle-particle interactions is a frictional-collisional model based on a Coulomb rheology for the frictional part and the kinetic theory of granular flows for the collisional part.
+The two contributions are simply added as described in \cite{chauchat2017}.
+The drag force is modelled using the \cite{schiller1933} closure multiplied by a hindrance function of \cite{richardson1954}.
+In order to describe the influence of the slope, the total pressure is solved and the gravity vector is rotated while the mesh is kept aligned with the $x$ and $y$ axis.
+
+\subsection{Datasets}
+\label{sec:datasets}
+
+269 experiments have been performed on the 3 experimental set-ups described in the previous section.
+They are complemented by 19 numerical simulations using SedFoam.
+Altogether, they are classified into 4 different datasets corresponding to the 3 experimental set-ups and the numerical modelling respectively.
+These 4 datasets are described below.
+
+\paragraph{Dataset 1}
+
+The first dataset is gathered using set-up 1.
+A first series of experiments is carried out with silica sand grains with $\rho_p=2650\textrm{kg/m}^{3}$ and $d \sim 120~\mu\textrm{m}$ while varying the bottom slope from $0^\circ$ to $7^\circ$.
+Moreover, for $\alpha = 7^\circ$, the settling velocity is varied by using glass beads (Silibeads$^\copyright$) with $\rho_p=2500\textrm{kg/m}^{3}$ and mean diameter ranging $d$ from $60~\mu\textrm{m}$ to $250~\mu\textrm{m}$, corresponding to $V_{\text s} \in [0.3, 3.2]~\textrm{cm}~\textrm{s}^{-1}$.
+In most series of experiments, the initial volume fraction $\phi$  is systematically varied in the range $[0.003, 0.15]$.
+%
+A few experiments with saline water (homogeneous gravity currents) performed at $\alpha=7^\circ$ are also included in the dataset, however not discussed in this chapter.
+
+\paragraph{Dataset 2}
+
+The second dataset is gathered using set-up 2.
+Three different types of particles have been used in order to vary $d$, $\rho_{\text p}$ and as a direct consequence $V_{\text s}$ and $\mathcal{S}t$.
+A first series of experiments is carried out with the same silica sand grains as in dataset 1, with the same stirring procedure and varying the bottom slope from $7^\circ$ to $15^\circ$ and $\phi$ from 0.01 to 0.1.
+%
+A second series of experiments is carried out with PMMA particles with $\rho_{\text p}=1190~\textrm{kg/m}^{3}$ and diameter $d\sim 250~\mu\textrm{m}$.
+Part of the runs had a manual stirring with a rod and $H_0=H_{\text a}$, while others deal with a release of a deposited layer of particles and no mixing leading to $H_0<H_{\text a}$.
+Here, the bottom slope is varied between $0^\circ$ and $45.4^\circ$, and the initial volume fraction is kept nearly constant, i.e.
+$\phi \in [0.01, 0.02]$.
+%
+A third series of experiments is carried out with hydrogel particles of densities $\rho_{\text p}\in \{1003, 1045\}~\textrm{kg/m}^{3}$ and $d\in \{0.3, 1.6\}~\textrm{cm}$.
+For most of them, the flume was inclined to $45.4^\circ$ and experiments began with particles deposited inside the lock (no stirring, $H_0<H_{\text a}$).
+In these cases, hydrogel physical properties, among which deformability, make volume fraction within the deposited layer reach values up to nearly 1 (full packing), way above the classical 0.65 for solid spherical monodisperse granular materials.
+These initially deposited runs are highlighted in each figure by red-edged symbols.
+Additionally, a couple of hydrogel runs have been made at a lower slope angle and were released from a stirred configuration.
+
+\paragraph{Dataset 3}
+
+The third dataset is gathered using setup 3.
+The experiments are performed using a neutrally-buoyant fluid-particle mixture made of polystyrene beads of typical diameter $d \sim 1~\textrm{mm}$ with density $\rho_{\text p} \sim 1038.5~\textrm{kg/m}^{3}$.
+Saltwater is added to the lock for density matching.
+The density difference between the lock and the ambient fluid is kept nearly constant for each experiment, equal to $25 \textrm{kg}~\textrm{m}^{-3}$.
+Only the initial volume fraction is varied in the range $[0.03$  $0.55]$.
+
+\paragraph{Dataset 4}
+
+This dataset has been obtained using the two-fluid model described above.
+The domain size is $L_0+L_1=3.8$~m long and $W=0.5$~m height and the typical grid resolution is $\Delta x=1.25$~mm in the streamwise direction and it varies between $\Delta y \in [0.25 ; 3.5]$ mm in the wall-normal direction with a geometrical progression.
+The time step is adaptative following criteria on the Courant number, $C_{FL}=\max(U) \Delta t / \Delta x < 0.3$ for the phase velocities as well as for the relative velocity between the phases (the maximum allowed time-step during the simulation is set to $\max(\Delta t) = 10^{-3}$~s).
+The lock is $L_0=0.3$~m wide and $H_0=0.5$~m high, corresponding to the domain dimensions.
+The numerical schemes are second order in time and space: backward for time, a blend of centred and upwind schemes for advection (limitedLinear in OpenFOAM) and a centred scheme for diffusion.
+The initial condition is fluid and particle phase at rest ($u_f=u_p=0$) and the initial volume fraction of particles $\phi$ is set to a constant value in the region $x\in[-0.5 ; 0]$ m and $y\in[0;0.5]m$.
+At $t=0$, the governing equations are numerically integrated until $t_{end}=80$s maximum.
+The front position is determined as the most downstream position of the isocontour of volume fraction $\phi =10^{-3}$ at a frequency of 10Hz.
+This criterion has been demonstrated to be robust for all the simulations presented in this work.
+The particle properties are $e=0.8$ (restitution coefficient for binary collisions) and $\mu_{\text s}=0.53$ (macroscopic friction coefficient).
+Two sets of simulations have been performed.
+One mimics silica sand beads in water $\rho_{\text p}/\rho_{\text f}=2.65$ with diameter $d_p=60~\mu$m and initial volume fractions $\phi\in[0.015, 0.1]$.
+The other one mimics PMMA particles in water $\rho_{\text p}/\rho_{\text f}=1.19$ with diameters $d \in [100, 650]~\mu$m and initial volume fractions $\phi \in [0.0144, 0.5]$.
+For both configurations, the slope angle $\alpha$ has been varied in the range $\alpha \in [0^\circ, 45^\circ]$.
+
+\section{Macroscopic current description}
+\label{sec:current_description}
+\subsection{Dimensional analysis at the macroscopic scale of the current}
 \label{sec:dimensionlessmap}
 
-Here, the dimensionless parameters are based on a characteristic length $L_0$ and a characteristic velocity $U_0=\sqrt{g' H_0}$ where $L_0$ and $H_0$ are the horizontal length and depth of the initial container, respectively, $g'$ is the modified gravity $r g \cos{\alpha}$ and $r = (\rho_m - \rho_a)/\rho_a$ with $\rho_m$ and $\rho_a$ the current density and ambient fluid density respectively.
-Note that $\rho_m = \phi \rho_p+(1-\phi)\rho_f$ with $\rho_p$ the particles density and $\rho_f$ the fluid phase density initially in the current.
-In the set of experiments presented here $\rho_f\le \rho_a$.
-The velocity scale $U_0$ corresponds to an idealized transfer of initial potential energy to kinetic energy along the slope, i.e.
+%%%%%%Figure 2  
+\begin{figure}
+	\centering
+	\includegraphics{figure2.pdf}
+	\caption{\textbf{Explored parameter space.} Each symbol corresponds to a single experimental run. Red-edged symbols correspond to experimental runs starting with deposited particles inside the lock.}
+	\label{fig:fig2}
+\end{figure}
+%%%%%%%%%%
+
+Following section~\ref{sec:intro} and \ref{sec:methods}, the relevant length, velocity and time scales for the propagation of a lock-release gravity current are the lock-length $L_{0}$, the buoyancy velocity $U_{0}=\sqrt{g' H_0 \cos\alpha}$ and $t_{0} = L_{0}/U_{0}$, respectively.
+%
+Note that here, the reduced gravity $g' = g (\rho_{\text{m}}-\rho_{\text{a}})/\rho_{\text{a}}$ is different from $\tilde{g}$ defined in \eqref{eq:solUc} in order to include the dependence of the front velocity on the density ratio $\rho_{\text{m}}/\rho_{\text{a}}$ in the definition of $U_{0}$ (see \eqref{eq:Froude_Huppert1980} in section~\ref{sec:intro_lockrelease}), and in the resulting dimensionless quantities.
+As such, the velocity scale $U_0$ corresponds to an idealized transfer of initial potential energy to kinetic energy along the slope, i.e.
 a collapse of the initial column.
 
-Accordingly,
+In addition to the particle volume fraction $\phi$ and bottom slope $\alpha$, we define the following control dimensionless parameters
 \begin{equation}
@@
 \begin{equation}
-	\displaystyle Re = \frac{U_0 H_0\rho_m}{\eta_f},
+	\displaystyle \mathcal{A}t = \frac{\rho_{\text{m}}-\rho_{\text{a}}}{\rho_{\text{a}}},
 \end{equation}
 \begin{equation}
-	\displaystyle At = \frac{\rho_m-\rho_a}{\rho_a},
+	\displaystyle \mathcal{R}e = \frac{U_0 H_0\rho_{\text m}}{\eta_{\text f}},
 \end{equation}
 \begin{equation}
-	\displaystyle St = \frac{V_s}{U_0},
+	\displaystyle \mathcal{S}t = \frac{L_{0} V_{\text s}}{H_{0} U_0},
 \end{equation}
-with $\eta_f$ the interstitial fluid viscosity and $V_s$ the Stokes settling velocity in the interstitial fluid phase labelled $f$.
+%
+with $\eta_{{\text f}}$ is the interstitial fluid viscosity.
+Note that $\mathcal{S}t=S/a$ with $S$ is the dimensionless settling velocity as defined in section \ref{sec:intro_PLGC}.
+As an estimation of $V_{\text s}$, we neglect possible turbulent drag effects and particle-particle interactions in the settling process, leading to the Stokes settling velocity $V_{\text s} = d^{2}(\rho_{\text{p}} - \rho_{\text{f}})g/(18\eta_{{\text f}})$.
+%
+In the following, we assume that the particle influence on the current dynamics can be quantified by $\phi$ and $\mathcal{S}t$, and discard other dimensionless parameters involving grain properties (e.g.
+$H_{0}/d$, $\rho_{\text p}/\rho_{\text{f}}$).
+%
+Likewise, excess buoyancy effects will be taken into account through $U_{0}$, as $\mathcal{A}t$ is mostly used to quantify non-Boussinesq effects, which can be neglected here as for the large majority of the experiments and simulations, $\mathcal{A}t < 0.2$.
 
-Then, the dynamics of the current are characterized by several dimensionless variables.
-The most common one to be analyzed in gravity-driven avalanches is the dimensionless front velocity, also referred to as the Froude number,
-\begin{equation}
-	\displaystyle \mathcal{F}r =\frac{U_c}{U_0},
-\end{equation}
-where $U_c$ is the front velocity of the current.
-Note that $U_c$ is time-dependent, and so is $\mathcal{F}r$, i.e.
-that $\mathcal{F}r$ as defined here is not a control parameter.
-(The Froude number which would be defined as a control parameter for a lock-release configuration would be $Fr\equiv 1$, and is thus not relevant to predict the flow dynamics.)
+The dimensionless control parameters $\phi$, $\alpha$, $\mathcal{S}t$ and $\mathcal{R}e$, used in the 3 experimental datasets and in the numerical dataset, are shown in figure~\ref{fig:fig2}.
+Here, $\phi$, $\alpha$ and $\mathcal{S}t$ are reported against $\mathcal{R}e$.
+It is shown that they roughly cover the range $\phi \in [0.001, 0.5]$ (see description of dataset 2 in section~\ref{sec:datasets} for experiments with $\phi > 0.5$), $\alpha \in [0,45]^\circ$, $\mathcal{S}t \in [3{\times}10^{-3}, 10^{1}]$ and $\mathcal{R}e \in [10^3, 10^6]$.
+To finish with, the aspect ratio $a=H_0/L_0$ is varied in the range $[0.5, 5]$, but not systematically.
+As its influence can not be systematically quantified, this will rather be qualitatively discussed in section~\ref{sec:conclusion}.
 
-\section{Methods}
+\subsection{Macroscopic description of the current dynamics}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 2
-\begin{figure}
-	\centering
-	% \includegraphics{Figures/Figure_lit_plat/Figure_lit_plat.pdf}
-	\includegraphics{figure2.pdf}
-	\caption{\textbf{Explored parameter space.} Each symbol correspond to a single experimental run.}
-	\label{fig:fig2}
-\end{figure}
-
-\subsection{Experimental setups}
-
-Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed on the front dynamics, often referred to as the $\mathcal{F}r$ number, and the front shape.
-Then prediction of the different flow regimes associated with PLGCs shall be entirely included into these observables.
-In order to characterize the different flow regimes for PLGCs induced by a finite volume of material initially at “rest”, 3 dam-break-type devices have been used to cover a wide range of control parameters, as $\alpha \in [0^o,30^o]$, $\phi\in[0.1\%,20\%]$, $At\in[???]$, $Re\in[???]$, $St\in[???]$, $a\in??$, as specifically defined in section \ref{sec:dimensionlessmap}.
-These experimental devices are supplemented with Euler-Euler numerical simulations to confirm the relevance of such flow regime description to characterize the dynamics of PLGCs.
-
-\paragraph{Set-up 1}
-
-Specifications of IMFT set-up
-
-\paragraph{Set-up 2}
-
-Specifications of LEGI set-up
-
-\paragraph{Set-up 3}
-
-Specifications of LEMTA set-up
-
-\subsection{Datasets}
-
-On these 3 devices, a total of $N$ experiments are presented here, complemented by $N$ numerical simulations using SedFoam (see description below).
-They are classified into 4 different datasets, described below.
-
-\paragraph{Dataset 1}
-
-Description Dataset LEGI
-
-\paragraph{Dataset 2}
-
-Description Dataset IMFT
-
-\paragraph{Dataset 3}
-
-Description Dataset LEGI/IMFT
-
-\paragraph{Dataset 4}
-
-Description Dataset LEMTA
-
-\paragraph{SedFoam}
-
-Description Dataset SedFoam
-
-\subsection{Fitting procedure and front position analysis}
-
-For each experimental/numerical run, we extract the current front position as a function of time (see Fig.
-\ref{fig:fig3}a).
-A few selected runs are shown on figure~\ref{fig:fig3}b in non-dimensional time coordinates.
-After a short transient acceleration phase, most currents exhibit at first a constant velocity corresponding to the so-called \emph{slumping} regime.
-After this, the velocity can decrease, either because of settling induced buoyancy loss, or because the current enters the \emph{inertial} regime.
-
-It can be qualitatively seen that increasing the bottom slope increases the slumping velocity, while increasing the Stokes number shortens the duration of the constant-velocity regime.
-%
-To properly quantify that, we fit all front propagation curves (except Dataset 4) with the following equation:
-
-\begin{equation}
-	\label{eq:fit_eq}
-	\frac{x_{\rm f}}{L_{0}} = x_{i} + \mathcal{F}r \frac{t}{t_{0}} - \frac{\tilde{\lambda}}{2} \left(\frac{t}{t_{0}}\right)^{2},
-\end{equation}
-where $\mathcal{F}r$ corresponds now to the non-dimensional current velocity during the slumping regime, and $\tilde{\lambda}$ to the dissipation.
-The fit is performed between $t/t_{0} = 5$, to avoid the initial transient phase, and $t/t_{0} = 30$, the transition time from the slumping to the inertial regime~\citep{refs}.
-Note that $x_{i}$ is kept constant equal to 0 for Dataset 1, 2 and 3, but had to be left free to adjust for the SedFoam simulations \textcolor{red}{comment this}.
-For runs with strong dissipation ($\tilde{\lambda} > 2 {\times} 10^{-2}$), we add a quartic term to \eqref{eq:fit_eq} to improve the fit quality and obtain a better estimation of $\mathcal{F}r$ and $\tilde{\lambda}$.
-We checked that our results does not depend on the presence of this quartic term, and that the linear and quadratic ones remain much larger.
-
-For Dataset 4, the time resolution as well as the experiment duration does not allow to extract the dissipation. Therefore, we simply fit the whole available points with a linear function.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure 3
-\begin{figure*}
+%%%%%%%%%%%%%%%%%%%% Figure 3
+\begin{figure*}[ht]
 	\centering
@@
 \end{figure*}
+%%%%%%%%%%%%%%%%%%%%
+
+The current dynamics is characterized by the evolution of several dimensionless variables.
+Here, as stated in section~\ref{sec:intro}, we aim at giving a macroscopic description of the initial current dynamics across ranges of several parameters spanning multiple orders of magnitude.
+The wide diversity of experiments and simulations results in a large variety of currents which does not allow for an exhaustive description of their individual dynamics.
+We rather choose to focus on a single variable, the front position $x_{\text{f}}$, which has been shown in the literature to be a good proxy to describe the current dynamics~\citep[e.g.][]{bonnecaze1993particle, Chowdhury2011, Adduce2012}.
+
+Figure~\ref{fig:fig3}(a) shows the front position of all experiments and simulations in dimensional coordinates, and figure~\ref{fig:fig3}(b) exhibits some selected runs in dimensionless coordinates.
+After a short transient acceleration phase, most currents exhibit at first a constant velocity phase, which corresponds to the so-called \emph{slumping} regime discussed in section \ref{sec:intro}.
+Afterwards, the velocity becomes time-dependent, often decreasing, due to one or multiple of the attenuation processes described in section~\ref{sec:intro}.
+%
+Qualitatively, it can be seen that increasing the bottom slope increases the slumping velocity while increasing the Stokes number shortens the duration of the constant-velocity phase (see fig~\ref{fig:fig3}).
+
+In order to properly quantify the slumping regime, i.e.
+the associated front velocity and duration, it is necessary to prescribe the relevant attenuation process leading to the regime transition.
+Yet, this is a priori unknown given the variety of involved processes in the wide range of PLGCs parameters used here.
+We thus propose to overcome this issue by fitting experimental and numerical front evolution by a second-order temporal expansion of the front position $x_{\text f}(t)$ as
+\begin{equation}
+	\label{eq:fit_eq}
+	\frac{x_{\text f}}{L_{0}} = \frac{x_{i}}{L_0} + \mathcal{F}r \left[\frac{t}{t_{0}} - \frac{1}{\tau} \left(\frac{t}{t_{0}}\right)^{2}\right],
+\end{equation}
+where $\mathcal{F}r = U_{\text{c}}/U_{0}$ corresponds to the dimensionless current velocity during the slumping regime and $\tau$ to a dimensionless attenuation time scale related to $X_{\text{t}}$ as
+\begin{equation}
+	\label{eq:link_Xt_tau}
+	\frac{X_{\text{t}}}{L_{0}} \propto \tau\mathcal{F}r.
+\end{equation}
+Note that the previous expansion \eqref{eq:fit_eq} assumes a constant velocity and constant deceleration trend of the front evolution.
+The latter is relevant to extract the time scale of existence of the slumping regime, i.e.
+the linear component, as long as its dynamics has only weakly deviated from such linear evolution.
+It is thus required to consider such a solution only during a brief time scale beyond the linear slumping regime.
+%as defined here, $\mathcal{F}r$ is not a control parameter, but a dimensionless variable characteristic of the initial current dynamics (The Froude number which would be defined as a control parameter for a lock-release configuration would be $\mathcal{F}r \equiv U_{0}/U_{0} = 1$, and is thus not relevant to describe the flow dynamics).
+
+In the following, we fit all front propagation curves (except for dataset 3) with \eqref{eq:fit_eq} between $t/t_{0} = 5$, to avoid the initial transient phase, and $t/t_{0} = 30$, the transition time from the slumping to the inertial regime~\citep[similar to $X_{\text{t}}/L_{0}\approx 10$ as mentioned in section \ref{sec:intro_lockrelease}; see also e.g.][]{Rottman1983, Sher2015, Ottolenghi2016}.
+Note that $x_{i}$ is kept constant equal to 0 for datasets 1 and 2, but had to be left free to adjust for the SedFoam simulations.
+This is probably due to the influence of the initial condition.
+In most of the experiments, due to the stirring, the initial condition exhibits velocity fluctuations as well as concentration fluctuations which might influence the early current behaviour.
+In the 2D RANS two-fluid simulations using sedFOAM, the velocities are zero and the concentration is homogeneous, both without any fluctuations.
+For runs with strong attenuation ($\tau < 10^{2}$), we add a quartic term to \eqref{eq:fit_eq} to improve the fit quality and obtain a better estimation of $\mathcal{F}r$ and $\tau$.
+We checked that our results do not depend on the presence of this quartic term and that the linear and quadratic ones remain much larger.
+Finally, for dataset 3, the time resolution and the experiment duration do not allow for inferring the quadratic term.
+Therefore, we simply fit the whole available points with a linear function.
 
 \section{Results}
+\label{sec:results}
 
-\subsection{Influence of the slope on the front velocity}
+\subsection{Influence of the bottom slope, $\alpha$}
+\label{sec:influence_slope}
 
-The influence of the slope angle $\alpha$ on the dimensionless front velocity $\mathcal{F}r$ is presented in figure \ref{fig:fig4}.
-Even if the entire set of experiments and numerical simulations are reported here, results are highlighted for a nearly constant $St\approx 5\times 10^{-2}$ and $\phi < 0.4$ for clarity.
+In this section, we focus on the influence of the bottom slope $\alpha$ on the current dynamics.
+Hence, figure~\ref{fig:fig4} shows the dimensionless front velocity $\mathcal{F}r$ as a function of $\alpha$ for the entire set of experiments.
 
-\begin{figure*}
+Despite the scatter in our experimental points, $\mathcal{F}r$ clearly exhibits an increasing trend with $\alpha$.
+Note however that the exact shape of this trend remains unclear, especially close to $\alpha=0^\circ$, where the various datasets may suggest slightly different trends.
+Note also the large scatter at $\alpha=45^\circ$, mostly related to experimental data points with $\phi > 0.45$ (transparent markers) and a settled initial condition (red-edged markers, see description of dataset 2 in section~\ref{sec:datasets}).
+%
+Such evolution of $\mathcal{F}r$ with $\alpha$ has only been reported few studies of homogeneous saline currents with either experiments on a smaller range of slopes $\alpha$ \citep{Maxworthy2007} or numerical simulations for small aspect-ratio $a = 0.1$ \citet{Birman2007}.
+This suggests a mechanism independent of the properties of the dense fluid, either saline or particle-laden.
+Interestingly, our data exhibits a much larger dependency of $\mathcal{F}r$ on $\alpha$ than reported in these simulations at small $a$.
+This difference is thus probably due to a significant difference in the lock aspect-ratio, which has been shown to significantly affect $\mathcal{F}r$ for $\alpha=0$ (see \citet{Bonometti2011} for instance), rather than from the presence of particles, as further discussed in sections~\ref{sec:influence_stokes} and~\ref{sec:influence_phi}.
+
+It shall be noted that this dependency of the slumping velocity on the bottom slope exists despite the slumping equilibrium being unmodified by the presence of a slope-induced acceleration (see section~\ref{sec:intro} for details).
+As suggested in \citet{Gadal2023}, the latter acceleration may instead affect the complicated dynamics of the early transient acceleration phase, involving strong vertical motions and significant interfacial friction~\citep{Cantero2007}.
+%
+One can write an energetic balance between the initial state and the end of this early acceleration phase \citep{Gadal2023}
+\begin{equation}
+	\label{eq:energetic_balance}
+	\left[\frac{1}{2}\rho_{0}u_{\text c}^{2} - B \Delta\rho g \sin\alpha L\right] - A \Delta\rho g \cos\alpha h_{0} =  - \frac{1}{2}c_{\text d}\rho_{0}u_{\text c}^{2} \frac{L}{h_{0}},
+\end{equation}
+where $A$, $B$ and $c_{\text d}$ are scaling constants, and $L$ accounts for the along slope distance over which the current moves during this phase.
+\citet{Cantero2007} found it to be independent of the lock aspect ratio or initial buoyancy, $L = 0.3 h_{0}$.
+Hence, we use here $L = D h_{0}$ with $D$ constant, which results in
+\begin{equation}
+	\label{eq:Fr_th}
+	\mathcal{F}r = \frac{U_{\text c}}{U_0} = \frac{Fr_0}{\sqrt{1 + C_{\text D}}} \sqrt{1 + \frac{\tan\alpha}{S}},
+\end{equation}
+where $Fr_0$, $C_{\text D}$ and $S$ are new constants depending on $A$, $B$, $c_{\text d}$ and $D$.
+As shown in Fig.~\ref{fig:fig4}, this simple model is able to reproduce the observed variation of the dimensionless velocity in the bottom slope in the range [$5^\circ$, $35^\circ$].
+%
+For $\alpha < 5^\circ$, experimental data points are significantly scattered, and \eqref{eq:Fr_th} almost systematically overpredicts the $\mathcal{F}r$ values.
+A plausible explanation for this discrepancy could be the difference in the width-to-depth ratio of the flumes (and associated wall dissipation) and in aspect ratios of the lock, which are not addressed in the energetic balance leading to \eqref{eq:Fr_th}.
+%
+For $\alpha > 35^\circ$, the measured dimensionless velocities ($\mathcal{F}r$) slowly stand above equation (\ref{eq:Fr_th}) curve.
+In this region of large slope angles, it is likely that the equilibrium between the forces becomes a lot more in favour of the so-called slope effect than the pressure gradient.
+Consequently, the scaling constants may have to be adapted to this major change in the influence of the forces.
+
+\begin{figure*}[ht]
 	\centering
 	\includegraphics{figure4.pdf}
-	\caption{\textbf{Influence of the bottom slope.} (a) Current Froude number as a function of the bottom slope.
-(b) Current Froude number corrected by a prefector $\sqrt{a}$ as a function of the bottom slope.
-Here, plain symbols corresponds to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
+	\caption{\textbf{Influence of the bottom slope} on the current Froude number.
+Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
+The black plain line indicates the prediction of \eqref{eq:Fr_th} for $Fr_{0} = 0.5$, $C_{\text D} = 0.4$ and $S = 0.25$.
+The black dashed line indicates the fit to the data from the numerical simulation of \citet{Birman2007} for saline homogeneous gravity currents.}
 	\label{fig:fig4}
@@
 
-\subsection{Influence of $St$}
+\subsection{Influence of the Stokes number, $\mathcal{S}t$}
+\label{sec:influence_stokes}
 
-\begin{figure*}
+\begin{figure*}[ht]
 	\centering
 	\includegraphics{figure5.pdf}
-	\caption{\textbf{Influence of particle settling.} Current Froude number $\mathcal{F}r$ and attenuation $\tilde{\lambda}$ as a function of the ratio between the Stokes number and the lock aspect-ratio, for different ranges of bottom slopes.
-Here, we selected runs for $\phi < 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).}
+	\caption{\textbf{Influence of particle settling.} Current Froude number $\mathcal{F}r$ and attenuation $1/\tau$ as a function of the ratio between the Stokes number and the lock aspect-ratio, for different ranges of bottom slopes.
+Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
+Dotted lines mark either the average of points with $\phi < 0.45$ (a, c, e, g) or the 0 baseline (b, d, f, h).
+On (e, f), the dashed line indicates the average value for homogeneous saline currents ($St \equiv 0$) of dataset 1.
+On (b, d, f, h), the black line is $1/\tau = 0.27 (St - 0.015)$.}
 	\label{fig:fig5}
@@
 
-Discuss figure \ref{fig:fig5}
+In this section, we focus on the influence of settling on the current dynamics.
+The influence of the Stokes number $\mathcal{S}t$ on the dimensionless front velocity $\mathcal{F}r$ and attenuation $\tau$ is presented in figure~\ref{fig:fig5} for different bottom slopes $\alpha$.
+
+At the first order, the dimensionless front velocity does not depend on $\mathcal{S}t$ (see figure~\ref{fig:fig5}a, c, e).
+However, this is not the case for the attenuation time scale $\tau$.
+For $\alpha \approx 7^\circ$, the attenuation ($1/\tau$) is constant below a threshold $St_{\text c} \simeq 10^{-2}$, but increases rather linearly above this value (see the black line in figure~\ref{fig:fig5}(d), note the semi-log x-scale).
+
+These two regimes can be understood as follows.
+First, assuming that the current decelerates due to its transition from the slumping regime towards the so-called \emph{inertial} regime, combining \eqref{eq:Xt_inertial} and \eqref{eq:link_Xt_tau} results in $\tau \sim cste$.
+Then, if the current decelerates due to a settling-induced loss of buoyancy, combining \eqref{eq:Xt_settling} and \eqref{eq:link_Xt_tau} results in
+%
+\begin{equation}
+	\frac{1}{\tau} \sim \mathcal{S}t.
+\end{equation}
+%
+This corresponds well to the observed trend in figure~\ref{fig:fig5}(d), hence confirming that above $St_{\text c} \simeq 10^{-2}$, particle settling controls the end of the slumping regime.
+Similar results can also be extrapolated for $\alpha = 0^\circ$ and $\alpha = 15^\circ$, although additional points are required to assess the dependence of the transitional threshold and of the slope of the linear increase with $\phi$ or $\alpha$.
+
+For larger slopes, our experiments do not allow us to discriminate the influence of the Stokes number (see figure~\ref{fig:fig5}(g,h)).
+From further exploring of these regimes might arise singular effects as, at large slopes, a significant component of the settling occurs along the slope.
+It might therefore contribute to increasing the current inertia, or at least to decreasing the expected attenuation.
 
@@
 	\caption{\textbf{Influence of particle volume fraction.} Current Froude number as a function of the initial particle volume fraction for (a) $\alpha \approx 0^\circ$, and (b) $\alpha \approx 45^\circ$.
-The inset in (a) shows the plot in linear scale.}
+The inset in (a) shows the plot in linear scale.
+The black line indicates the prediction of \eqref{eq:Fr_th_eta} for $Fr_{0} = 0.5$, $C_{\text D} = 0.4$, $S = 0.25$ and $Re_{\text c} = 500$.}
 	\label{fig:fig6}
@@
 
-Discuss figure \ref{fig:fig6}
+In this section, we focus on the influence of the particle volume fraction $\phi$ on the current dynamics.
+Hence, figure~\ref{fig:fig6} shows the dimensionless front velocity $\mathcal{F}r$ as a function of $\phi$ for $\alpha=0^\circ$ and $\alpha=45^\circ$.
+
+For $\alpha=0^\circ$, figure~\ref{fig:fig6}(a) shows that $\mathcal{F}r$ is independent of $\phi$ up to a critical value $\phi_{c} \simeq 0.45$, above which the Froude number abruptly decreases.
+This results from the large increase in effective viscosity induced by the presence of particles~\citep{stickel2005fluid}, which probably enhances dissipation at the bottom and on the sidewalls.
+Furthermore, this decrease in Froude number is associated with a reduction of the height of the current in the presence of a strong mass gradient near the front, as noted by \citet{Schneider2023}.
+
+The energetic balance presented in section~\ref{sec:influence_slope} can be adapted to include viscous dissipation effects, by replacing the constant $c_{\text d}$ in \eqref{eq:energetic_balance} by:
+\begin{equation}
+	c_{\text d}\left(1 + \frac{E}{\mathcal{R}e}\frac{U_0}{U_{\text c}}\frac{\eta_{\text eff}}{\eta_{\text f}} \right),
+\end{equation}
+where $E$ is a new scaling constant, and $\eta_{\text eff}$ is the effective suspension viscosity accounting for the presence of particles. This results in
+\begin{equation}
+	\label{eq:Fr_th_eta}
+	\begin{split}
+		\mathcal{F}r & =  \frac{1}{1 + C_{\text D}} \left[-\frac{Re_{\text c}}{\mathcal{R}e}\frac{\eta_{\text eff}}{\eta_{\text f}} \right. \\
+		&\left. +\sqrt{\left(\frac{Re_{\text c}}{\mathcal{R}e}\frac{\eta_{\text eff}}{\eta_{\text f}} \right)^{2} + Fr_0^{2} \left(1 + \left[1 + C_{\text D}\right]\frac{\tan\alpha}{S} \right)}  \right],
+	\end{split}
+\end{equation}
+where $Re_{\text c}$ is a new constant.
+The effect of the particles on the effective viscosity is based on the correlation proposed by \citet{krieger1959mechanism} and validated in many configurations:
+\begin{equation}
+	\label{eq:Krieger}
+	\frac{\eta_{\text eff}}{\eta_{\text f}} = \left(1 - \frac{\phi}{\phi_{\text m}}\right)^{-(5/2)\phi_{\text m}},
+\end{equation}
+where $\phi_{\text m} = 0.585$ is the maximum volume fraction~\citep{boyer2011unifying}.
+%
+As shown in figure~\ref{fig:fig6}(a), the combination of equations \eqref{eq:Fr_th_eta} and \eqref{eq:Krieger} can reproduce quantitatively the observed influence of $\phi$ on the dimensionless velocity for $\alpha=0^\circ$.
+This agreement does not depend on the choice of the viscosity model, and using other available rheological models from the literature~\citep{stickel2005fluid, boyer2011unifying} leads to the same conclusions.
+
+For $\alpha=45^\circ$, a decreasing trend of the dimensionless front velocity is also observed despite the dispersion of experimental points.
+However, the shape of this trend is less clear, especially because data at large $\phi$ correspond to a different initial condition (red-edged markers, see description of dataset 2 in section~\ref{sec:datasets}), which makes the interpretation difficult.
+The combination of \eqref{eq:Fr_th_eta} and \eqref{eq:Krieger} is shown for completeness.
+
+\section{Conclusion and perspectives}
+\label{sec:conclusion}
+
+\begin{figure*}[ht]
+	\centering
+	\includegraphics{figures/figure7.pdf}
+	\caption{Comparison between 2D (lines) and 3D (grey scale) two-fluids numerical simulation at $t = 15~\textrm{s}$ ($t/t_{0} = 45$) for $\alpha = 45^\circ$, $\phi = 0.1$ and $St = 2{\times}10^{-2}$.}
+	\label{fig:fig7}
+\end{figure*}
+
+In this chapter, we summarize the effect of various dimensionless parameters on the early propagation of finite-volume turbidity currents.
+Hence, by using data from multiple experimental devices and from two-fluid numerical simulations, we explore a wide range of $\phi$, $\alpha$, $\mathcal{S}t$ and $\mathcal{R}e$.
+%
+By reducing the early dynamics to a constant-velocity propagation at a dimensionless velocity $\mathcal{F}r$ during a dimensionless characteristic time $\tau$ (see~\eqref{eq:fit_eq} for additional details), we show that:
+\begin{itemize}
+	\item $\mathcal{F}r$ is independent of $\mathcal{S}t$, but increases with $\alpha$ and drastically decreases with $\phi$ above a critical value $\phi_{\text c} \approx 0.45$.
+These variations are well explained by a simple energy balance during the early transient acceleration phase occurring before the slumping regime for which the front evolves at constant velocity.
+This early stage dynamics is associated with a non-conservative and slope-dependent transfer of the initial potential energy towards the along-slope kinetic energy (see \eqref{eq:Fr_th}, \eqref{eq:Fr_th_eta}).
+The role of the particles in the dissipation is captured by considering an effective suspension viscosity model (\eqref{eq:Fr_th_eta}).
+
+	\item $1/\tau$ depends on $\mathcal{S}t$ above a transition value $St_{\text c} \approx 10^{-2}$.
+Below $St_{\text c}$, the currents are then similar to a gravity current induced by a saline dense fluid, while above $St_{\text c}$, dissipation $1/\tau$ increases as $1/\tau \sim \mathcal{S}t$, or equivalently the duration of the slumping regime decreases as $\tau \sim \mathcal{S}t^{-1}$.
+\end{itemize}
+Although probably hidden by the first-order dependency (velocity) of the dynamics on these parameters, note that we also expect $\tau$ to depend on $\alpha$ and $\phi$.
+%
+Importantly, all these variations can also be quantitatively reproduced by the 2D vertical numerical simulations using a two-fluid model.
+This validates the use of continuous approaches over looking at the grain scale dynamics to understand and model these parts of the current dynamics.
+
+% \begin{itemize}
+%     \item Rationalize/Compare PLGCs dynamics using multiple devices
+%     \item Si on ecrit vfront = \mathcal{F}r (1-1/tau t2), \mathcal{F}r (alpha,phi) et \mathcal{F}r ne depend pas de \mathcal{S}t.
+attenuation, tau (alpha, phi, \mathcal{S}t) on a montre la dependece en (\mathcal{S}t) et celle en alpha, phi car domine par l ordre 1.
+%     \item Numerique diphasique EUler-Euler 2D, permet de retrouver ces tendances. Donc l echelle du grain n est pas primordiale pour comprendre tout ca.
+% \end{itemize}
+
+Altogether, the influence of $\phi$, $\mathcal{S}t$, $\alpha$ on the slumping dynamics of PLGCs characterised in the present study, highlights the importance of a better understanding of the early stage transfer of initial potential energy towards the along-slope kinetic energy.
+To predict the longtime scale evolution of PLGCs, the knowledge and the characterisation of their early-stage dynamics remain fundamental to prescribe an initial input for large-scale modelling.
+
+However, details of the current shape and related mixing remain unclear.
+In particular, even even if the front velocity is well captured by the 2D simulations, they poorly capture the current shapes, as shown in figures~\ref{fig:fig1} and~\ref{fig:fig7}.
+This would require 3D simulations to properly capture the turbulent mixing at the current interface.
+For validation, advanced experimental metrologies are thus required to extract accurate measurements of the turbulent mixing at the interface and responsible for ambient entrainment in the current.
+These should be an important focus for future work.
+%
+Moreover, we also found that the choice in kinetic theory closures for the particle phases in the simulations can modify the measured velocities up to $30~\%$ (especially at larger slopes), which emphasizes the role of particle-particle interactions in the macroscopic dynamics of PLGCs.
+These are particularly important in zones of large volume fractions and are therefore key to quantifying the ability of a current to erode a loose sediment bottom or the dynamics of currents originating from deposited sediment.
+Both are common situations in natural flows and require the focus of future works.
+
+%\begin{itemize}
+%   \item Details of the current shape and associated mixing remains unclear
+%  \item Then 3D numerical simulations are required, as 2D pourly capture the shape.
+%\end{itemize}
+
+% \julien{
+% \begin{itemize}
+%     \item Influence of kinetic theory closures suggests that the particle-particle interactions play a role in the dynamics of the PLGC.
+%     \item 3D effects matter on the turbulent mixing at the current interface (add a figure with 2 snapshots)
+%     \item Influence of the initial condition in the numerical simulations: noise on volume fraction, velocities, ... Relevant for 2D RANS simulations? 
+% \end{itemize}
+% }
+
+% \cyril{
+% \begin{itemize}
+%     \item Influence of $a = H_{0}/L_{0}$ ? (add a figure $Fr = f(a)$ for $\alpha = 0$, $\alpha = 7$ ?)
+%     \item Influence of $H_{0}/d$ for setting/maintaining particles in suspension (`pore pressure')? Relevant for experiments with initial deposited experiments and for strong settling?
+%     \item Part of $V_{s}\sin\alpha$ that increases current inertia ?
+% \end{itemize}
+% }
+
+% \begin{itemize}
+%     \item Future works need to focus on the link between this dynamics and the ability to erode the bottom, as observed in natural flow.
+% \end{itemize}
+
+\subsection*{Data availability}
+
+Data and processing scripts are available at [public repository DOI] (will be completed upon publication).
 
 \subsection*{Acknowledgment}
+We thank Jean-Dominique Barron (IMFT) and Sébastien Cazin (IMFT) for their support in carrying out the experiments on setup-up 1.
+%
+Sylvain Dauge is acknowledged for the full design of setup 2. Hervé Bellot and Brivaël Collin are thanked for their experimental support.
+%
+We also acknowledge Hugo Divel for his contribution to the preliminary lock-release simulations using SedFoam.
+%
+Finally, we would like to acknowledge the contributors of the open-source numerical Python libraries, including Matplotlib \citep{Hunter2007}, NumPy \citep{Harris2020} and SciPy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
 
 \subsection*{Fundings}
+We acknowledge financial support from the French National Research Agency Grants, ANR-19-CE30-0041/PALAGRAM.
 

### 2024-02-21 analysis/analysis.py
temp commit

@@ -2,15 +2,14 @@ import glob
 import os
 import shutil
 
-import numpy as np
 import formatting_parameters as fp
+import numpy as np
 from lmfit import Model
 from lmfit.model import save_modelresult
 from netCDF4 import Dataset
 
 
-def create_variable(netcdf_group, name, data, dimensions=None, std=None,
-                    unit=None, comments=None, type='float64'):
+def create_variable(netcdf_group, name, data, dimensions=None, std=None, unit=None, comments=None, type="float64"):
     if dimensions is not None:
         var = netcdf_group.createVariable(name, type, (dimensions))
     else:
@@ -25,58 +24,58 @@ def create_variable(netcdf_group, name, data, dimensions=None, std=None,
 
 
 def polyFULL(t, Fr, lamb, xi, c, d):
-    return xi + Fr*(t - lamb*t**2 + c*t**3 + d*t**4)
+    return xi + Fr * (t - lamb * t**2 + c * t**3 + d * t**4)
 
 
 def Stokes_Velocity(d, mu, rho_p, rho_f, g):
-    return d**2*(rho_p - rho_f) * g/mu/18
+    return d**2 * (rho_p - rho_f) * g / mu / 18
 
 
 def determine_fit_props(author, alpha, run, params):
-    params['xi'].vary = False
-    params['Fr'].vary = True
-    params['lamb'].vary = True
-    params['c'].vary = False
-    params['d'].vary = False
+    params["xi"].vary = False
+    params["Fr"].vary = True
+    params["lamb"].vary = True
+    params["c"].vary = False
+    params["d"].vary = False
     t_bounds = [5, 30]
     #
-    if author == 'Jean':
+    if author == "Jean":
         t_bounds = [0, 40]
-        params['lamb'].vary = False
-        params['xi'].vary = True
-    if author == 'Julien':
-        params['xi'].vary = True
-        if run == 'run42b_front.nc':
+        params["lamb"].vary = False
+        params["xi"].vary = True
+    if author == "Julien":
+        params["xi"].vary = True
+        if run == "run42b_front.nc":
             t_bounds = [0, 8]
-        if run == 'run37_front.nc':
+        if run == "run37_front.nc":
             t_bounds = [0, 20]
-        if run == 'run42_front.nc':
+        if run == "run42_front.nc":
             t_bounds = [0, 20]
[…]

### 2024-02-21 paper/figures/figure_scripts/figure2.py

@@ -13,34 +13,49 @@ from netCDF4 import Dataset
 
 
 # %% Load data
-path_data = '../../../data/output_data'
-list_runs = glob.glob(os.path.join(path_data, '*.nc'))
-datasets = np.array([Dataset(run) for run in list_runs if Dataset(
-    run).particle_type != 'saline water'])
+path_data = "../../../data/output_data"
+list_runs = glob.glob(os.path.join(path_data, "*.nc"))
+datasets = np.array([Dataset(run) for run in list_runs if Dataset(run).particle_type != "saline water"])
 
 # %% create data vectors
-alpha, phi, Re, St, H0 = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
-                                    d.variables['Re'][:].data, d.variables['St'][:].data,
-                                    d.variables['H0'][:].data,
-                                    ] for d in datasets]).T
+alpha, phi, Re, St, H0 = np.array(
+    [
+        [
+            d.variables["alpha"][:].data,
+            d.variables["phi"][:].data,
+            d.variables["Re"][:].data,
+            d.variables["St"][:].data,
+            d.variables["H0"][:].data,
+        ]
+        for d in datasets
+    ]
+).T
 
-authors, particles = np.array([[d.author, d.particle_type,
-                                ] for d in datasets]).T
+authors, particles = np.array(
+    [
+        [
+            d.author,
+            d.particle_type,
+        ]
+        for d in datasets
+    ]
+).T
 
-Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
-              else d.variables['H0'][:].data for d in datasets])
+Ha = np.array(
+    [d.variables["H_a"][:].data if "H_a" in d.variables.keys() else d.variables["H0"][:].data for d in datasets]
+)
 
 # %% graphic vector for plots
 dataset_idx = np.vectorize(tp.datasets.get)(authors)
 
 markers = np.vectorize(tp.marker_style.get)(particles)
-markers[dataset_idx == '4'] = 's'
+markers[dataset_idx == "4"] = "s"
 
[…]

### 2024-02-21 paper/figures/figure_scripts/figure4.py

@@ -9,7 +9,7 @@ from matplotlib.colors import to_rgba
 from models import Birman, Froude
 from netCDF4 import Dataset
 
-plt.rcParams['xtick.top'] = False
+plt.rcParams["xtick.top"] = False
 
 
 def ang2sin(ang):
@@ -21,24 +21,40 @@ def sin2ang(sin):
 
 
 # %% Load data
-path_data = '../../../data/output_data'
-list_runs = glob.glob(os.path.join(path_data, '*.nc'))
+path_data = "../../../data/output_data"
+list_runs = glob.glob(os.path.join(path_data, "*.nc"))
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% create data vectors
-alpha, phi, Fr, St, H0 = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
-                                    d.variables['Fr'][:].data, d.variables['St'][:].data,
-                                    d.variables['H0'][:].data,
-                                    ] for d in datasets]).T
+alpha, phi, Fr, St, H0 = np.array(
+    [
+        [
+            d.variables["alpha"][:].data,
+            d.variables["phi"][:].data,
+            d.variables["Fr"][:].data,
+            d.variables["St"][:].data,
+            d.variables["H0"][:].data,
+        ]
+        for d in datasets
+    ]
+).T
 
-Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
-              else d.variables['H0'][:].data for d in datasets])
-authors, particles = np.array([[d.author, d.particle_type,
-                                ] for d in datasets]).T
+Ha = np.array(
+    [d.variables["H_a"][:].data if "H_a" in d.variables.keys() else d.variables["H0"][:].data for d in datasets]
+)
+authors, particles = np.array(
+    [
+        [
+            d.author,
+            d.particle_type,
+        ]
+        for d in datasets
+    ]
+).T
 
 # %% graphic specifications
 # %% masks for plot
-mask_phi = (phi < tp.phi_c)
+mask_phi = phi < tp.phi_c
 
 # %% graphic vector for plots
[…]

### 2024-02-21 paper/figures/figure_scripts/figure6.py

@@ -6,40 +6,57 @@ import matplotlib.pyplot as plt
 import matplotlib.transforms as mtransforms
 import numpy as np
 import template as tp
-from netCDF4 import Dataset
 from models import Froude, Krieger_viscosity
+from netCDF4 import Dataset
 
-plt.rcParams['figure.constrained_layout.hspace'] = 0
-plt.rcParams['figure.constrained_layout.h_pad'] = 0.005
+plt.rcParams["figure.constrained_layout.hspace"] = 0
+plt.rcParams["figure.constrained_layout.h_pad"] = 0.005
 
 
 # %% Load data
-path_data = '../../../data/output_data'
-list_runs = glob.glob(os.path.join(path_data, '*.nc'))
+path_data = "../../../data/output_data"
+list_runs = glob.glob(os.path.join(path_data, "*.nc"))
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% create data vectors
-alpha, phi, St, H0, L0, Fr = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
-                                        d.variables['St'][:].data, d.variables['H0'][:].data,
-                                        d.variables['L0'][:].data, d.variables['Fr'][:].data,
-                                        ] for d in datasets]).T
+alpha, phi, St, H0, L0, Fr = np.array(
+    [
+        [
+            d.variables["alpha"][:].data,
+            d.variables["phi"][:].data,
+            d.variables["St"][:].data,
+            d.variables["H0"][:].data,
+            d.variables["L0"][:].data,
+            d.variables["Fr"][:].data,
+        ]
+        for d in datasets
+    ]
+).T
 
-Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
-              else d.variables['H0'][:].data for d in datasets])
-authors, particles = np.array([[d.author, d.particle_type,
-                                ] for d in datasets]).T
+Ha = np.array(
+    [d.variables["H_a"][:].data if "H_a" in d.variables.keys() else d.variables["H0"][:].data for d in datasets]
+)
[…]

### 2024-02-21 paper/main.tex
Pass doc to single column for working
Adding file for answering referees

@@
-\documentclass[twocolumn]{article}
+% \documentclass[twocolumn]{article}
+\documentclass[12pt]{article}
 \usepackage[utf8]{inputenc}
@@
 
-\newcommand*{\cyril}{\textcolor{ForestGreen}}
-\newcommand*{\marie}{\textcolor{Cyan}}
-\newcommand*{\julien}{\textcolor{Brown}}
-\newcommand*{\yvan}{\textcolor{Blue}}
-
-\definecolor{jeanc}{RGB}{253,129,43}
-\newcommand*{\jean}{\textcolor{jeanc}}
+\usepackage{setspace}
+\doublespacing
 
@@
 
-\date{{\it\color{red}Deadline: May 31$^{st}$ - repoussé (mi juillet ou EuroMech?)}}
-
 \begin{document}
-
-\twocolumn[
 
@@
 \end{abstract}
-
-	\vspace{0.3cm}
-	{\it \color{red}REMINDER: Word limit of about 10,000 per article, plus 7 figures.}\\
-	\vspace{0.3cm}
-]
 
@@
 For a finite volume lock-release device on a mild enough slope, the slumping regime necessarily ends when the rarefaction wave associated with finite volume reaches the current front \citep[e.g.][]{Rottman1983, Hogg2006}.
-This leads to a new regime usually referred to as the inertial regime in the literature.
+This leads to a new regime usually referred to as the self-similar regime in the literature.
 This is observed on the front evolution as $U_{\text{c}}$ is no longer time independent but evolves as $t^{-1/3}$.
@@
 Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear.
-Overall, the value of $F$ in \eqref{eq:solUc} during the slumping inertial regime remains to clarify varying the influence of the mechanisms discussed so far.
+Overall, the value of $F$ in \eqref{eq:solUc} during the slumping regime remains to clarify varying the influence of the mechanisms discussed so far.
 

### 2024-02-21 paper/reply_to_referees/first_review/reply.tex
Adding file for answering referees

\documentclass[12pt]{article}
\usepackage{a4wide}
\usepackage[usenames]{color}

\usepackage{graphics}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{pifont}
\usepackage{float}
\usepackage{natbib}
\usepackage[top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}

\usepackage[dvipsnames]{xcolor}
\newcommand{\xx}{\textcolor{WildStrawberry}}
\newcommand{\pc}{\textcolor{ForestGreen}}
\newcommand{\cg}{\textcolor{Mulberry}}
\newcommand{\new}{\textcolor{blue}}
\newcommand{\todo}{\textcolor{red}}

\newcommand{\MinorReply}[3]{\filbreak\noindent
\textbf{Referee:}
\emph{#1 -- #2}

\textbf{Reply:} {\color{blue}{#3}}
}

\newcommand{\MajorReply}[2]{\noindent
\textbf{Referee:}
\emph{#1}

\medskip
\noindent\textbf{Reply:} {\color{blue}#2}
\vspace*{0.4 cm}}

\begin{document}

\noindent Dear Editor,\\

\noindent We thank you very much for considering our chapter ``Chapter 16: Particle-laden gravity currents: the lock-release slumping regime at the laboratory scale", by C. Gadal \emph{et al}, for publication in a book for the \emph{Geophysical Monograph Series}. We are grateful for the thoughtful referee reports, and for their comments and suggestions. \\

\noindent
Please find below a detailed reply to each of the comments and questions by the referees. Additionally, places where the text has been correspondingly edited, appear in red (removed text) and blue (added text) in the file \emph{diff.pdf}. \\

\noindent Sincerely, \\

\noindent The authors

%__________________________________________________________________________
\newpage
\section{Reply to referee 1}

\MajorReply{The authors discuss gravity currents in the lock-release configuration, where the densest phase is represented by sediments subject to a progressive decrease in concentration during propagation. The work is supported by an extensive data set of three experimental configurations with varying bottom slope. Two of these configurations are simulated with an open source code to which a module, "SedFoam", published by some of the co-authors of this paper, has been added. This is followed by a description of the experiments and an interpretation of the results, categorising the effects of some of the many parameters controlling the physical process. The paper is of particular interest, even though it does not provide clear results for some of the investigations carried out.
}{}

\begin{itemize}
    \item \MinorReply{Introduction}{The literature, although rich, lacks a number of contributions of some importance, mainly because they are themselves experimental. See, for example, the work of Zemach et al. 2017, which deals with some more complex configurations, with non-rectangular geometry, with stratified fluid, possibly in the partial depth lock release configuration. The various contributions should be catalogued and described in an appropriate way.}{}

    \item \MinorReply{Figure 1}{the q-v panels look out of place and the sequence is horizontal while the experiments are vertical. It is recommended that they be placed in a separate figure.}{}

    \item \MinorReply{Figure 3a}{Figure 3a is not meaningful due to the large amount of data shown. It is advisable to scale it down in dimensionless form, as is done in Figure 3b for 'selected runs', which should be specified.}{}

    \item \MinorReply{Figure 3b}{the alpha symbol is flanked by an arrow which appears to be meaningless. I assume that the arrow should be reversed.}{}

    \item \MinorReply{References}{the book by Ungarish should be better updated to the last edition.}{}
\end{itemize}

%__________________________________________________________________________
\newpage
\section{Reply to referee 1}

\MajorReply{This is generally a good contribution with a lot of interesting results. The following are few comments the authors might want to make use.}{}

\begin{itemize}
    \item \MinorReply{Line 64}{Line 64 is unclear}{}
    \item \MinorReply{Line 78}{The sentence starting in line 78 is too convoluted. Needs restructuring.}{}
    \item \MinorReply{Line 102}{what does it mean by $h_c$ can have different scaling?}{}
    \item \MinorReply{Lines 104--107}{It is unclear what the authors mean gt detachment. Also “similarly” may need to change to “similar”}{}
    \item \MinorReply{Section 1.1}{gives several different expressions for $X_t$. Some clarification as to where each apply will be useful.}{}
    \item \MinorReply{Lines 133-134}{where do length and time scale of particles play a role?}{}
    \item \MinorReply{Line211-212}{what does it mean 2D vertical and turbulence averaged? Are these 2D RANS simulations? Are the 2 coordinates horizontal, since vertical is averaged?}{}
    \item \MinorReply{In figure 2}{the color of the symbols outline must be the same. Otherwise it is hard to relate to the color legend}{}
    \item \MinorReply{figure 2}{some of the volume fraction data goes to 1.    }{}
    \item \MinorReply{}{Curve fits with such a large data scatter, requires some justification}{}
    \item \MinorReply{Line 461}{even even}{}
    \item \MinorReply{Line 466}{This should be an important….}{}
\end{itemize}

\end{document}

### 2024-03-01 paper/main.tex
temp commit modifying introduction

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{cuted}
@@
 	%
-	The objective of the present chapter is to provide a description at the macroscopic scale of the early moments of the flow, namely the \emph{slumping regime}, with respect to the relevant dimensionless parameters.
+	The objective of the present chapter is to describe the macroscopic scale of the early moments of the flow, namely the \emph{slumping regime}, with respect to the relevant dimensionless parameters.
 For this, we combine a total of 288 runs from three different lock-release devices and from two-fluids numerical simulations, which allow us to cover a large range of particle types (size and density), volume fractions, bottom slopes and geometries.
@@
 The latter comparison confirms the role of initial energy transfer from the initial state towards the slumping phase on the resulting dynamics.
-This initial process and its characterisation remain crucial to prescribe relevant initial conditions for large-scale predictive modelling.
+This initial process and its characterization remain crucial to prescribe relevant initial conditions for large-scale predictive modeling.
 
@@
 Density currents have been widely considered at the laboratory scale for their ubiquitous occurrence in geophysical flows \citep{Hopfinger1983,Simpson1999,Dufek2016}.
-As they are mostly associated with unsteady and non-uniform dynamics in nature, the lock-release device appeared to be a popular canonical configuration to mimic their behaviour at the laboratory scale \citep{Hacker1996,Shin2004,Nogueira2014}.
+As they are mostly associated with unsteady and non-uniform dynamics in nature, the lock-release device appeared to be a popular canonical configuration to mimic their behavior at the laboratory scale \citep{Hacker1996,Shin2004,Nogueira2014}.
 In particular, studies often focused on characterising the front dynamics and the front shape as a macroscopic description of the underlying complex dynamics \citep{Rottman1983,Marino2005,Hogg2006}.
@@
 The main reason is that most density currents are characterized by a small density ratio between current and ambient, making the assumption of inertialess ambient fluid unlikely.
-A dynamic front condition is now required, and is often inspired from the steady front configuration derived by \citet{Benjamin1968}, now also linking the ambient height $H_{\text a}$ to the front velocity.
-This front relation includes two unknowns, a front velocity and a front height, and is often used as a front boundary condition in shallow water numerical solver for gravity currents \cite[e.g.][]{Ungarish2009}.
+%
+Hence, a dynamic front condition is now required, typically inspired by the steady front configuration derived by \citet{Benjamin1968}, also involving the ambient height $H_{\text a}$.
+This front relation includes two unknowns, a front velocity and a front height, and is used as a front boundary condition in the depth-averaged modeling of gravity currents \cite[e.g.][]{ungarish2020gravity}.
 Then, past studies managed to link the front velocity to the current height close to the front.
@@
 	\label{eq:Froude_Huppert1980}
-	F = 0.5 \sqrt{\frac{\rho_{\text c}}{\rho_{\text a}}}\left(\frac{H_{\text a}}{H_0}\right)^{1/3}.
+	F = 0.5 \sqrt{\frac{\rho_{\text c}}{\rho_{\text a}}}\left(\frac{H_{\text a}}{H_0}\right)^{-1/3}.
 \end{equation}
 %
-The latter is relevant for not too deep ambient $H_{\text a}$ compared to the reservoir height $H_0$, which is actually the case in most experiments.
+The latter is relevant only when typically $H_0/H_{\text a} > 0.2$, which is the case in most experiments.
 Many configurations correspond to full-depth release ($H_{\text a}=H_0$), which leads to $F = 0.5(\rho_{\text c}/\rho_{\text a})^{1/2}$.
@@
 
-For a finite volume lock-release device on a mild enough slope, the slumping regime necessarily ends when the rarefaction wave associated with finite volume reaches the current front \citep[e.g.][]{Rottman1983, Hogg2006}.
-This leads to a new regime usually referred to as the self-similar regime in the literature.
+For a finite volume lock-release device, the opening of the gate also generates a backward propagating rarefaction wave or bore (depending on $H_{0}/H_{\rm a}$), which reflects on the wall at the end of the reservoir and then propagates towards the front~\citep[e.g.][]{ungarish2020gravity}.
+As it evolves faster than the current front, it eventually reaches it, ending the slumping regime and leading to a new regime usually referred to as the self-similar regime~\citep[e.g.][]{Rottman1983, Hogg2006}.
 This is observed on the front evolution as $U_{\text{c}}$ is no longer time independent but evolves as $t^{-1/3}$.
-A simple approach can be useful to provide a rapid estimation of the position $X_{\text t}$ or time $T_{\text t}$ at which this transition occurs.
-For that purpose, it is worth mentioning that such a transition occurs as a rarefaction wave evolves faster than the front.
-Then, assuming the front velocity $U_{\text{c}}$ to be different from the rarefaction wave velocity at the interface $\sqrt{\tilde{g}H_0}$, which is assumed to be constant along its entire propagation path for simplicity, this leads to a transition length scale $X_{\text t}$ written as
+%
+Experiments, numerical simulations and theoretical studies using shallow-water modeling have shown that the position $X_{\text t}$ at which this transition occurs typically writes as:
+%%
 \begin{equation}
 	\label{eq:Xt_inertial}
-	\displaystyle X_{\text t}\propto  \frac{2L_0 F}{1 - F}.
+	\displaystyle X_{\text t} = \mathcal{A} L_{0},
 \end{equation}
-Note that here $2$ comes from the path of propagation of the wave: emanating from the gate, it propagates backwards to the end of the reservoir, reflects on the wall and then propagates towards the front.
-During the slumping regime, the previous estimation of $F$ \eqref{eq:Froude_Huppert1980} for full depth release $H_0=H_{\text a}$ leads to $X_{\text t}\propto 2L_0/[2(\rho_{\text a}/\rho_{\text c})^{1/2}-1]$.
-Note that, although more advanced methods based on single-layer shallow-water models predict a similar result, $X_{\text t} = 2 L_0 F (1 + F/2)^{1/2}$ \citep{Hogg2006}, full depth-release experiments have found much larger transitional positions, i.e $X_{\text t} \approx 10 L_{0}$~\citep{Rottman1983}.
+%%
+where $\mathcal{A}$ is an increasing function of $H_{0}/H_{\rm a}$, and may depend on other parameters of the problem, e.g.
+$\rho_{\rm c}/\rho_{a}$.
+In the general case, its determination requires the numerical resolution of the two-layer shallow-water system of equations~\citep{Ungarish2005}.
+In the limit of deep currents (typically $H_{0}/H_{\rm a} \lesssim 0.25$), the system can be approximated by the one-layer shallow-water model, for which an analytical solution is available, $X_{\text t} = 2 L_0 F (1 + F/2)^{1/2}$ \citep{Hogg2006}.
+In the present case of full-depth release ($H_{0} = H_{\rm a}$), experiments and the resolution of the two-layer shallow-water model lead to $X_{\text t} \approx 8 \text{--} 12 L_{0}$~\citep{Rottman1983, Ungarish2005}.
 
-Other mechanisms can modify the dynamics of the slumping regime which could lead to other transitions of the front velocity at different positions $X_{\text t}$ or times $T_{\text t}$, prior to the influence of the rarefaction wave.
-This is basically part of the topic of the present chapter.
+Other mechanisms can modify the dynamics of the slumping regime, possibly leading to other possible positions $X_{\text t}$ or times $T_{\text t}$ at which the velocity is no longer time-independent.
+Depending on the relative intensity of these mechanisms, this can occur prior to the influence of the rarefaction wave or bore.
+Hence, determining the various regimes related to the prevalence of each mechanism is part of the topic of the present chapter.
 
@@
 A similar regime can be induced by inertial friction, for which the full derivation remains less obvious \citep[see for instance][for the case of a constant-influx current]{Hogg2001}.
-In any case, such transition can be seen as the time or position at which the current is strongly affected by dissipation, which actually encompasses entrainment for miscible gravity currents.
+In any case, such transition can be seen as the time or position at which the current is strongly affected by dissipation, which encompasses entrainment for miscible gravity currents.
 The slumping regime then ends when dissipation induced by bottom friction and/or entrainment at the miscible interface, overcomes the current dynamics.
 %
-Again using a simple idealised approach, the latter transition can be estimated from a dynamic condition balancing an estimation of the shear stress $\rho_{\text c} C_d U_{\text{c}}^2$ ($C_d$ being some drag coefficient induced by friction at the bottom and/or by entrainment at the interface) with inertia over the current height $h_{\text c}$, $\rho_{\text c} U_{\text{c}}^2 h_{\text c} /X_t$.
-In the latter expression, $h_{\text c}$ can have different scaling.
-In the case of a clearly detached current head, the constant height scales with the initial height $H_0$.
-When assuming that the current remains attached to the initial column at transition \citep[similarly to][but for inertial drag instead of viscous dissipation]{Bougouin2022}, mass conservation between the initial column and a simplified box geometry implies $h_{\text c} = H_0 L_0/X_t$.
+Again using a simple idealized approach, the latter transition can be estimated from a dynamic condition balancing an estimation of the shear stress $\rho_{\text c} C_d U_{\text{c}}^2$ ($C_d$ being some drag coefficient induced by friction at the bottom and/or by entrainment at the interface) with inertia over the current height $h_{\text c}$, $\rho_{\text c} U_{\text{c}}^2 h_{\text c} /X_t$.
+In the latter expression, $h_{\text c}$ can have different scalings.
+In the case of a detached current head, the constant height scales with the initial height $H_0$.
+When assuming that the current remains attached to the initial column at transition \citep[similar to][but for inertial drag instead of viscous dissipation]{Bougouin2022}, mass conservation between the initial column and a simplified box geometry implies $h_{\text c} = H_0 L_0/X_t$.
 %
@@
 
-Particle-Laden Gravity Currents (PLGCs) refer to as gravity currents, or density currents, are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid \citep[e.g.][]{Hopfinger1983,Middleton1993,Kneller2000,Meiburg2010,wells2021}.
+Particle-Laden Gravity Currents (PLGCs), also referred to as gravity currents or density currents, are driven by a mass difference between a heavy fluid-particle mixture and a lighter ambient liquid \citep[e.g.][]{Hopfinger1983, Middleton1993,Kneller2000, Meiburg2010,wells2021}.
 Different subclasses of PLGCs can be distinguished depending on the relative density of the different phases constituting the system; i.e.
@@
 This is the most commonly studied subclass of PLGCs, usually referred to as turbidity currents.
-At the other end, PLGCs can also be induced by a neutrally-buoyant fluid-particle mixture flowing into a lighter ambient fluid.
+At the other end, PLGCs can also be induced by a neutrally buoyant fluid-particle mixture flowing into a lighter ambient fluid.
 
 A major difference between PLGCs and compositional currents is therefore the existence of typical length and time scales associated with the particle dynamics.
-Nevertheless, the prediction of these finite scales is difficult as it shall strongly depend on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface which modifies entrainment and associated dissipation.
-In turn, this can modify the macroscopic behaviour of the current, for instance, the dynamics of the front.
+Nevertheless, the prediction of these finite scales is difficult as it should strongly depend on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface which modifies entrainment and associated dissipation.
+In turn, this can modify the macroscopic behavior of the current, for instance, the dynamics of the front.
 Accordingly, these dynamics shall be controlled by several dimensionless parameters that can lead to a wide variety of flow regimes.
@@
 
-Beyond the transition from the slumping regime to other regimes discussed in the previous section (still relevant for PLGCs although parametrisation of the processes involves more complex mechanisms induced by the fluid-particle interaction), here, a new transition towards the deposition of the particle can emerge.
+Beyond the transition from the slumping regime to other regimes discussed in the previous section (still relevant for PLGCs although parametrization of the processes involves more complex mechanisms induced by the fluid-particle interaction), here, a new transition towards the deposition of the particle can emerge.
 In particular, based on a kinematic approach balancing time scales related to current propagation and particle settling, leads to
@@
 %
-Note that such transition shall disappear for the neutrally-buoyant situation of PLGCs.
-However, it has been observed that mixing at the interface can modify the local concentration of particles at the front, resulting in behaviour similar to that of settling~\cite{Schneider2023}.
+Note that such transition shall disappear for the neutrally buoyant situation of PLGCs.
+However, it has been observed that mixing at the interface can modify the local concentration of particles at the front, resulting in behavior similar to that of settling~\cite{Schneider2023}.
 Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear.
@@
 The two contributions are simply added as described in \cite{chauchat2017}.
-The drag force is modelled using the \cite{schiller1933} closure multiplied by a hindrance function of \cite{richardson1954}.
+The drag force is modeled using the \cite{schiller1933} closure multiplied by a hindrance function of \cite{richardson1954}.
 In order to describe the influence of the slope, the total pressure is solved and the gravity vector is rotated while the mesh is kept aligned with the $x$ and $y$ axis.
@@
 They are complemented by 19 numerical simulations using SedFoam.
-Altogether, they are classified into 4 different datasets corresponding to the 3 experimental set-ups and the numerical modelling respectively.
+Altogether, they are classified into 4 different datasets corresponding to the 3 experimental set-ups and the numerical modeling respectively.
 These 4 datasets are described below.
@@
 The third dataset is gathered using setup 3.
-The experiments are performed using a neutrally-buoyant fluid-particle mixture made of polystyrene beads of typical diameter $d \sim 1~\textrm{mm}$ with density $\rho_{\text p} \sim 1038.5~\textrm{kg/m}^{3}$.
+The experiments are performed using a neutrally buoyant fluid-particle mixture made of polystyrene beads of typical diameter $d \sim 1~\textrm{mm}$ with density $\rho_{\text p} \sim 1038.5~\textrm{kg/m}^{3}$.
 Saltwater is added to the lock for density matching.
@@
 This dataset has been obtained using the two-fluid model described above.
-The domain size is $L_0+L_1=3.8$~m long and $W=0.5$~m height and the typical grid resolution is $\Delta x=1.25$~mm in the streamwise direction and it varies between $\Delta y \in [0.25 ; 3.5]$ mm in the wall-normal direction with a geometrical progression.
+The domain size is $L_0+L_1=3.8$~m long and $W=0.5$~m height and the typical grid resolution is $\Delta x=1.25$~mm in the streamwise direction and it varies between $\Delta y \in [0.25; 3.5]$ mm in the wall-normal direction with a geometrical progression.
 The time step is adaptative following criteria on the Courant number, $C_{FL}=\max(U) \Delta t / \Delta x < 0.3$ for the phase velocities as well as for the relative velocity between the phases (the maximum allowed time-step during the simulation is set to $\max(\Delta t) = 10^{-3}$~s).
 The lock is $L_0=0.3$~m wide and $H_0=0.5$~m high, corresponding to the domain dimensions.
-The numerical schemes are second order in time and space: backward for time, a blend of centred and upwind schemes for advection (limitedLinear in OpenFOAM) and a centred scheme for diffusion.
+The numerical schemes are second order in time and space: backward for time, a blend of centered and upwind schemes for advection (limitedLinear in OpenFOAM) and a centered scheme for diffusion.
 The initial condition is fluid and particle phase at rest ($u_f=u_p=0$) and the initial volume fraction of particles $\phi$ is set to a constant value in the region $x\in[-0.5 ; 0]$ m and $y\in[0;0.5]m$.
@@
 	\includegraphics{figure3.pdf}
-	\caption{(a) Current front position as a function of time for all experimental runs (b) Selected runs plotted in non-dimensional coordinates.}
+	\caption{(a) Dimensionless current front position as a function of dimensionless time for all experimental runs (b) Zoom on the early dynamics for selected runs.
+Here, $\alpha$ and $\mathcal{S}t$ increase along the black arrows.}
 	\label{fig:fig3}
@@
 The current dynamics is characterized by the evolution of several dimensionless variables.
-Here, as stated in section~\ref{sec:intro}, we aim at giving a macroscopic description of the initial current dynamics across ranges of several parameters spanning multiple orders of magnitude.
+Here, as stated in section~\ref{sec:intro}, we aim to give a macroscopic description of the initial current dynamics across ranges of several parameters spanning multiple orders of magnitude.
 The wide diversity of experiments and simulations results in a large variety of currents which does not allow for an exhaustive description of their individual dynamics.
@@
 After a short transient acceleration phase, most currents exhibit at first a constant velocity phase, which corresponds to the so-called \emph{slumping} regime discussed in section \ref{sec:intro}.
-Afterwards, the velocity becomes time-dependent, often decreasing, due to one or multiple of the attenuation processes described in section~\ref{sec:intro}.
+Afterward, the velocity becomes time-dependent, often decreasing, due to one or multiple of the attenuation processes described in section~\ref{sec:intro}.
 %
@@
 Note that the previous expansion \eqref{eq:fit_eq} assumes a constant velocity and constant deceleration trend of the front evolution.
-The latter is relevant to extract the time scale of existence of the slumping regime, i.e.
+The latter is relevant to extract the time scale of the existence of the slumping regime, i.e.
 the linear component, as long as its dynamics has only weakly deviated from such linear evolution.
@@
 This is probably due to the influence of the initial condition.
-In most of the experiments, due to the stirring, the initial condition exhibits velocity fluctuations as well as concentration fluctuations which might influence the early current behaviour.
+In most of the experiments, due to the stirring, the initial condition exhibits velocity fluctuations as well as concentration fluctuations which might influence the early current behavior.
 In the 2D RANS two-fluid simulations using sedFOAM, the velocities are zero and the concentration is homogeneous, both without any fluctuations.
@@
 \end{equation}
-where $A$, $B$ and $c_{\text d}$ are scaling constants, and $L$ accounts for the along slope distance over which the current moves during this phase.
+where $A$, $B$ and $c_{\text d}$ are scaling constants, and $L$ accounts for the alongslope distance over which the current moves during this phase.
 \citet{Cantero2007} found it to be independent of the lock aspect ratio or initial buoyancy, $L = 0.3 h_{0}$.
@@
 For $\alpha > 35^\circ$, the measured dimensionless velocities ($\mathcal{F}r$) slowly stand above equation (\ref{eq:Fr_th}) curve.
-In this region of large slope angles, it is likely that the equilibrium between the forces becomes a lot more in favour of the so-called slope effect than the pressure gradient.
+In this region of large slope angles, the equilibrium between the forces likely becomes a lot more in favor of the so-called slope effect than the pressure gradient.
 Consequently, the scaling constants may have to be adapted to this major change in the influence of the forces.
@@
 
-Altogether, the influence of $\phi$, $\mathcal{S}t$, $\alpha$ on the slumping dynamics of PLGCs characterised in the present study, highlights the importance of a better understanding of the early stage transfer of initial potential energy towards the along-slope kinetic energy.
-To predict the longtime scale evolution of PLGCs, the knowledge and the characterisation of their early-stage dynamics remain fundamental to prescribe an initial input for large-scale modelling.
+Altogether, the influence of $\phi$, $\mathcal{S}t$, $\alpha$ on the slumping dynamics of PLGCs characterized in the present study, highlights the importance of a better understanding of the early stage transfer of initial potential energy towards the along-slope kinetic energy.
+To predict the longer-time evolution of PLGCs, the knowledge and the characterization of their early-stage dynamics remain fundamental to prescribing an initial input for large-scale modeling.
 
 However, details of the current shape and related mixing remain unclear.
-In particular, even even if the front velocity is well captured by the 2D simulations, they poorly capture the current shapes, as shown in figures~\ref{fig:fig1} and~\ref{fig:fig7}.
+In particular, even if the front velocity is well captured by the 2D simulations, they poorly capture the current shapes, as shown in figures~\ref{fig:fig1} and~\ref{fig:fig7}.
 This would require 3D simulations to properly capture the turbulent mixing at the current interface.
 For validation, advanced experimental metrologies are thus required to extract accurate measurements of the turbulent mixing at the interface and responsible for ambient entrainment in the current.
-These should be an important focus for future work.
+This should be an important focus for future work.
 %

### 2024-03-01 paper/reply_to_referees/first_review/reply.tex
temp commit modifying introduction

@@
 \usepackage[dvipsnames]{xcolor}
-\newcommand{\xx}{\textcolor{WildStrawberry}}
-\newcommand{\pc}{\textcolor{ForestGreen}}
-\newcommand{\cg}{\textcolor{Mulberry}}
-\newcommand{\new}{\textcolor{blue}}
-\newcommand{\todo}{\textcolor{red}}
+\newcommand{\julien}{\textcolor{WildStrawberry}}
+\newcommand{\laurent}{\textcolor{ForestGreen}}
+\newcommand{\tocheck}{\textcolor{Mulberry}}
+\newcommand{\marie}{\textcolor{red}}
 
@@
 2017, which deals with some more complex configurations, with non-rectangular geometry, with stratified fluid, possibly in the partial depth lock release configuration.
-The various contributions should be catalogued and described in an appropriate way.}{}
+The various contributions should be catalogued and described in an appropriate way.}{The introduction has been rewritten according to various comments of the referees and other readers.
+Some references have been added.
+Note that we can not provide an extensive catalog of the complete literature related gravity currents in such an introduction, bound to have a reasonable length.
+\tocheck{TO DO.}}
 
-    \item \MinorReply{Figure 1}{the q-v panels look out of place and the sequence is horizontal while the experiments are vertical. It is recommended that they be placed in a separate figure.}{}
+    \item \MinorReply{Figure 1}{the q-v panels look out of place and the sequence is horizontal while the experiments are vertical.
+It is recommended that they be placed in a separate figure.}{\tocheck{There are already many figures compared to the text length.
+As this is properly described in the caption, we prefer to leave it as one figure presenting all configurations.}}
 
     \item \MinorReply{Figure 3a}{Figure 3a is not meaningful due to the large amount of data shown.
-It is advisable to scale it down in dimensionless form, as is done in Figure 3b for 'selected runs', which should be specified.}{}
+It is advisable to scale it down in dimensionless form, as is done in Figure 3b for 'selected runs', which should be specified.}{\tocheck{Do we do it?}}
 
-    \item \MinorReply{Figure 3b}{the alpha symbol is flanked by an arrow which appears to be meaningless. I assume that the arrow should be reversed.}{}
+    \item \MinorReply{Figure 3b}{the alpha symbol is flanked by an arrow which appears to be meaningless.
+I assume that the arrow should be reversed.}{\tocheck{The upward arrow flanking the alpha symbol means that alpha is increasing.
+We prefer to leave it this way.}}
 
-    \item \MinorReply{References}{the book by Ungarish should be better updated to the last edition.}{}
+    \item \MinorReply{References}{the book by Ungarish should be better updated to the last edition.}{Done.}
 \end{itemize}
@@
 \begin{itemize}
-    \item \MinorReply{Line 64}{Line 64 is unclear}{}
-    \item \MinorReply{Line 78}{The sentence starting in line 78 is too convoluted. Needs restructuring.}{}
-    \item \MinorReply{Line 102}{what does it mean by $h_c$ can have different scaling?}{}
-    \item \MinorReply{Lines 104--107}{It is unclear what the authors mean gt detachment. Also “similarly” may need to change to “similar”}{}
-    \item \MinorReply{Section 1.1}{gives several different expressions for $X_t$. Some clarification as to where each apply will be useful.}{}
+    \item \MinorReply{Line 64}{Line 64 is unclear}{Done.}
+    \item \MinorReply{Line 78}{The sentence starting in line 78 is too convoluted. Needs restructuring.}{Done.}
+    \item \MinorReply{Line 102}{what does it mean by $h_c$ can have different scaling?}{The two scalings are explained in the two following sentences.}
+    \item \MinorReply{Lines 104--107}{It is unclear what the authors mean gt detachment.
+Also “similarly” may need to change to “similar”}{\laurent{laurent}.
+Je serai pour enlever le scaling des colonnes lourdes non-Boussinesq.
+C'est pas clair pour moi, et je pense que ce que l'on dit est faux : dans un dam-break de courant de gravité Bussinesq, on a le courant qui reste attaché à la colonne au début tout en ayant $h_{\rm c} \propto H_{0}$.}
+    \item \MinorReply{Section 1.1}{gives several different expressions for $X_t$.
+Some clarification as to where each apply will be useful.}{The physical process associated to each of the expression of $X_{t}$ is clearly stated each time.
+Hence, each expression applies when this process dominates the other.
+Characterizing these transitions is exactly one of the purposes of the data presented here.
+This is now specified in the text.
+\tocheck{TO DO}.}
     \item \MinorReply{Lines 133-134}{where do length and time scale of particles play a role?}{}
-    \item \MinorReply{Line211-212}{what does it mean 2D vertical and turbulence averaged? Are these 2D RANS simulations? Are the 2 coordinates horizontal, since vertical is averaged?}{}
-    \item \MinorReply{In figure 2}{the color of the symbols outline must be the same. Otherwise it is hard to relate to the color legend}{}
-    \item \MinorReply{figure 2}{some of the volume fraction data goes to 1.    }{}
+    \item \MinorReply{Line 211-212}{what does it mean 2D vertical and turbulence averaged?
+Are these 2D RANS simulations?
+Are the 2 coordinates horizontal, since vertical is averaged?}{\julien{julien}}
+    \item \MinorReply{In figure 2}{the color of the symbols outline must be the same.
+Otherwise it is hard to relate to the color legend}{The color of the outline of the symbols is used to show the difference between two different kinds of runs, as explained in the caption.
+We prefer to leave it this way.}
+    \item \MinorReply{figure 2}{some of the volume fraction data goes to 1.}{\marie{marie}}
     \item \MinorReply{}{Curve fits with such a large data scatter, requires some justification}{}
-    \item \MinorReply{Line 461}{even even}{}
-    \item \MinorReply{Line 466}{This should be an important….}{}
+    \item \MinorReply{Line 461}{even even}{Done.}
+    \item \MinorReply{Line 466}{This should be an important….}{Done.}
 \end{itemize}

### 2024-03-03 paper/main.tex
new fig1

@@
 Other mechanisms can modify the dynamics of the slumping regime, possibly leading to other possible positions $X_{\text t}$ or times $T_{\text t}$ at which the velocity is no longer time-independent.
-Depending on the relative intensity of these mechanisms, this can occur prior to the influence of the rarefaction wave or bore.
+Depending on the relative intensity of these mechanisms, this may occur prior to the influence of the rarefaction wave or bore.
 Hence, determining the various regimes related to the prevalence of each mechanism is part of the topic of the present chapter.
@@
 
-A major difference between PLGCs and compositional currents is therefore the existence of typical length and time scales associated with the particle dynamics.
+A major difference between PLGCs and compositional currents is therefore the existence of typical length and time scales associated with the particle dynamics, such as settling or particle/particle interactions.
 Nevertheless, the prediction of these finite scales is difficult as it should strongly depend on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface which modifies entrainment and associated dissipation.

### 2024-03-03 paper/main.tex
Finishing text modifications

@@
 They often occur in natural and industrial situations, among which a typical situation is the release of a finite volume.
-Here, we focus on such `dam-break' situations, which are studied using lock-release devices at the laboratory scale.
+Here, we focus on such `dam-break' situations, which are studied using lock-release devices at the laboratory scale, and more specifically on Boussinesq turbidity currents generated from full-depth releases of vertical reservoirs.
 	%
@@
 This initial process and its characterization remain crucial to prescribe relevant initial conditions for large-scale predictive modeling.
-
 \end{abstract}
@@
 As they are mostly associated with unsteady and non-uniform dynamics in nature, the lock-release device appeared to be a popular canonical configuration to mimic their behavior at the laboratory scale \citep{Hacker1996,Shin2004,Nogueira2014}.
-In particular, studies often focused on characterising the front dynamics and the front shape as a macroscopic description of the underlying complex dynamics \citep{Rottman1983,Marino2005,Hogg2006}.
+It consists of generating the current by releasing a fixed volume of fluid initially locked behind a gate (see figure~\ref{fig:fig1}).
+In particular, studies often focused on characterizing the front dynamics and the front shape as a macroscopic description of the underlying complex dynamics \citep{Rottman1983,Marino2005,Hogg2006}.
 
@@
 This front relation includes two unknowns, a front velocity and a front height, and is used as a front boundary condition in the depth-averaged modeling of gravity currents \cite[e.g.][]{ungarish2020gravity}.
-Then, past studies managed to link the front velocity to the current height close to the front.
+Then, past studies managed to link the front velocity to the current height close to the front~\citep[e.g.][]{Benjamin1968, Huppert1980, Rottman1983, Ungarish2005}.
 However, the link between the front velocity and the initial conditions then becomes much more difficult to predict.
-Models have been proposed to circumvent this issue, leading for instance to \citep{Huppert1980}
+From the front law of \citet{Huppert1980}, one can use the typical scaling of the front height as $0.5 H_{0}$ to obtain :
 \begin{equation}
 	\label{eq:Froude_Huppert1980}
-	F = 0.5 \sqrt{\frac{\rho_{\text c}}{\rho_{\text a}}}\left(\frac{H_{\text a}}{H_0}\right)^{-1/3}.
+	F \simeq 0.45 \sqrt{\frac{\rho_{\text c}}{\rho_{\text a}}}\left(\frac{H_{\text a}}{H_0}\right)^{-1/3}.
 \end{equation}
 %
-The latter is relevant only when typically $H_0/H_{\text a} > 0.2$, which is the case in most experiments.
-Many configurations correspond to full-depth release ($H_{\text a}=H_0$), which leads to $F = 0.5(\rho_{\text c}/\rho_{\text a})^{1/2}$.
+Note that the latter is relevant only when typically $H_0/H_{\text a} \gtrsim 0.15$, which is the case in most experiments.
+Many configurations correspond to full-depth release ($H_{\text a}=H_0$), which leads to $F \simeq 0.45(\rho_{\text c}/\rho_{\text a})^{1/2}$.
 Even if several experiments and numerical simulations confirm such a result, the exact value of $F$ remains still unclear and probably dependent on several physical processes.
@@
 
-For a finite volume lock-release device, the opening of the gate also generates a backward propagating rarefaction wave or bore (depending on $H_{0}/H_{\rm a}$), which reflects on the wall at the end of the reservoir and then propagates towards the front~\citep[e.g.][]{ungarish2020gravity}.
-As it evolves faster than the current front, it eventually reaches it, ending the slumping regime and leading to a new regime usually referred to as the self-similar regime~\citep[e.g.][]{Rottman1983, Hogg2006}.
+For a finite volume lock-release device, the opening of the gate also generates a backward propagating rarefaction wave or bore (depending on $H_{0}/H_{\rm a}$), which reflects on the wall at the end of the reservoir and then propagates towards the front \citep[e.g.][]{ungarish2020gravity}.
+As it evolves faster than the current front, it eventually reaches it, ending the slumping regime and leading to a new regime usually referred to as the self-similar regime \citep[e.g.][]{Rottman1983, Hogg2006}.
 This is observed on the front evolution as $U_{\text{c}}$ is no longer time independent but evolves as $t^{-1/3}$.
@@
 $\rho_{\rm c}/\rho_{a}$.
-In the general case, its determination requires the numerical resolution of the two-layer shallow-water system of equations~\citep{Ungarish2005}.
-In the limit of deep currents (typically $H_{0}/H_{\rm a} \lesssim 0.25$), the system can be approximated by the one-layer shallow-water model, for which an analytical solution is available, $X_{\text t} = 2 L_0 F (1 + F/2)^{1/2}$ \citep{Hogg2006}.
-In the present case of full-depth release ($H_{0} = H_{\rm a}$), experiments and the resolution of the two-layer shallow-water model lead to $X_{\text t} \approx 8 \text{--} 12 L_{0}$~\citep{Rottman1983, Ungarish2005}.
+In the general case, its determination requires the numerical resolution of the two-layer shallow-water system of equations.
+However, in the limit of deep currents (typically $H_{0}/H_{\rm a} \lesssim 0.25$), the system can be approximated by the one-layer shallow-water model, for which an analytical solution is available, $X_{\text t} = 2 L_0 F (1 + F/2)^{1/2}$ \citep{Hogg2006}.
+In the present case of full-depth releases ($H_{0} = H_{\rm a}$), experiments and the resolution of the two-layer shallow-water model lead to $X_{\text t} \approx 8 \text{--} 12 L_{0}$ \citep{Rottman1983, Ungarish2005}.
 
@@
 Again using a simple idealized approach, the latter transition can be estimated from a dynamic condition balancing an estimation of the shear stress $\rho_{\text c} C_d U_{\text{c}}^2$ ($C_d$ being some drag coefficient induced by friction at the bottom and/or by entrainment at the interface) with inertia over the current height $h_{\text c}$, $\rho_{\text c} U_{\text{c}}^2 h_{\text c} /X_t$.
-In the latter expression, $h_{\text c}$ can have different scalings.
-In the case of a detached current head, the constant height scales with the initial height $H_0$.
-When assuming that the current remains attached to the initial column at transition \citep[similar to][but for inertial drag instead of viscous dissipation]{Bougouin2022}, mass conservation between the initial column and a simplified box geometry implies $h_{\text c} = H_0 L_0/X_t$.
-%
-Both cases respectively lead to
+In the case of Boussinesq currents, the height is relatively constant along the current, $h_{\text c} \propto H_{0}$, hence leading to
 \begin{equation}
 	\label{eq:Xt_drag}
-	\displaystyle X_{\text t}= {\frac{H_0}{C_d}}\quad \mbox{or} \quad X_{\text t}=  \sqrt{\frac{H_0 L_0}{C_d}}.
+	\displaystyle X_{\text t} \propto {\frac{H_0}{C_d}}.
 \end{equation}
+Note that this is not the case for non-Boussinesq currents, whose height significantly decreases from the position of the gate to the nose of the current \citep{Ungarish2011}.
+While the height at the nose still scales with $H_{0}$, the average height can then be obtained through mass conservation, $h_{\text c} \propto H_0 L_0/X_t$, hence leading to another possible scaling, $X_{\text t} \propto \sqrt{H_0 L_0/C_d}$ \citep[similar to][but for inertial drag instead of viscous dissipation]{Bougouin2022}.
 
@@
 	\label{eq:Xt_slope}
-	X_{\text t}=H_0\tan^{-1}{(\alpha)}.
+	X_{\text t} \propto H_0\tan^{-1}{(\alpha)}.
 \end{equation}
@@
 	\caption{\textbf{Experimental set-ups used in this study.} (a--c) Sketches of the three experimental set-ups.
-Below are shown snapshots of corresponding experimental images: (d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, ${S}t=0.027$; (i--k) Set-up 2, $\alpha = 45.4^\circ$, $\phi \sim 1.5~\%$, ${S}t=0.06 $; (l--p) Set-up 3, $\alpha = 0^\circ$, $\phi \sim 55\%$, ${S}t \equiv 0$; (q--v) SedFoam, $\alpha = 45.4^\circ$, $\phi \sim 1.7~\%$, $\mathcal{S}t=0.06$.}
+Below are shown snapshots of corresponding experimental images: (d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, ${S}t=0.027$; (i--k) Set-up 2, $\alpha = 45.4^\circ$, $\phi \sim 1.5~\%$, ${S}t=0.06 $; (l--p) Set-up 3, $\alpha = 0^\circ$, $\phi \sim 55\%$, ${S}t \equiv 0$.
+Below the horizontal black line are shown snapshots of a numerical simulation using SedFoan: (q--v) $\alpha = 45.4^\circ$, $\phi \sim 1.7~\%$, $\mathcal{S}t=0.06$.}
 	\label{fig:fig1}
@@
 Note that such transition shall disappear for the neutrally buoyant situation of PLGCs.
-However, it has been observed that mixing at the interface can modify the local concentration of particles at the front, resulting in behavior similar to that of settling~\cite{Schneider2023}.
+However, it has been observed that mixing at the interface can modify the local concentration of particles at the front, resulting in behavior similar to that of settling \cite{Schneider2023}.
 Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear.
@@
 
-Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed in the front dynamics, often referred to as the Froude number, and the front shape.
-Then prediction of the different flow regimes associated with PLGCs shall be somewhat included in these observables.
-After presenting the experimental and numerical methods used to investigate PLGCs dynamics in section~\ref{sec:methods}, we introduce the main dimensionless quantities used and macroscopic observations in section~\ref{sec:current_description}.
+Even if the dynamics of PLGCs is associated with several complex processes, a general consequence is observed in the front dynamics, often referred to as the Froude number.
+Then prediction of the different flow regimes associated with PLGCs shall be somewhat included in this observable, on which we focus the present chapter.
+After presenting in section~\ref{sec:methods} the experimental and numerical methods used to investigate the dynamics of Boussinesq finite-volume PLGCs currents generated from full-depth ($H_{0} \simeq H_{a}$) releases of vertical reservoirs ($a >1$), we introduce the main dimensionless quantities used and macroscopic observations in section~\ref{sec:current_description}.
 We discuss our results in section~\ref{sec:results} and conclude in section~\ref{sec:conclusion}.
@@
 The wide diversity of experiments and simulations results in a large variety of currents which does not allow for an exhaustive description of their individual dynamics.
-We rather choose to focus on a single variable, the front position $x_{\text{f}}$, which has been shown in the literature to be a good proxy to describe the current dynamics~\citep[e.g.][]{bonnecaze1993particle, Chowdhury2011, Adduce2012}.
+We rather choose to focus on a single variable, the front position $x_{\text{f}}$, which has been shown in the literature to be a good proxy to describe the current dynamics \citep[e.g.][]{bonnecaze1993particle, Chowdhury2011, Adduce2012}.
 
@@
 
-In the following, we fit all front propagation curves (except for dataset 3) with \eqref{eq:fit_eq} between $t/t_{0} = 5$, to avoid the initial transient phase, and $t/t_{0} = 30$, the transition time from the slumping to the inertial regime~\citep[similar to $X_{\text{t}}/L_{0}\approx 10$ as mentioned in section \ref{sec:intro_lockrelease}; see also e.g.][]{Rottman1983, Sher2015, Ottolenghi2016}.
+In the following, we fit all front propagation curves (except for dataset 3) with \eqref{eq:fit_eq} between $t/t_{0} = 5$, to avoid the initial transient phase, and $t/t_{0} = 30$, the transition time from the slumping to the inertial regime \citep[similar to $X_{\text{t}}/L_{0}\approx 10$ as mentioned in section \ref{sec:intro_lockrelease}; see also e.g.][]{Rottman1983, Sher2015, Ottolenghi2016}.
 Note that $x_{i}$ is kept constant equal to 0 for datasets 1 and 2, but had to be left free to adjust for the SedFoam simulations.
@@
 It shall be noted that this dependency of the slumping velocity on the bottom slope exists despite the slumping equilibrium being unmodified by the presence of a slope-induced acceleration (see section~\ref{sec:intro} for details).
-As suggested in \citet{Gadal2023}, the latter acceleration may instead affect the complicated dynamics of the early transient acceleration phase, involving strong vertical motions and significant interfacial friction~\citep{Cantero2007}.
+As suggested in \citet{Gadal2023}, the latter acceleration may instead affect the complicated dynamics of the early transient acceleration phase, involving strong vertical motions and significant interfacial friction \citep{Cantero2007}.
 %
@@
 For $\alpha=0^\circ$, figure~\ref{fig:fig6}(a) shows that $\mathcal{F}r$ is independent of $\phi$ up to a critical value $\phi_{c} \simeq 0.45$, above which the Froude number abruptly decreases.
-This results from the large increase in effective viscosity induced by the presence of particles~\citep{stickel2005fluid}, which probably enhances dissipation at the bottom and on the sidewalls.
+This results from the large increase in effective viscosity induced by the presence of particles \citep{stickel2005fluid}, which probably enhances dissipation at the bottom and on the sidewalls.
 Furthermore, this decrease in Froude number is associated with a reduction of the height of the current in the presence of a strong mass gradient near the front, as noted by \citet{Schneider2023}.
@@
 \end{equation}
-where $\phi_{\text m} = 0.585$ is the maximum volume fraction~\citep{boyer2011unifying}.
+where $\phi_{\text m} = 0.585$ is the maximum volume fraction \citep{boyer2011unifying}.
 %
 As shown in figure~\ref{fig:fig6}(a), the combination of equations \eqref{eq:Fr_th_eta} and \eqref{eq:Krieger} can reproduce quantitatively the observed influence of $\phi$ on the dimensionless velocity for $\alpha=0^\circ$.
-This agreement does not depend on the choice of the viscosity model, and using other available rheological models from the literature~\citep{stickel2005fluid, boyer2011unifying} leads to the same conclusions.
+This agreement does not depend on the choice of the viscosity model, and using other available rheological models from the literature \citep{stickel2005fluid, boyer2011unifying} leads to the same conclusions.
 
@@
 
-In this chapter, we summarize the effect of various dimensionless parameters on the early propagation of finite-volume turbidity currents.
+In this chapter, we summarize the effect of various dimensionless parameters on the early propagation of Boussinesq finite-volume turbidity currents generated from full-depth ($H_{0} \simeq H_{a}$) releases of vertical reservoirs ($a >1$).
 Hence, by using data from multiple experimental devices and from two-fluid numerical simulations, we explore a wide range of $\phi$, $\alpha$, $\mathcal{S}t$ and $\mathcal{R}e$.

### 2024-03-03 paper/main.tex
Finishing up review and sending to Laurent

@@
 
+\begin{figure*}
+	\centering
+	\includegraphics{figure4.pdf}
+	\caption{\textbf{Influence of the bottom slope} on the current Froude number.
+Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
+The black plain line indicates the prediction of \eqref{eq:Fr_th} for $Fr_{0} = 0.45$, $C_{\text D} = 0.02$ and $S = 0.25$.
+The black dashed line indicates the fit to the data from the numerical simulation of \citet{Birman2007} for saline homogeneous gravity currents.}
+	\label{fig:fig4}
+\end{figure*}
+
 In this section, we focus on the influence of the bottom slope $\alpha$ on the current dynamics.
@@
 %
-One can write an energetic balance between the initial state and the end of this early acceleration phase \citep{Gadal2023}
+One can write an energetic balance for the entire current between its initial state and the end of this early acceleration phase \citep{Gadal2023}
 \begin{equation}
@@
 \end{equation}
-where $A$, $B$ and $c_{\text d}$ are scaling constants, and $L$ accounts for the alongslope distance over which the current moves during this phase.
+where $A$ and $B$ are scaling constants, $c_{\text d}$ is a drag coefficient typically on the order of $10^{-2}$ -- $10^{-3}$ \citep{hogg2004effects}, and $L$ accounts for the alongslope distance over which the current moves during this phase.
 \citet{Cantero2007} found it to be independent of the lock aspect ratio or initial buoyancy, $L = 0.3 h_{0}$.
@@
 where $Fr_0$, $C_{\text D}$ and $S$ are new constants depending on $A$, $B$, $c_{\text d}$ and $D$.
-As shown in Fig.~\ref{fig:fig4}, this simple model is able to reproduce the observed variation of the dimensionless velocity in the bottom slope in the range [$5^\circ$, $35^\circ$].
+Note that $C_{\rm d}$ has little influence if $c_{\text d}$ is kept within the reasonable values mentioned previously, and $Fr_0$ is strongly restricted by the values in $\theta = 0^\circ$, such that the only truly adjustable parameter is $S$.
+As shown in Fig.~\ref{fig:fig4}, this simple dimensional model is able to reproduce the observed variation of the dimensionless velocity in the bottom slope in the range [$5^\circ$, $35^\circ$] while keeping the scaling constants within reasonable values.
 %
@@
 
-\begin{figure*}[ht]
-	\centering
-	\includegraphics{figure4.pdf}
-	\caption{\textbf{Influence of the bottom slope} on the current Froude number.
-Plain symbols correspond to $\phi < 0.45$, while transparent symbols are for $\phi \geq 0.45$ (see section~\ref{sec:influence_phi} for details on the influence of $\phi$).
-The black plain line indicates the prediction of \eqref{eq:Fr_th} for $Fr_{0} = 0.5$, $C_{\text D} = 0.4$ and $S = 0.25$.
-The black dashed line indicates the fit to the data from the numerical simulation of \citet{Birman2007} for saline homogeneous gravity currents.}
-	\label{fig:fig4}
-\end{figure*}
-
 \subsection{Influence of the Stokes number, $\mathcal{S}t$}
@@
 
-\begin{figure*}[ht]
+\begin{figure*}
 	\centering
@@
 The inset in (a) shows the plot in linear scale.
-The black line indicates the prediction of \eqref{eq:Fr_th_eta} for $Fr_{0} = 0.5$, $C_{\text D} = 0.4$, $S = 0.25$ and $Re_{\text c} = 500$.}
+The black line indicates the prediction of \eqref{eq:Fr_th_eta} for $Fr_{0} = 0.45$, $C_{\text D} = 0.02$, $S = 0.25$ and $Re_{\text c} = 500$.}
 	\label{fig:fig6}
@@
 %
+Here, the only adjustable parameter is $Re_{\text c}$ as all the other parameters have been fixed previously.
 As shown in figure~\ref{fig:fig6}(a), the combination of equations \eqref{eq:Fr_th_eta} and \eqref{eq:Krieger} can reproduce quantitatively the observed influence of $\phi$ on the dimensionless velocity for $\alpha=0^\circ$.

### 2024-03-03 paper/reply_to_referees/first_review/reply.tex
Finishing up review and sending to Laurent

@@
 Gadal \emph{et al}, for publication in a book for the \emph{Geophysical Monograph Series}.
-We are grateful for the thoughtful referee reports, and for their comments and suggestions.
+We are grateful for the thoughtful referee reports and for their comments and suggestions.
 \\
@@
 Some references have been added.
-Note that we can not provide an extensive catalog of the complete literature related gravity currents in such an introduction, bound to have a reasonable length.
-\tocheck{TO DO.}}
+Note that we can not provide an extensive catalog of the complete literature related to gravity currents in such an introduction, bound to have a manuscript of reasonable length.
+We have then focused on the literature that shares the most scientific objectives with the present contribution.}
 
     \item \MinorReply{Figure 1}{the q-v panels look out of place and the sequence is horizontal while the experiments are vertical.
-It is recommended that they be placed in a separate figure.}{\tocheck{There are already many figures compared to the text length.
-As this is properly described in the caption, we prefer to leave it as one figure presenting all configurations.}}
+It is recommended that they be placed in a separate figure.}{There are already many figures compared to the text length, and we do not wish to emphasize the numerical simulations by adding a dedicated figure.
+Hence, we modified figure 1 to better separate the q-v panels from the rest.}
 
     \item \MinorReply{Figure 3a}{Figure 3a is not meaningful due to the large amount of data shown.
-It is advisable to scale it down in dimensionless form, as is done in Figure 3b for 'selected runs', which should be specified.}{\tocheck{Do we do it?}}
+It is advisable to scale it down in dimensionless form, as is done in Figure 3b for 'selected runs', which should be specified.}{Done.}
 
     \item \MinorReply{Figure 3b}{the alpha symbol is flanked by an arrow which appears to be meaningless.
-I assume that the arrow should be reversed.}{\tocheck{The upward arrow flanking the alpha symbol means that alpha is increasing.
-We prefer to leave it this way.}}
+I assume that the arrow should be reversed.}{The flanking arrows have been removed, and the figure caption has been modified accordingly.}
 
@@
     \item \MinorReply{Line 64}{Line 64 is unclear}{Done.}
+
     \item \MinorReply{Line 78}{The sentence starting in line 78 is too convoluted. Needs restructuring.}{Done.}
-    \item \MinorReply{Line 102}{what does it mean by $h_c$ can have different scaling?}{The two scalings are explained in the two following sentences.}
+
+    \item \MinorReply{Line 102}{what does it mean by $h_c$ can have different scaling?}{This sentence has been removed, and this part of the text re-written for the sake of clarity.}
+
     \item \MinorReply{Lines 104--107}{It is unclear what the authors mean gt detachment.
-Also “similarly” may need to change to “similar”}{\laurent{laurent}.
-Je serai pour enlever le scaling des colonnes lourdes non-Boussinesq.
-C'est pas clair pour moi, et je pense que ce que l'on dit est faux : dans un dam-break de courant de gravité Bussinesq, on a le courant qui reste attaché à la colonne au début tout en ayant $h_{\rm c} \propto H_{0}$.}
+Also “similarly” may need to change to “similar”}{This part of the text has been re-written for the sake of clarity.}
+
     \item \MinorReply{Section 1.1}{gives several different expressions for $X_t$.
-Some clarification as to where each apply will be useful.}{The physical process associated to each of the expression of $X_{t}$ is clearly stated each time.
-Hence, each expression applies when this process dominates the other.
-Characterizing these transitions is exactly one of the purposes of the data presented here.
-This is now specified in the text.
-\tocheck{TO DO}.}
-    \item \MinorReply{Lines 133-134}{where do length and time scale of particles play a role?}{}
+Some clarification as to where each apply will be useful.}{The physical process associated with each expression of $X_{t}$ is stated along the section when $X_{t}$ is defined.
+Now, the first transition to occur will depend on the prevailing mechanism, and as such on where the experimental run is in the parameter space.
+This is now clarified in the introduction.}
+
+    \item \MinorReply{Lines 133-134}{where do length and time scale of particles play a role?}{This sentence has been modified by adding examples.}
+
     \item \MinorReply{Line 211-212}{what does it mean 2D vertical and turbulence averaged?
@@
 Are the 2 coordinates horizontal, since vertical is averaged?}{\julien{julien}}
+
     \item \MinorReply{In figure 2}{the color of the symbols outline must be the same.
@@
 We prefer to leave it this way.}
+
     \item \MinorReply{figure 2}{some of the volume fraction data goes to 1.}{\marie{marie}}
-    \item \MinorReply{}{Curve fits with such a large data scatter, requires some justification}{}
+
+    \item \MinorReply{}{Curve fits with such a large data scatter, requires some justification}{Here, we do not fit a proper analytical model or check the validity of its prediction against the data.
+We merely aim to show that a simple dimensional model incorporating the required mechanisms is able to reproduce the observed variation of the data while keeping the scaling constants within reasonable values.
+In addition, we now clarify that the only true adjustable parameters are $S$ for the slope effect and $Re_{c}$ for the effect of the volume fraction, as $C_{\rm d}$ and $Fr_{0}$ are strongly bound by values in the literature.}
+
     \item \MinorReply{Line 461}{even even}{Done.}
+
     \item \MinorReply{Line 466}{This should be an important….}{Done.}

### 2024-03-03 paper/figures/figure_scripts/figure4.py

@@ -68,7 +68,7 @@ markers[dataset_idx == "4"] = "s"
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 facecolors = np.array([to_rgba(c, a) for c, a in zip(facecolors, alphas)])
 
-mask_nosuspended = (authors == "Rastello") & (H0 / Ha < 1)
+mask_nosuspended = (authors == "Marie Rastello") & (H0 / Ha < 1)
 edgecolors = np.array([to_rgba("k", a) for a in alphas])
 edgecolors[mask_nosuspended] = np.array([to_rgba("tab:red", 0.4) for a in alphas[mask_nosuspended]])
 
@@ -82,7 +82,7 @@ plot_idxs = np.lexsort((random_order, zorders))
 phi_m = 0.585
 alpha_plot = np.linspace(0, 50, 500)
 eta = 0
-Fr_th = Froude(alpha_plot, eta, 7e4)
+Fr_th = Froude(alpha_plot, eta, 7e4, Cd=0.02, Fr0=0.45)
 
 # %% figure
 fig, ax = plt.subplots(1, 1, figsize=tp.large_figure_size, constrained_layout=True)

### 2024-03-03 paper/figures/figure_scripts/figure5.py

@@ -10,49 +10,65 @@ from lmfit import Model
 from matplotlib.colors import to_rgba
 from netCDF4 import Dataset
 
-plt.rcParams['figure.constrained_layout.hspace'] = 0
-plt.rcParams['figure.constrained_layout.h_pad'] = 0.0005
+plt.rcParams["figure.constrained_layout.hspace"] = 0
+plt.rcParams["figure.constrained_layout.h_pad"] = 0.0005
 
 
 def lambda_var(x, a, c, th):
-    return np.piecewise(x, [x < th, x >= th], [lambda x: c, lambda x: a*(x - th) + c])
+    return np.piecewise(x, [x < th, x >= th], [lambda x: c, lambda x: a * (x - th) + c])
 
 
 # %% Load data
-path_data = '../../../data/output_data'
-list_runs = glob.glob(os.path.join(path_data, '*.nc'))
+path_data = "../../../data/output_data"
+list_runs = glob.glob(os.path.join(path_data, "*.nc"))
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% create data vectors
-alpha, phi, St, H0, L0, Fr, lamb = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
-                                              d.variables['St'][:].data, d.variables['H0'][:].data,
-                                              d.variables['L0'][:].data, d.variables['Fr'][:].data,
-                                              d.variables['lamb'][:].data
-                                              ] for d in datasets]).T
+alpha, phi, St, H0, L0, Fr, lamb = np.array(
+    [
+        [
+            d.variables["alpha"][:].data,
+            d.variables["phi"][:].data,
+            d.variables["St"][:].data,
+            d.variables["H0"][:].data,
+            d.variables["L0"][:].data,
+            d.variables["Fr"][:].data,
+            d.variables["lamb"][:].data,
+        ]
+        for d in datasets
+    ]
+).T
 
-Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
-              else d.variables['H0'][:].data for d in datasets])
-authors, particles = np.array([[d.author, d.particle_type,
[…]

### 2024-03-03 paper/figures/figure_scripts/figure6.py

@@ -68,7 +68,7 @@ plot_idxs = np.lexsort((random_order, zorders))
 phi_m = 0.585
 phi_plot = np.logspace(-3, phi_m, 200)
 eta = Krieger_viscosity(phi_plot, phi_m)
-Fr_th0 = Froude(0, eta, 7e4)
+Fr_th0 = Froude(0, eta, 7e4, Cd=0.02, Fr0=0.45)
 Fr_th45 = Froude(45, eta, 7e4)
 
 # %% masks for plot

### 2024-03-03 paper/figures/figure_scripts/figure2.py
Update figure BW and run evey figure to be sure

@@ -54,7 +54,7 @@ markers[dataset_idx == "4"] = "s"
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 edgecolors = np.full_like(facecolors, "k")
 
-mask_nosuspended = (authors == "Rastello") & (H0 / Ha < 1)
+mask_nosuspended = (authors == "Marie Rastello") & (H0 / Ha < 1)
 edgecolors[mask_nosuspended] = "tab:red"
 
 zorders = np.vectorize(lambda dataset: tp.dataset_zorder[dataset])(dataset_idx)

### 2024-03-03 paper/figures/figure_scripts/figure5.py

@@ -66,7 +66,7 @@ markers[dataset_idx == "4"] = "s"
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 facecolors = np.array([to_rgba(c, a) for c, a in zip(facecolors, alphas)])
 
-mask_nosuspended = (authors == "Rastello") & (H0 / Ha < 1)
+mask_nosuspended = (authors == "Marie Rastello") & (H0 / Ha < 1)
 edgecolors = np.array([to_rgba("k", a) for a in alphas])
 edgecolors[mask_nosuspended] = np.array([to_rgba("tab:red", 0.4) for a in alphas[mask_nosuspended]])
 

### 2024-03-03 paper/figures/figure_scripts/figure6.py

@@ -55,7 +55,7 @@ markers[dataset_idx == "4"] = "s"
 facecolors = np.vectorize(tp.color_datasets.get)(dataset_idx)
 edgecolors = np.full_like(facecolors, "k")
 
-mask_nosuspended = (authors == "Rastello") & (H0 / Ha < 1)
+mask_nosuspended = (authors == "Marie Rastello") & (H0 / Ha < 1)
 edgecolors[mask_nosuspended] = "tab:red"
 
 zorders = np.vectorize(lambda dataset: tp.dataset_zorder[dataset])(dataset_idx)

### 2024-03-03 paper/figures_B&W/figure_scripts_B&W/figure1.py

@@ -6,22 +6,21 @@ import matplotlib.pyplot as plt
 import matplotlib.transforms as mtransforms
 import numpy as np
 import template as tp
-from matplotlib.colors import to_rgba
 from matplotlib.path import Path
 from template import to_grayscale
 
-plt.rcParams['figure.constrained_layout.hspace'] = 0
-plt.rcParams['figure.constrained_layout.wspace'] = 0
-plt.rcParams['figure.constrained_layout.h_pad'] = 0.005
-plt.rcParams['figure.constrained_layout.w_pad'] = 0.005
+plt.rcParams["figure.constrained_layout.hspace"] = 0
+plt.rcParams["figure.constrained_layout.wspace"] = 0
+plt.rcParams["figure.constrained_layout.h_pad"] = 0.005
+plt.rcParams["figure.constrained_layout.w_pad"] = 0.005
 
 
 def sind(x):
-    return np.sin(x*np.pi/180.)
+    return np.sin(x * np.pi / 180.0)
 
 
 def cosd(x):
-    return np.cos(x*np.pi/180.)
+    return np.cos(x * np.pi / 180.0)
 
 
 def Rotation_matrix(theta):
@@ -32,51 +31,51 @@ def Rotation_matrix(theta):
 # ## Sketches parameters
 # colors
 # color_water = 'aliceblue'
-color_water = to_grayscale('lightcyan')
+color_water = to_grayscale("lightcyan")
 # color_water_salt = 'lightgreen'
-color_water_salt = 0.9*to_grayscale('#F3F7D4')
-color_sed = to_grayscale('peru')
-color_walls = 'k'
+color_water_salt = 0.9 * to_grayscale("#F3F7D4")
+color_sed = to_grayscale("peru")
+color_walls = "k"
 alpha_water = 1
 
 # dimension parameters
 tank_height = 45  # cm
 tank_length = 165  # cm
 door_pos = 15  # cm
-door_height = 1.1*tank_height
-door_pad = 0*tank_height
+door_height = 1.1 * tank_height
+door_pad = 0 * tank_height
 y_bottom = 0  # cm, at the end of the canal
 x_bottom = 0  # cm, at the end of the canal
 mixing_height = tank_height
-mixing_pad = 0.2*tank_height
-mixing_width = 0.15*mixing_height
+mixing_pad = 0.2 * tank_height
+mixing_width = 0.15 * mixing_height
 
 
[…]

### 2024-03-03 paper/figures_B&W/figure_scripts_B&W/figure3.py

@@ -4,104 +4,137 @@ import sys
 
 import matplotlib.pyplot as plt
 import matplotlib.transforms as mtransforms
-from matplotlib.lines import Line2D
 import numpy as np
 import template as tp
-from netCDF4 import Dataset
 from lmfit.model import load_modelresult
+from matplotlib.lines import Line2D
+from netCDF4 import Dataset
+
+
+def polyFULL(t, Fr, lamb, xi, c, d):
+    return xi + Fr * (t - lamb * t**2 + c * t**3 + d * t**4)
 
 
 # %% Load data
-path_data = '../../../data/output_data'
-list_runs = sorted(glob.glob(os.path.join(path_data, '*.nc')))
-list_fitresults = sorted(glob.glob(os.path.join(path_data, 'fitresult*')))
+path_data = "../../../data/output_data"
+list_runs = sorted(glob.glob(os.path.join(path_data, "*.nc")))
+list_fitresults = sorted(glob.glob(os.path.join(path_data, "fitresult*")))
 
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% mask data
 particles = np.array([d.particle_type for d in datasets]).T
-mask = (particles != 'saline water')
+mask = particles != "saline water"
 
-zorder_setups = {'Cyril': -9, 'Rastello': -10,
-                 'Jean': -6, 'Julien': -7, 'Cyril/Marie': -10}
+zorder_setups = {
+    "Cyril Gadal": -9,
+    "Marie Rastello": -10,
+    "Jean Schneider": -6,
+    "Julien Chauchat": -7,
+    "Marie Rastello - Cyril Gadal": -10,
+}
 
 # %% figure
-layout = [['legend', 'legend'], ['(a)', '(b)']]
-fig, axarr = plt.subplot_mosaic(layout, figsize=(tp.large_figure_width, 0.55*tp.large_figure_width), constrained_layout=True,
-                                gridspec_kw={'height_ratios': [0.001, 1]})
+layout = [["legend", "legend"], ["(a)", "(b)"]]
+fig, axarr = plt.subplot_mosaic(
+    layout,
+    figsize=(tp.large_figure_width, 0.55 * tp.large_figure_width),
+    constrained_layout=True,
+    gridspec_kw={"height_ratios": [0.001, 1]},
+)
 
 # All runs
-ax = axarr['(a)']
[…]

### 2024-03-03 paper/figures_B&W/figure_scripts_B&W/figure4.py

@@ -9,7 +9,7 @@ from matplotlib.colors import to_rgba
 from models import Birman, Froude
 from netCDF4 import Dataset
 
-plt.rcParams['xtick.top'] = False
+plt.rcParams["xtick.top"] = False
 
 
 def ang2sin(ang):
@@ -21,24 +21,40 @@ def sin2ang(sin):
 
 
 # %% Load data
-path_data = '../../../data/output_data'
-list_runs = glob.glob(os.path.join(path_data, '*.nc'))
+path_data = "../../../data/output_data"
+list_runs = glob.glob(os.path.join(path_data, "*.nc"))
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% create data vectors
-alpha, phi, Fr, St, H0 = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
-                                    d.variables['Fr'][:].data, d.variables['St'][:].data,
-                                    d.variables['H0'][:].data,
-                                    ] for d in datasets]).T
+alpha, phi, Fr, St, H0 = np.array(
+    [
+        [
+            d.variables["alpha"][:].data,
+            d.variables["phi"][:].data,
+            d.variables["Fr"][:].data,
+            d.variables["St"][:].data,
+            d.variables["H0"][:].data,
+        ]
+        for d in datasets
+    ]
+).T
 
-Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
-              else d.variables['H0'][:].data for d in datasets])
-authors, particles = np.array([[d.author, d.particle_type,
-                                ] for d in datasets]).T
+Ha = np.array(
+    [d.variables["H_a"][:].data if "H_a" in d.variables.keys() else d.variables["H0"][:].data for d in datasets]
+)
+authors, particles = np.array(
+    [
+        [
+            d.author,
+            d.particle_type,
+        ]
+        for d in datasets
+    ]
+).T
 
 # %% graphic specifications
 # %% masks for plot
-mask_phi = (phi < tp.phi_c)
+mask_phi = phi < tp.phi_c
 
[…]

### 2024-03-03 paper/figures_B&W/figure_scripts_B&W/figure6.py

@@ -6,40 +6,57 @@ import matplotlib.pyplot as plt
 import matplotlib.transforms as mtransforms
 import numpy as np
 import template as tp
-from netCDF4 import Dataset
 from models import Froude, Krieger_viscosity
+from netCDF4 import Dataset
 
-plt.rcParams['figure.constrained_layout.hspace'] = 0
-plt.rcParams['figure.constrained_layout.h_pad'] = 0.005
+plt.rcParams["figure.constrained_layout.hspace"] = 0
+plt.rcParams["figure.constrained_layout.h_pad"] = 0.005
 
 
 # %% Load data
-path_data = '../../../data/output_data'
-list_runs = glob.glob(os.path.join(path_data, '*.nc'))
+path_data = "../../../data/output_data"
+list_runs = glob.glob(os.path.join(path_data, "*.nc"))
 datasets = np.array([Dataset(run) for run in list_runs])
 
 # %% create data vectors
-alpha, phi, St, H0, L0, Fr = np.array([[d.variables['alpha'][:].data, d.variables['phi'][:].data,
-                                        d.variables['St'][:].data, d.variables['H0'][:].data,
-                                        d.variables['L0'][:].data, d.variables['Fr'][:].data,
-                                        ] for d in datasets]).T
+alpha, phi, St, H0, L0, Fr = np.array(
+    [
+        [
+            d.variables["alpha"][:].data,
+            d.variables["phi"][:].data,
+            d.variables["St"][:].data,
+            d.variables["H0"][:].data,
+            d.variables["L0"][:].data,
+            d.variables["Fr"][:].data,
+        ]
+        for d in datasets
+    ]
+).T
 
-Ha = np.array([d.variables['H_a'][:].data if 'H_a' in d.variables.keys()
-              else d.variables['H0'][:].data for d in datasets])
-authors, particles = np.array([[d.author, d.particle_type,
-                                ] for d in datasets]).T
+Ha = np.array(
[…]

### 2024-03-03 paper/figures_B&W/figure_scripts_B&W/template.py

@@ -11,19 +11,19 @@ from scipy.constants import golden
 
 
 def to_grayscale(color):
-    return np.array([np.sum(to_rgb(color))/3 for i in range(3)])
+    return np.array([np.sum(to_rgb(color)) / 3 for i in range(3)])
 
 
 # %% Default parameters (matplotlib.rcParams)
-style.use('tableau-colorblind10')
+style.use("tableau-colorblind10")
 
 # %% Loading custom style
-plt.style.use(os.path.join(os.path.dirname(__file__), 'src/style.mplstyle'))
+plt.style.use(os.path.join(os.path.dirname(__file__), "src/style.mplstyle"))
 
 # %%
 # Defining some constants
 inches_per_cm = 0.3937
-regular_aspect_ratio = 1/golden
+regular_aspect_ratio = 1 / golden
 
 
 # %% Figure sizing
@@ -31,17 +31,16 @@ regular_aspect_ratio = 1/golden
 full_text_width = 492.982  # page width, in pt
 column_width = 240.089  # one column width, in pt
 
-large_figure_width = full_text_width*0.35136*0.1*inches_per_cm  # in inches
-half_figure_width = column_width*0.35136*0.1*inches_per_cm  # in inches
+large_figure_width = full_text_width * 0.35136 * 0.1 * inches_per_cm  # in inches
+half_figure_width = column_width * 0.35136 * 0.1 * inches_per_cm  # in inches
 
-half_figure_size = np.array([1, regular_aspect_ratio])*half_figure_width
-large_figure_size = np.array([1, regular_aspect_ratio])*large_figure_width
+half_figure_size = np.array([1, regular_aspect_ratio]) * half_figure_width
+large_figure_size = np.array([1, regular_aspect_ratio]) * large_figure_width
 
-half_figure_size_inv = np.array([1, 1/regular_aspect_ratio])*half_figure_width
-large_figure_size_inv = np.array(
-    [1, 1/regular_aspect_ratio])*large_figure_width
+half_figure_size_inv = np.array([1, 1 / regular_aspect_ratio]) * half_figure_width
+large_figure_size_inv = np.array([1, 1 / regular_aspect_ratio]) * large_figure_width
 
-rcParams['figure.figsize'] = large_figure_size
+rcParams["figure.figsize"] = large_figure_size
 
 # %% colors
 
[…]

### 2024-03-06 paper/main.tex
small modifs

@@
 where $Fr_0$, $C_{\text D}$ and $S$ are new constants depending on $A$, $B$, $c_{\text d}$ and $D$.
-Note that $C_{\rm d}$ has little influence if $c_{\text d}$ is kept within the reasonable values mentioned previously, and $Fr_0$ is strongly restricted by the values in $\theta = 0^\circ$, such that the only truly adjustable parameter is $S$.
+Note that $C_{\rm d}$ has little influence if $c_{\text d}$ is kept within the reasonable values mentioned previously, and $Fr_0$ is fixed by the values in $\theta = 0^\circ$ (also corresponding to previous literature), such that the only truly adjustable parameter is $S$.
 As shown in Fig.~\ref{fig:fig4}, this simple dimensional model is able to reproduce the observed variation of the dimensionless velocity in the bottom slope in the range [$5^\circ$, $35^\circ$] while keeping the scaling constants within reasonable values.

### 2024-03-14 paper/reply_to_referees/reply.tex
small modifs
Last modifs

--- 2024-03-03 paper/reply_to_referees/first_review/reply.tex
@@
 \newpage
-\section{Reply to referee 1}
+\section{Reply to referee 2}
 
@@
 Are these 2D RANS simulations?
-Are the 2 coordinates horizontal, since vertical is averaged?}{\julien{julien}}
+Are the 2 coordinates horizontal, since vertical is averaged?}{Indeed, 2D vertical and turbulence-averaged means 2D RANS.
+Only turbulence is averaged the wall-normal direction is not.
+The sentence has been modified to clarify this.}
 
@@
 
-    \item \MinorReply{figure 2}{some of the volume fraction data goes to 1.}{\marie{marie}}
+    \item \MinorReply{figure 2}{some of the volume fraction data goes to 1.}{Indeed, hydrogel particles are deformable.
+As a consequence, at rest, they experience deformation under the weight of the particles above, hence resulting in denser packing compared to hard spheres, close to 1 for the smallest hydrogel particles used here.
+This was already mentioned in the description of Dataset 2, so we did not modify the text.}
 

### 2024-03-14 paper/main.tex

@@
 This model available for download at Zenodo \citep{bonamy2023} is implemented in the open-source computational fluid dynamics toolbox OpenFOAM and has been extensively validated for various sediment transport problems such as sheet-flow \citep{cheng2016,chauchat2022,mathieu2021,mathieu2022}, scour \citep{nagel2020,tsai2022}, ripples \citep{salimi2021a,salimi2021b}, avalanches \citep{montella2021} and granular collapse \citep{montella2023}.
-The simulations presented herein are 2D vertical and turbulence-averaged.
+The simulations presented herein are 2D Reynolds-averaged considering the streamwise and wall-normal direction.
 The turbulence model is a $k-\epsilon$ model with additional two-phase flow terms for density stratification and turbulent drag dissipation.
@@
 %
-Finally, we would like to acknowledge the contributors of the open-source numerical Python libraries, including Matplotlib \citep{Hunter2007}, NumPy \citep{Harris2020} and SciPy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
+We also acknowledge the contributors of the open-source numerical Python libraries, including Matplotlib \citep{Hunter2007}, NumPy \citep{Harris2020} and SciPy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
+%
+Finally, we also would like to acknowledge the two anonymous reviewers for their comments and Prof. Marius Ungarish for his extensive suggestions.
 

### 2024-03-14 paper/main.tex
update repo for publication
update doi

Fixing typo

Add ref sedfoam

small modif

@@
 However, the link between the front velocity and the initial conditions then becomes much more difficult to predict.
-From the front law of \citet{Huppert1980}, one can use the typical scaling of the front height as $0.5 H_{0}$ to obtain :
+From the front law of \citet{Huppert1980}, one can use the typical scaling of the front height as $0.5 H_{0}$ to obtain:
 \begin{equation}
@@
 Below are shown snapshots of corresponding experimental images: (d--h) Set-up 1, $\alpha = 7^\circ$, $\phi \sim 3~\%$, ${S}t=0.027$; (i--k) Set-up 2, $\alpha = 45.4^\circ$, $\phi \sim 1.5~\%$, ${S}t=0.06 $; (l--p) Set-up 3, $\alpha = 0^\circ$, $\phi \sim 55\%$, ${S}t \equiv 0$.
-Below the horizontal black line are shown snapshots of a numerical simulation using SedFoan: (q--v) $\alpha = 45.4^\circ$, $\phi \sim 1.7~\%$, $\mathcal{S}t=0.06$.}
+Below the horizontal black line are shown snapshots of a numerical simulation using SedFoam: (q--v) $\alpha = 45.4^\circ$, $\phi \sim 1.7~\%$, $\mathcal{S}t=0.06$.}
 	\label{fig:fig1}
@@
 This model available for download at Zenodo \citep{bonamy2023} is implemented in the open-source computational fluid dynamics toolbox OpenFOAM and has been extensively validated for various sediment transport problems such as sheet-flow \citep{cheng2016,chauchat2022,mathieu2021,mathieu2022}, scour \citep{nagel2020,tsai2022}, ripples \citep{salimi2021a,salimi2021b}, avalanches \citep{montella2021} and granular collapse \citep{montella2023}.
-The simulations presented herein are 2D Reynolds-averaged considering the streamwise and wall-normal direction.
+The simulations presented herein are 2D Reynolds-averaged (2D-RANS) in the streamwise and wall-normal direction.
 The turbulence model is a $k-\epsilon$ model with additional two-phase flow terms for density stratification and turbulent drag dissipation.
